### 2023-05-29 paper/paper.tex
feat(paper): Add paper

\documentclass[conference,final,a4paper]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{minted}
\usepackage{multicol}
\usepackage{microtype}
\usepackage[hyphens]{url}

\usepackage{pdfcomment}
\newcommand{\todo}[1]{\pdfcomment[color={0.045 0.278 0.643},icon=Note]{#1}}

\begin{document}

\title{Improving reproducibility of scientific software using Nix/NixOS}

\author{\IEEEauthorblockN{Max Hausch}
\IEEEauthorblockA{Stuttgart, Germany \\
st175425@stud.uni-stuttgart.de}
\and
\IEEEauthorblockN{Simon Hauser}
\IEEEauthorblockA{Stuttgart, Germany \\
st148883@stud.uni-stuttgart.de}
}

\maketitle

\begin{abstract}
  Ensuring the reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
  However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations and software environments.
  In this paper, we present a possible solution for these challenges by utilizing Nix and NixOS.
  Nix is a package manager and functional language that allows to mitigate these problems by guaranteeing that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
  NixOS is a purely functional Linux distribution, built on top of Nix that enables the build of reproducible systems including configuration files, packages and their dependencies.
  We present a case study on improving the reproducibility of preCICE, an open-source coupling library, and some of its main adapters using Nix and NixOS.
  Using this approach, we demonstrate how to create a reproducible and self-contained environment for preCICE and highlight the benefits of using Nix and NixOS for managing software and system configurations, resulting in improved reproducibility.
  In addition, we compare the performance and usability of Nix with two widely used high-performance computing (HPC) solutions, Spack and EasyBuild.
  This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.
\end{abstract}

\begin{IEEEkeywords}
  Reproducibility, Nix, NixOS, Spack, EasyBuild, preCICE, HPC.
\end{IEEEkeywords}

\section{Introduction}

In scientific research, it's crucial to be able to reproduce and verify results.
Reproducibility ensures that experiments can be repeated and findings can be validated, which is essential for reliable and credible research.
Being able to replicate experiments and computations is important for verifying scientific claims.
However, achieving reproducibility in scientific software has been a challenge due to complex dependencies, conflicting software environments, and changing software systems.
Problems arise from dependencies, library versions, and system configurations, leading to inconsistencies across different computing environments.
Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are time-consuming, prone to errors, and not scalable.
This calls for a more efficient and automated solution.

In the past years, the Nix package manager\cite{Dolstra_2004} and NixOS\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising solutions to address these challenges.
This paper explores the use of Nix/NixOS to improve the reproducibility of scientific software, focusing on a case study of preCICE\cite{preCICEv2} and its adapters and solvers.

In high-performance computing (HPC), where performance and efficiency are critical, managing software dependencies and configurations becomes even more challenging.
By using Nix, researchers and practitioners in HPC could easily reproduce computational experiments, ensuring that the same software stack, libraries, and configurations are utilized consistently.
Reproducibility of scientific results can be achieved using the Nix package manager as described by Devresse et al.\cite{Devresse_2015}
This not only streamlines the deployment process but also facilitates collaboration and sharing of software environments, making it easier for researchers to validate and build upon each other's work.
Additionally, because packages in Nix are highly customizable, optimized builds for specific HPC clusters can be realized.

We also compare Nix, EasyBuild, and Spack as package managers in the realm of scientific software and if Nix is feasible for being used in a scientific context at all.

There are several other papers which discuss and highlight the importance of reproducibly of scientific experiments.
In their paper, Dalle lists technical factors that lead to issues regarding reproducibly of in silico simulation experiments.\cite{Dalle_2012}
\begin{enumerate}
    \item software bugs: if bugs get fixed in future versions of the same software, they may distort results
    \item software availability: loosing track of old software versions may lead to unreproducible experiments
    \item floating points numbers: due to interpretation errors of floating point numbers different simulators could produce different results
    \item computer and operating systems evolutions: evolving software stacks and dependencies could lead to researchers not being able to reproduce results of an experiment
\end{enumerate}
Nix could probably solve three out of these four factors:
\begin{enumerate}
    \item using Nix, software --- including its bugs --- can be built reproducibly
    \item using Nix, even old software can be rebuild and used\footnote{\url{https://blinry.org/nix-time-travel/}}
    \item Nix cannot really help with floating point issues that affect different codes
    \item using NixOS, entire operating systems can be reproducibly built, including complete dependencies and kernels
\end{enumerate}

One suggestion to enhance reproducibility by using a virtual machine (VM) running on a hypervisor software, e.g. VirtualBox\footnote{\url{https://www.virtualbox.org/}} is made by Kim.\cite{Kim_2019}
A major advantage compared to experiments with a non-reproducible ad-hoc environment is, that other researchers can simply copy the VirtualBox image containing the whole operating system, including the software stack and all dependencies.
However, when using the original VM image to build on top of simulations and expanding research, practitioners often have to install several other pieces of software which could cause altering dependencies of existing packages which in turn could lead to falsification of the initial results.
In this scenario, using Nix and NixOS could address this challenge, because here, dependencies of software stacks are always complete and self-contained.
So even if researchers want to expand research, new software can be installed alongside existing packages without interference.

\section{Background}

In our case study, we use Nix for building and packaging preCICE adapters and their respective solvers.
This chapter gives a short introduction into Nix and preCICE.

\subsection{Nix}

Nix is a general purpose build tool and purely functional package manager that provides some features that help to build software derivations reproducibly.
It does so, by recursively calculating a hash over all the inputs of a derivation and its dependencies to ensure the completeness of the whole derivation.
If any of the inputs changes, all dependent derivations will have to be rebuilt.

All contents, including source files and resulting build artifacts, are stored inside the Nix store.
Per default, the Nix store resides in the path \texttt{/nix/store} on the file system and all file- and directory names stored inside the Nix store are prefixed with a cryptographic hash of the inputs of the specific file/directory followed by a human-readable name and a version number.
This eases checking for integrity the Nix store and ensures that builds using the same inputs are performed only once.

These are two examples for possible directory names for the preCICE 2.5.0 package:

\begin{itemize}
  \item \texttt{/nix/store/0a5gw3l\ldots-precice-2.5.0}
  \item \texttt{/nix/store/9gpyg37\ldots-precice-2.5.0}
\end{itemize}

Hashing the inputs and storing them in the manner described above provides several features that other package managers cannot provide.
Nix enables users to have multiple versions of the same software installed at the same time without the softwares raising conflicts.
As each build output references their complete dependency graphs inside the Nix store, they don't interfere with other build outputs.
An example of a graphical representation of a dependency graph can be seen in Fig. \ref{fig:nix-graph}.
Nix modifies environment variables, such as the \texttt{PATH} variable, to make calling the binaries more convenient to users.
This fact allows multiple users to safely install software without the need of elevated permissions, as users cannot exchange the binaries of other users.

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{coreutils.png}
    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -ocoreutils.png}}
    \label{fig:nix-graph}
\end{figure}

Package derivations for the Nix package manager are written in a domain specific language which is also called Nix.
Derivations are the ``build plans'' on how Nix should build a specific software, which resulting build artifacts should be stored where, what the inputs are and so forth.
There are several different phases during a build, which contain shell code that is being run one phase after the other.
Those phases default to a given behavior, depending on the build system used.
All phases can be overridden freely to give users a maximum of flexibility when building and packaging software for Nix.
The most important phases users will override are the \texttt{patchPhase}, \texttt{buildPhase}, \texttt{installPhase} and \texttt{checkPhase}.
During the build, a special bash variable \texttt{\$out} is set that contains the path to the target directory inside the Nix store.

As the Nix language is purely functional, evaluating and building the derivation with the same inputs, yields the same outputs.
This is a critical factor where Nix's reproducibility comes from.
Also, Nix builds are sandboxed so they can reproducibly be run, meaning that there is no internet access possible during a build.

Another benefit of the reproducibility are remote builds, so a build cluster can build packages and serve its Nix store to other machines running Nix.
The upstream build cluster is called ``hydra'' and Nix defaults to using the upstream hydra as a substituter, i.e. fetching all packages from the hydras binary cache.
When evaluating a derivation, Nix checks if the hashed path is existent in the local Nix store.
If it's not, then it queries all known binary caches, which per default is only the upstream binary cache, and fetches the path and all its dependent paths until the full closure is copied into the local Nix store.

Software patches are quite easy to apply in Nix.
One can simply provide a list of \texttt{.patch} files inside the Nix file describing a derivation, Nix will then include the patch files in the calculation of the hash.
This new store path can probably not be found in the upstream binary cache, so Nix will fall back to building from source.
Nix also provides overriding all inputs freely, so every package can be flexibly adjusted to the users needs.

\subsection{preCICE}

preCICE (Precise Code Interaction Coupling Environment) is an open-source software library designed to facilitate the coupling of different computational simulation softwares.
It provides an interface that allows different simulation tools to exchange data and work together in a collaborative manner, enabling multi-physics as well as multi-scale simulations.

Many scientific simulations require complex combinations of multiple solvers, each specialized in a particular aspect of the problem.
For example, in fluid-structure interaction problems, where fluid flow interacts with deformable structures, separate solvers can be used to model the fluid dynamics and structural mechanics.
preCICE acts as a bridge between all solvers participating in the simulation, allowing them to exchange data and synchronize their computations.

preCICE offers a flexible and generic framework for coupling simulations.
It supports a wide range of solvers, including finite element, finite volume, and particle-based methods.
The library provides methods for interpolating data between meshes, managing communication between solvers, and handling time stepping and convergence.
It can handle different types of coupling scenarios, such as one-way or bidirectional couplings, loose or tight couplings, and static or dynamic meshes.
The core principle of preCICE's functionality is shown visually in Fig. \ref{fig:precice}.

\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{precice-figure.png}
    \caption{A visual representation on how preCICE works in principle, see: \url{https://precice.org}}
    \label{fig:precice}
\end{figure}

One main idea behind preCICE is to enable the combination of existing simulation tools without requiring significant modifications to their original source code.
Instead, preCICE provides an abstraction layer that sits between the individual solvers, allowing them to communicate and exchange data efficiently.
This approach promotes code reusability and reduces the effort required for coupling simulations.

preCICE is an open-source project, which means its source code is freely available and can be modified and distributed by the community.
This fosters collaboration and encourages the development of new features and improvements by researchers and engineers worldwide.

Overall, preCICE simplifies the process of coupling different simulation programs, enabling the simulation of complex multi-physics problems.
It promotes interoperability, code reusability, and collaboration, making it a valuable tool in various engineering and scientific domains.

\section{Case study}

To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution, which is currently preCICE \texttt{v2211.0}.
The preCICE distribution consists out of the preCICE core library, as well as some tools, bindings, adapters, tutorials, the website and documentation, and the preCICE VM\footnote{\url{https://precice.org/installation-distribution.html}}.

We are mainly interested in the tools, adapters including their solvers and the preCICE VM.
For each of these entities, we present possible challenges, how to circumvent those and also comment on how software can be designed to make it as easy as possible for other users to build and package the software in question.

\subsection{preCICE}

The preCICE library itself is already available in the nixpkgs repository, so we don't have to package the software.
Looking at the package definition and the preCICE source code, the library is quite easy to build and package as it uses CMake as a build system.

\subsection{Tools}

The preCICE distribution comes with a few tools that ease working with preCICE.
They are mainly meant for debugging and visualizing several aspects of preCICE and preCICE configuration files.

\subsubsection{ASTE}

This package is a collection of tools which help in development of new adapters.
It stands for Artificial Solver Testing Environment, so it allows for debugging adapter mapping setups and other similar tasks.

ASTE needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
However, this feature is needed for ASTE.
As already mentioned, Nix allows easy overriding of inputs, so in this case, we can look at the package definition of VTK.
The parameter \texttt{enablePython} can simply be set to \texttt{true}, then we also need to supply the version of python we want to compile against, so in our case \texttt{python = python3;}.

After enabling python support in VTK, the package can be built without further modifications.

Only for testing, we additionally need to replace the absolute paths to \texttt{/bin/bash} and \texttt{/usr/bin/time}, then all tests succeed without any problems.

\subsubsection{config-visualizer}

The precice-config-visualizer package can be used to generate images that represent a specified preCICE config file to check graphically what is configured in the file.

This tool is already packaged upstream, so we can just use it.
As it is a python package, it is quite easy to build with Nix, so the package definition is straight forward.

\subsection{Bindings}

Bindings are used to offer interfaces to other programming languages.

\subsubsection{python-bindings}

The \texttt{pyprecice} package is already available upstream, however it is broken and does not compile as it is not able to find the python module \texttt{pkgconfig}.
After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.

\subsubsection{Julia-bindings}

Julia support in NixOS is currently still in its early stages and cannot be declaratively be defined by Nix\footnote{\url{https://github.com/NixOS/nixpkgs/issues/20649}}.

\subsection{Adapters}

A preCICE distribution comes bundled with several adapters, which are currently not in the upstream nixpkgs repository.
Many of the adapters use solvers that aren't packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.\\

\subsubsection{CalculiX}

The preCICE adapter for CSM code CalculiX, requires the original CalculiX source code and provides a new Makefile in the adapter repository.
By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it's possible to override this location with a make variable.
It then builds the original code and the adapter code together, which results in a combined binary.

The Makefile also has hard-coded dependency location for \texttt{SPOOLES}, \texttt{ARPACK} and \texttt{YAML} so these need to be replaced with the equivalent \texttt{pkg-config} calls, because Nix doesn't have these dependencies at the hard-coded locations, and generally only finds dependencies with \texttt{pkg-config}.
Additionally, there is no \texttt{install} target provided by the Makefile so the resulting binary needs to be installed manually by copying it to \texttt{\$out}.\\

\subsubsection{code\_aster}

The code\_aster preCICE adapter is a python file that needs to be placed in the code-aster lib directory.
For this, the code\_aster solver needs to be packaged at the latest stable version 14.6.
It uses a custom build system based on a \texttt{setup.py} file, that invokes build and install phases for dependencies and in the end for the solver itself.

The stable package also distributes the dependencies as pinned tarballs.
On Nix almost none of these dependencies, complete the build stage with the \texttt{setup.py}, so we need to pull them out, package them separately as Nix packages and configured the build system with a \texttt{setup.cfg}.
In this file, we are able to disable the installation of the pinned dependencies and provide paths to the dependencies instead.

With this approach, we disable HDF5, which is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
HDF5 is packaged with multiple outputs.
One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, this is, because most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
Additionally, code\_aster requires medfile and scotch, which are both available upstream and can be used as a dependency.

Another dependency, which is also provided by code\_aster is metis, which partially builds, so we do not replace this dependency, but we need to fix the second part of the build, which includes the metis programs like gpmetis, ndmetis, mpmetis and more.
The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \texttt{/home/karypis/local/lib}.
After removing this line, it also installs, so to resolve this issue in the buildPhase, we unpack the metis tarball, remove this line with \texttt{sed} and then repack the directory.

The last 2 dependencies, are not available and need to be packaged, this includes mumps and tfel, also known as mfront.

Mumps builds using a Makefile and can be configured using Makefile.inc, by default it provides a couple of these configurations for systems like Ubuntu, so we use one of these provided configuration files to write our own custom config file.
It defines the location for dependencies like scotch, metis, parmetis, liblapack, blas, scalapack and blacs.
Almost all the dependencies, are available upstream, but we need to recompile scotch with additional build flags, \texttt{scotch ptscotch esmumps ptesmumps} using Nix overrideAttrs feature.
For packaging blacs, which also uses a Makefile, we configure the build using a custom \texttt{Bmake.inc} file, which is used to set compiler flags, the path to the mpi and bash installation.

The last missing dependency for code\_aster, tfel, uses CMake as a build system which makes packaging it trivial with Nix and we only have to set a couple of CMake flags.

After successfully packaging all dependencies, the build phase of code\_aster completes, but the installation part has another issue, that also affects new versions of Ubuntu and other distributions which use a python version greater than 3.9.
There is a forum post without any resolution from the developers or the community, so we need to manually patch the bug.

The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, which works if your version is 3.9.x but not with 3.10.x because then the code calculates \texttt{lib64/python3.1/site-packages} rather than the expected \texttt{lib64/python3.10/site-packages}.

After patching this bug code\_aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib} and we provide symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/} as code\_aster expects configuration files in these directories.

Last but not least, we can copy the code\_aster adapter at the required location which concludes this adapter installation.\\

\subsubsection{deal.II}

In order to package this adapter, we first need to package deal.II, which uses CMake as build system and works without any adjustments.
The same applies for the adapter, which also uses CMake and needs preCICE as a dependency.
When supplied with the inputs, the adapter builds without any problems.
As an example for the Nix language, you can see the abbreviated Nix code for the dealii.II adapter in Fig. \ref{lst:dealii-adapter-nix}.\\

\begin{figure*}
    \begin{multicols}{2}
    \normalsize
    \begin{minted}{nix}
{ lib, stdenv, fetchFromGitHub, cmake,
  precice, dealii, enable3d ? false }:

stdenv.mkDerivation rec {
  pname = "precice-dealii-adapter";
  version = "unstable-2022-09-23";

  src = fetchFromGitHub {
    owner = "precice";
    repo = "dealii-adapter";
    rev = "dbb25...8367c";
    sha256 = "sha256-pPQ2...2jlflgUE=";
  };

  nativeBuildInputs = [ cmake ];
  buildInputs = [ precice dealii ];

  cmakeFlags = lib.optionals enable3d [
    "-DDIM=3"
  ];

  installPhase = ''
    mkdir -p $out/bin
    cp elasticity $out/bin/elasticity
  '';
}
    \end{minted}
    \end{multicols}
    \caption{A sample of Nix code used to package the dealii-adapter, see: \url{https://github.com/precice/dealii-adapter/}}
    \label{lst:dealii-adapter-nix}
    \hrulefill
    \vspace*{4pt}
\end{figure*}

\subsubsection{DUNE}

This package uses a combination of CMake files and custom build scripts which makes the build process quite tedious.
This might be an issue because of the ongoing migration to CMake from autotools and might be resolved in the future.

It's also a minor issue that the dune-project is a collection of repositories rather than a monorepository, because users have to correctly clone all repositories so the build system finds all relevant information.
We clone all of the required dune repos into a directory and additionally clone the adapter into the same directory.

For the build and install, process we need to manually set \texttt{\$DUNE\_CONTROL\_PATH} and \texttt{\$DUNE\_PY\_DIR} environment variables, both variables also need to be set at runtime.
Additionally, we need to patch the python install process because the current CMake file that handles this install tries to access the internet with \texttt{pip install} which isn't allowed in Nix' sandboxed builds.
Therefore, we need to patch this \texttt{pip} command with the options \texttt{--no-build-isolation --no-cache-dir --no-index --no-deps} and we need to remove the \texttt{--upgrade} option.\\

\subsubsection{FEniCS}

We are explicitly referring to the legacy version of FEniCS here, not the newer and still experimental FEniCSx.
This adapter uses the open source computing platform FEniCS, which is already packaged within nixpkgs, to provide preCICE support.

Although it is available, not all features the adapter needs are enabled, including PETSc support.
To fix this, we need to package PETSc4py, which provides python bindings for PETSc and is not present in the upstream nixpkgs repository.
The build process is possible without any issues, because it uses the internal \texttt{buildPythonApplication} build tool, but we need to add the additional build optional \texttt{build\_src --force} because the cython code needs to be rebuilt.
We also aren't able to enable tests for this package because tests depend on OpenMPI and the network, which isn't available within the Nix build sandbox.

After this, we can specify PETSc4py as dependency for FEniCS and can enable support for this feature.

Building the adapter with Nix is now possible with \texttt{buildPythonPackage} and doesn't require any special workarounds.
A majority of the \texttt{pytest} ran successfully, but two tests failed inside the Nix sandbox, which means all tests needed to be disabled.\\

\subsubsection{OpenFOAM}

There are two flavors of OpenFOAM, we are referring to the OpenFOAM fork of OpenCFD Ltd\footnote{\url{https://www.openfoam.com/}}.

OpenFOAM uses its own custom build system called \texttt{wmake} which is sometimes called with a wrapper script called \texttt{Allwmake}.
The build system sets 36 environment variables that are set by wmake, one step at a time by checking several parameters, e.g. the cpu architecture or the location of the source code, and concatenating some other environment variables together.

Allwmake and wmake rely on these environment variables to be set, as otherwise the tools will exit with a warning.
During runtime, before running any OpenFOAM commands, a small wrapper script has to be started that sets all relevant environment variables, or a user has to manually source a file inside the installation directory of OpenFOAM.
OpenFOAM uses modules for extensibility and flexibility which need the source code of OpenFOAM and also require wmake as a build tool.

For Nix, these properties are rather unfortunate.
To compile OpenFOAM with Nix, we first need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
Secondly, we render a shell script we call \texttt{set-openfoam-vars} that exports all necessary environment variables to the current shell session.
At the time of writing, all parameters such as the processor architecture are hard-coded for simplicity reasons but could be parametrized based on the Nix inputs to allow for optimized builds.
During the buildPhase, the rendered script will be sourced to set all variables, e.g. so the \texttt{OPENFOAM\_SRC\_PATH} points to \texttt{/build/openfoam} which is the default location for Nix builds inside the sandbox.
After sourcing the script, the single command \texttt{./Allwmake -j -q} is sufficient to start the build.

The installPhase then simply copies the necessary files and directories to \texttt{\$out}, replaces the mock \texttt{OPENFOAM\_SRC\_PATH} by the value of \texttt{\$out} and creates a wrapper for the \texttt{openfoam} shell script.

As the preCICE OpenFOAM-adapter is an OpenFOAM module, the derivation of the adapter needs OpenFOAM and wmake as build inputs.
The correct environment variables have to be set again, but this is rather easy now, as we have the \texttt{set-openfoam-vars} script that we can simply source during the buildPhase.
The last thing we need to change before launching the build with Allwmake is to set the target directory for the adapter.
A single shared library that can be copied to \texttt{\$out/lib/} is the only output for this build.

Debug and timing modes for the adapter can simply be enabled by setting the \texttt{debugMode} and \texttt{enableTimings} inputs to the derivation to true.

OpenFOAM is available on Spack and EasyBuild which both patch the OpenFOAM source files as those tools also have to set environment variables to make a build compatible with wmake.\\

\subsubsection{su2-adapter}

This package is a preCICE adapter for the CFD code SU2 and extends the original SU2 source code by patching the files.
For this, the adapter provides a custom script that needs to be executed before the build phase.
Nix offers the patchPhase which allows us to run the script in this phase and patch the original source, after that the \texttt{stdenv} automatically recognizes that the repository uses autotools for building the software and runs the configurePhase.

We additionally needed to add the two configureFlags \texttt{--with-include} and \texttt{--with-lib} to the preCICE install directory because otherwise autoconf does not find the preCICE installation directory.

\subsection{preCICE VM}

The current preCICE VM is built on top of Vagrant\footnote{\url{https://www.vagrantup.com/}} using Ubuntu as a base image.
When it is run for the first time, it further installs software, compiles programs and the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the ``latest'' version, e.g. the \texttt{main} branch of the preCICE repository.

Nix comes with the built-in functionality of producing qemu\footnote{\url{https://www.qemu.org/}} VM images.
We used Nix to define a VM image with NixOS as a base, that can be built reproducibly.
The image contains all the adapters and solvers from the case study and some additional custom tools that can only be seen in the official, like the \texttt{preciceToPNG} command.
Also, we generate an iso and a Vagrant VirtualBox file of the VM as additional outputs.

Most of the functionality that can be found in the official VM is also available in our NixOS VM and preCICE tutorials, like the \textit{perpendicular-flap} can be run successfully.

\subsection{Discussion}

After detailing the challenges of building and packaging each solver and their respective adapter, we want to discuss some common points between these scientific software packages.
preCICE, as a multi-component research software, covers many different simulation codes and therefore provides a good overview over different kinds of packages, making this case study representative for multi-component research software.\\

Generally speaking, there are major gaps regarding the ability to successfully package the software between each different software solution that can be traced back to the build system.
Custom build systems require an disproportional amount of additional work to make them runnable on any system.
OpenFOAM and Allwmake are prime examples of these issues because now package managers have to understand a whole new build system for exactly one package in order to successfully package it and make it easily available to users.
It is also worth mentioning that writing your own custom build system, requires additional time to maintain this build system, which could have been spent working on the software itself.
Packages that use common, industry proven package managers, like CMake and autotools make it easier for developers of the software to build their software as well as package managers to package it.
These build tools, generally also provide interfaces for dependency management which can be used to tell users and package maintainers which dependencies are needed, and it is usually better than only documenting required dependencies, because documentation needs to be kept in sync with the underlying code.
They also generally already provide proven interfaces for configuring packages with additional features, and these also might be self documenting.

The authors of the xSDK\footnote{\url{https://xsdk.info/}} came to a very similar conclusion.
Their goal is to ``identify, adapt, and adopt best practices in software engineering''\cite{xsdk-website} and ``to provide the foundation of this extensible scientific software ecosystem''\cite{xsdk-website}.
To achieve this they provide a list of package policies~\cite{xSDK2023}, which for example specifies as mandatory policy that packages ``must support portable installation through Spack''\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''\cite{xSDK2023}, which includes CMake and autoconf as examples.
If a package supports all these policies, they can be added to the growing list of packages in the xSDK, which currently among other things include PETSc, deal.ii and preCICE.
These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes like nix store paths, and common build systems like CMake and autoconf.
\begin{figure}[htbp]
  \centerline{\includegraphics{figures/xsdk.png}}
  \caption{xSDK Logo}
  \label{fig}
\end{figure}

Another finding is, that is also addressed in xSDK through portable installations, is the requirement to install software to be installed into \texttt{\$HOME} or requires files to be located at a specific location on your file system.
Most of the time, it is possible to work around this, but there are already available solutions for dependency management.
Both, the CalculiX-adapter and SU2-adapter do require the original source code to be present as they add new features on top or patch the source code.
The current solution of the user cloning the code at the specific location works, but it might be better to look into submodules because that also allows the developers to pin the version more easily rather than documenting it, as that could more easily lead to user error, when they try to package and install the adapters.

Additionally, hard-coding libraries to \texttt{/usr/lib} and \texttt{/usr/include} might work on one distribution, but might make it hard or even impossible to install a package on a different distribution or even a later version of the same distribution.
Solutions like \texttt{pkg-config} that can be configured, so it finds libraries outside the common paths, should always be preferred and work out of the box with the common paths like \texttt{/usr/lib}.

\section{Comparision}

This chapter defines metrics and uses these metrics to compare the Nix package manager~\cite{Dolstra_2004} with widely used HPC package managers Spack~\cite{Gamblin_2015} and EasyBuild~\cite{Geimer_2014}.
The logos of the different package managers can be seen in Fig. \ref{fig:pm-logos}.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{pm-logos.png}
    \caption{Logos of (left to right) Nix\cite{nix-logo}, Spack\cite{spack-logo} and EasyBuild\cite{easybuild-logo}.}
    \label{fig:pm-logos}
\end{figure}

\subsection{Metrics definition}

In this section, we define all metrics used in the comparison of the selected package managers.
The metrics are grouped with related metrics.\\

\subsubsection{Overview}
The first few metrics are grouped as Overview, where we present some basic stats about each package manager.
First, we look at the number of available packages, then at the number of unique packages, meaning all packages minus packages that are available at different versions and last we compare how packages can be implemented.\\

\subsubsection{Installation}
In this section, we look at the installation of all 3 package managers, how they can be updated and if root permissions are needed to either install it or later install packages.\\

\subsubsection{Usage}
In Usage, we present basic usages of the package managers.
We show how you can search for packages, install software, install software with a specific version, install software at a specific point in time, specify the compiler that is used to install a piece of software and last we show how to build optimized software for your architecture.\\

\subsubsection{Reproducibility}
Here we look into reproducibility of build software.
We first look into how the reproducibility is achieved, then if the architecture and cpu features are considered and latestly we look at the dependencies can affect reproducibility.\\

\subsubsection{Performance}
We will take a short look at the performance of the package manager and the resulting packages\\

\subsubsection{Testing}
Finally, we look into tests and see what features the package managers provide for validating that a package works as expected.

\subsection{Results}

This section presents the results we collected.
First we will provide a description and afterwards we will discuss the gathered results.\\

\subsubsection{Overview}

For the first set of metrics, we observed that in our comparison Nix provides the most packages with 83000 packages, 67000 of them unique, followed by Spack with around 7000 packages, 6000 of them unique.\footnote{\url{https://repology.org/repositories/statistics/total}}
Regarding EasyBuild, there are around 17000 and 3000 unique packages.\footnote{\url{https://docs.easybuild.io/release-notes/}}

While Nix has the most amount of packages, it is a general purpose package manager not specifically made for scientific software.
This means, that most of the packages are not focused on scientific software, but it still comes with important packages like HDF5 and mostly offers mpi support for those packages.

Spack and EasyBuild lay their focus more on the HPC/Simulation environment, so the main purpose of these package managers is to provide scientific software, but that also means that sometimes common packages, like openssl3.0 simply aren't available.

Both, Spack and EasyBuild, use python scripts to implement their packages, Nix on the other hand, uses with its own DSL.
Using a common language to define packages makes it easier for new people to get started packaging their own software because they might already be familiar with the language and generally, there is more documentation available for general purpose language, especially python.
The Nix DSL is only useful inside Nix and people looking to package their software for Nix, need to learn a new language that is only applicable within Nix.\\

\subsubsection{Installation}

Nix, the package manager, can be installed with a custom installation script.\footnote{\url{https://nixos.org/download.html\#download-nix}}
There are multiple ways to install it, with the Nix daemon, making it a multi-user installation, this installation requires root privileges for installation and installs it in \texttt{/nix}, after that normal users can use the Nix binary to communicate with a running daemon to download and install packages, this does not require root permissions.
The single user installation allows to setup the Nix package manager inside your home directory, this method does not require root for installation and usage.
The third way is to install Nix as part of NixOS and requires a new installation of the operating system, requires root for installation but not for usage.

Spack is being installed with git and can be cloned into any directory, which means it does not require root for installation and usage.\footnote{\url{https://spack.readthedocs.io/en/latest/getting_started.html\#installation}}

EasyBuild also has multiple ways to install it.\footnote{\url{https://docs.easybuild.io/installation/}}
It can be installed with \texttt{pip install} (or virtualenv/pipenv) and after that can be loaded with \texttt{module load EasyBuild}.
It can also be installed into a temporary directory using \texttt{pip} and afterwards properly installed with using the temporary installation to bootstrap an install into the latest release.
EasyBuild installation requires modifying \texttt{PATH}, \texttt{PYTHONPATH} and potentially \texttt{MODULEPATH} depending on the location, it is being installed in.
It also requires Tcl/C environment-modules or Tcl-only variant of environment modules or Lmod being installed and configured.
None the less EasyBuild can be installed without root privileges and generally can also be used without root, but EasyBuild has a package option called \texttt{osdependencies} with requires dependencies being installed on the operating system.
For example CMake (till 3.18) requires \texttt{openssl-devel}, \texttt{libssl-dev} and \texttt{libopenssl-devel} making it impossible to install for a normal user, if one of the above packages is not installed.
This also means that users can't install any package that requires the CMake build system without updating the package definitions.

Last but not least, updating the tools and their packages, is also widely different.
Nix has something that is called channels that need to be advanced upstream and after that users can update their local channels with \texttt{nix-channel --update}, or use the newer Nix flake technology which pins the inputs so users can choose freely when to update to which commit of the upstream nixpkgs repository.

Spack is being updated by fetching a new version of the local repository, which then can lead to merge conflicts if users change the provided package definitions.

EasyBuild can be updated using the \texttt{--install-latest-eb-release} the tool provides.\\

\subsubsection{Usage}

The basic usage of each tool is pretty similar.
All 3 package managers allow a user to query for a package than then can be installed.

Nix allows for downloading and installing software within the current shell session\footnote{\url{https://nix.dev/tutorials/first-steps/}}, that means that an only accessible within the shell and not actually installed on the system in the classical sense of making it available to multiple users of the system.
A user or an administrator can at any point \texttt{collect-garbage} and all packages not specified as installed system packages are being removed that includes packages installed in the current shell, as long as the shell session isn't still open.
If a user wants to install a specific software at a given version there are two main options made available by Nix.
Either there is an old package version definition provided by the packages, like \texttt{boost159} or a user can opt into overriding the package by overriding the \texttt{version} and the \texttt{src} \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, that will build the package given the specific source, which can be a remote tarball, revision in a repository or even a local directory.
With these options, users can override any aspect of the package definition, including the dependencies and even providing custom patches.
Each change to the package will trigger a new build because the input for the build was changed.

Spack and EasyBuild on the other hand, keep the old software versions in the specifications or keep the whole specification, so you can reinstall an old version by specifying the available version, for example, \texttt{spack install gcc@6.4.0}.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}}
If the required version is not available, you have to edit the current package definition and add the required version.
There is no elegant way of overriding the package in a command call and Spack even provides a command \texttt{spack edit} which does it in tree.
This can be an issue if you later want to update the git revision, as this might lead to merge conflicts as soon as you checkout a new revision, as described earlier.

To achieve reproducibility, it is necessary that we can install software from a previous revision of the packages.
Nix allows us to simply specify the channel revision we want to use to build our software, it then evaluates the requested packages and builds it.
This can easily be done without affecting the current system by specifying it in the shell command \texttt{nix-shell -p pan -I nixpkgs=pathToRevision/urlToTarball}, it now only uses this channel for building the requested package.
Spack, can achieve something similar by checking out your Spack version at a previous revision and then installing the package requested package, but this approach has a downside that it affects the whole Spack installation, and now also work with an older Spack tool.
On the other hand, EasyBuild, could be installed using pip at an earlier version, but it also offers something different called toolchains, which we will look at shortly.

Next, we look at how you can build a package with a specific compiler version.
The flexibility of Nix overrides allows different levels of rebuilds of packages with a specific compiler, you can either just rebuild the requested package with specific compiler, this for example rebuilds curl with clang8 \texttt{nix-shell -p '(curl.override \{ stdenv = clang8Stdenv;\})'} or you can rebuild the package and all required dependencies with a specific compiler by overriding the \texttt{stdenv} on nixpkgs level, \texttt{nix-build -E 'with (import <nixpkgs> \{ config.replaceStdenv = \{ pkgs \}: pkgs.clang8Stdenv; \}); curl'}.
This is possible because every c/c++/\ldots package, is build using the \texttt{stdenv} which defines the compiler and default compile flags.
While Nix provides a default stdenv, which is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, like the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
By default, there are some useful helpers provided, like \texttt{makeStaticLibraries} which allows compiling any library as a static library, or \texttt{enableDebugging} which sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.

Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}}
On this note, Spack tooling and interface to override compilers, flags and more is way more user friendly and readable than using and invoking the Nix DSL within a one line command call.

EasyBuild has a whole different system, where packages have defined toolchains.\footnote{\url{https://docs.easybuild.io/common-toolchains/}}
A toolchain is a set of packages defined at a specific version and users can create their own toolchains.
For example, EasyBuild provides a foss toolchain which contains packages like, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version like foss-2022b.
Packages are then provided with toolchains in mind and are never deleted, so installing a package with a different compiler can be achieved by selecting a different eb script with a different toolchain version.
On the other hand this results in a less elegant user flow, because if a user wants to install a package with a toolchain that isn't provided they have to fork a package definition and update the toolchain.\\

\subsubsection{Reproducibility}

\begin{figure}
    \centering
    \includegraphics[width=0.4\textwidth]{nix-tree.png}
    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --tree \$(nix build nixpkgs\#coreutils --print-out-paths)}}
    \label{fig:nix-tree}
\end{figure}

As previously mentioned, Nix computes an output hash over all its inputs that is unique for this input, so if anything in the inputs changes, like the version of a dependency, the output hash changes resulting in a different software build.
This guarantees that given the same set of inputs, we are able to compile the same software, including it's dependency tree as you can see in Fig. \ref{fig:nix-tree}.
One thing to mention is, that Nix currently only builds generic architecture build, so the inputs of a derivation don't include any information about the architecture and/or compiler features.
Another thing to mention is, that all software in the \texttt{/nix/store} can only link back to the \texttt{/nix/store} itself.
Nix validates this for every package and breaks links than allowing links back to a file in the home directory or the directory that was used for building the software.
This is required, to guarantee the best possible reproducibly.

Spack does something similar to Nix where it adds hashes to each installation path, but also stores additional files inside the package installation which include the complete spec of the package than can later be used to reproducibly build the same package.~\cite{Gamblin_2015}.
It also does something that Nix, currently doesn't support, which is it storing the architecture and cpu features that are used in this build, that way it can provide binary packages that are optimized for specific CPUs.
None the less it has one problem that Nix doesn't have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies.
\begin{itemize}
  \item libm.so
  \item libc.so
  \item libdl.so
  \item libnl-3.so
  \item libnl-route-3.so
  \item librt.so
  \item libutil.so
  \item libpthread.so
\end{itemize}
Most of these dependencies are stable, and usually don't change on a system update, but it also means that it simply can't guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.

EasyBuild does achieve reproducibility with its toolchains and the approach to keep old package definitions, meaning that they also keep old dependency definitions and each dependency is pinned at a specific version~\cite{Geimer_2014}.
Regarding, CPU architecture and features, EasyBuild does not do generic builds by default but rather prioritizes the idea of optimizing for a specific system, meaning defaulting to \texttt{-march=native} for all packages and it has to manually be disabled~\footnote{\url{https://docs.easybuild.io/controlling-compiler-optimization-flags/\#controlling\_compiler\_optimization\_flags\_optarch\_default}}.
This leads to inherently impure and irreproducible builds as building the same easyconfig on two different host/architectures results in different binaries.
On the other hand, the already mentioned option for \texttt{osdependencies} goes against the reproducibility idea, because any os dependency is not managed by the package manager and might change at any point.
This leads to the problem of binaries linking back to the operating system \texttt{/usr/lib}, that Spack also has.
For example a build of curl, uses libssl, libcrypto, libz, libm and libc, all of which are provided by the operating system, and for EasyBuild it includes libraries that change more frequently like libssl, libcrypto and libz.\\

\subsubsection{Performance}

With Nix' possibility to override stdenv it is also possible to set custom compiler flags and Nix provides an option \texttt{impureUseNativeOptimizations} to build packages optimized for your hardware, using the \texttt{-march=native} compiler flag, it can then be applied to the package, that is being installed, or every dependency in the dependency graph.~\footnote{\url{https://nixos.wiki/wiki/Build_flags}}
Nix can guarantee these compiler flags for every dependency because Nix wraps its compiler binaries in scripts, which then means it has more control over them and packages can not remove any flags by clearing environment variables.
The advantage of overriding the stdenv is that every package in a dependency tree can be optimized which might lead to better overall performance of packages, but it also has the downside that every package needs to be rebuild which might take several days, for example preCICE required a rebuild of 1500 packages, which took 2 days on an high-performance computing (HPC) cluster (2x AMD EPYC 7702).
So it might be sufficient to optimize only the package, which is easily possible with \texttt{(package.override)} like shown before or additionally optimize its direct dependencies, which is possible but more time-consuming to do, because one has to manually override each dependency.

Spack also allows to override cflags, using \texttt{spack install \ldots cflags='-march=native'}, so the package is then being optimized.
By setting cflags with \texttt{==}, the cflags are also applied to the dependencies, causing them to be rebuilt.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html\#compiler-flags}}
Additionally, there is a way to define flags for a specific registered compiler, but these are only respected by CMake and autotools packages.
Also, a package might also have unavoidable system dependencies and are not optimized, which might impact performance negatively.

Like previously mentioned, EasyBuild applies \texttt{-march=native} by default, but you can further customize your compiler flags using \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS}, \texttt{\$FFLAGS}, \texttt{\$F90FLAGS}.
It also has the same problem as Spack that it might have unavoidable unoptimized system package which might impact performance negatively.\\

\subsubsection{Testing}

We also shortly look at the capabilities of the package managers to run test and evaluate the built packages.
Nix has something that is called a checkPhase which can be run for each package after being built.
There are sane defaults for all available build systems, like \texttt{make test} or \texttt{ctest} for makefiles and CMake, \texttt{cargo check} and \texttt{cargo test} for rust packages, \texttt{go test} for go packages and \texttt{pytest} for python packages and more.
By default, this checkPhase is not enabled for every build system, like the \texttt{stdenv} build system, which is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
For these build systems, package maintainers have to manually disable the check phase or they can override the check phase if the tests aren't working out of the box and make the needed adjustments to resolve the issues.

Spack has something similar, but its not enabled by default, and users, not package maintainers, has to manually specify that they want to run test on installing the package with \texttt{--test=root}.

EasyBuild on the other hand does not offer this feature but has another feature called sanity checks, where EasyBuild ensures that paths and or binaries, specified by a package maintainer inside the package definition, are successfully installed.

Additionally, Nix offers a way for testing multiple packages in combination and or with a specific network setup by providing a way to write your own tests.
These tests are then running inside isolated qemu virtual machines and users are able to start as many VMs as they need and define the set of commands that need to run on each of the so-called nodes.
The tests are pythons scripts, and it is possible to wait until systemd targets are running, until ports are available or wait until commands are completed and compare the output and status codes.

\section{Nix on HPC}

High-performance computing refers to the utilization of powerful computational resources to tackle computationally intensive scientific problems.
It involves the parallel execution of tasks across multiple computing units, also called nodes in this context, enabling researchers to perform complex simulations, data analysis, and modeling at large scales.
HPC systems, such as clusters, supercomputers, and cloud-based infrastructures, usually offer immense processing power, storage capacity, and specialized hardware, supporting scientific research across various domains\footnote{\url{https://www.ibm.com/topics/hpc}}.

In many HPC environments, the \texttt{module} system is being used to load different pieces of software from a local repository of available softwares, such as compilers and different message passing interfaces (MPI).
With the module command, different users can use different software or different versions of the same software at the same time without interfering with each other.
One major drawback here is, that without further configuration, users can only load software modules that were previously installed into the cluster by the clusters administrators, so any other software that is needed for a simulation, including its dependencies, have to be compiled by the users that want to conduct an experiment.
Sometimes, this can be quite hard, as some software expects libraries or other software artifacts at some hard-coded location as \texttt{/usr/lib/\ldots}\footnote{\url{https://hpc-wiki.info/hpc/Modules}}.

As mentioned, Nix uses the Nix store, which lies in \texttt{/nix/store} by default and all software artifacts are linked against libraries inside the Nix store.
So if there is no such path as \texttt{/nix/store} on the HPC clusters filesystem, all software would need to be recompiled as all links and paths inside the binaries would point to a non existent location.
We'd need administrator rights to be allowed to setup the Nix store globally on the cluster and to install Nix globally for all users.
However, Nix supports a Linux kernel feature called user namespaces, which is a form of kernel based isolation of resources and other attributes like user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.

Using namespaces, it is possible for the Nix package manager to mount an arbitrary user-writable location to \texttt{/nix/store} inside a user namespace, so all programs running in that namespace can access the Nix store as if it would actually be installed under the default store path.
Another challenge is to get Nix to work without the needed libraries that can usually be found inside the Nix store, leading to a bootstrapping problem.
We can simply use a statically compiled version the Nix package manager, however, to circumvent this problem.
In our repository, we provide a setup bash script that simplifies download, configuration and installation of the static Nix binary which can then be used to easily build and access the solvers adapters and the preCICE VM.

\section{Conclusion and Outlook}

In conclusion, the case study has shown that packaging of scientific software isn't as trivial as expected.
The majority of the packages required some solution for a fundamental issue within the scientific software or its build system that isn't solvable for none experienced Nix users, which makes it hard for people to use it even for their own software.
Nix itself also is partly at fault, because of its unique nature, that every piece of software lies within \texttt{/nix/store} and that each store path is its own-isolated tree based on the Filesystem Hierarchy Standard (FHS).
But it is worth mentioning that not only Nix would benefit from more streamlined build systems and packaging, because both Spack and EasyBuild do something similar for their packages, where each package also lies in its own FHS.
The work of the xSDK people and their policy might be a step in the right direction but looking at the current list of software that is available within the latest xSDK release, 26 libraries, indicates that it is an uphill battle that might take several more years, especially if we consider that the first release happened in 2016.
We have also seen, that the current HPC solutions, Spack and EasyBuild make it somewhat possible to reproducibly build software but not to the degree in which Nix allows it.
That is because they focus on flexibility of installing packages.
Spacks way of allowing users to change the compiler of a package, easily enable additional features on installation, override cflags and more advanced features, for example regarding mpi, shows that the Nix tooling needs more work or a layer on top that strips away the uniqueness of the DSL for something more user- and beginner-friendly.

Lastly we looked at Nix on HPC and have seen that it's possible to get it running, and we are able to run simulations on one node, so we currently assume that with the correct networking setup it should be possible to run a simulation across multiple nodes.
It is currently not ideally solved on getting Nix running on HPC as a single user installation, so this can be improved in the future.

The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/}.
Also the \texttt{setup.sh} and \texttt{clean.sh} scripts used to install Nix on HPC can be found there.\\

While this case study and this comparison has provided valuable insights into the topic of reproducible software, there are still numerous unanswered questions that merit further investigation.
This includes a full evaluation if there are performance improvements using a full impure system rebuild, with \texttt{-march=native} compared to only rebuilding the package at hand like we suspect.
It would also be interesting to look at trying to reproduce a nixpkgs commit from 10 years ago, and if there might be issues and incompatibilities with newer Nix versions, which currently get released in a 6 weeks time windows.

Last but not least, it would be interesting to investigate Nix and NixOS on HPC even more.
It would be interesting to see how an HPC cluster completely build on NixOS could improve maintaining the HPC cluster, with slurm already being available as a nixpkgs package and module providing a systemd service.

Finally, it would be interesting to investigate a nix store that is shared between users and how it might optimize the amount of storage needed for a cluster as each user is not required to rebuild the same package in their respective home directories.

This summary shows, how diverse and complex Nix is, and how much more opportunities there are to be researched in the future.

\bibliographystyle{IEEEtran}
\bibliography{paper}

\end{document}

### 2023-10-27 paper/paper.tex
paper: Add "case study" to title
paper: Explain what is "not scalable" to traditional approaches

@@
 
-\title{Improving reproducibility of scientific software using Nix/NixOS}
+\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
 
@@
 Problems arise from dependencies, library versions, and system configurations, leading to inconsistencies across different computing environments.
-Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are time-consuming, prone to errors, and not scalable.
+Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are time-consuming, and prone to errors.
+It is not scalable and inefficient to make every researcher to tediously recreate the exact same environment with traditional approaches.
 This calls for a more efficient and automated solution.

### 2023-11-04 paper/paper.tex
paper: Highlight goals of the case study earlier

@@
 In the past years, the Nix package manager\cite{Dolstra_2004} and NixOS\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising solutions to address these challenges.
-This paper explores the use of Nix/NixOS to improve the reproducibility of scientific software, focusing on a case study of preCICE\cite{preCICEv2} and its adapters and solvers.
+This paper explores the use of Nix/NixOS to improve the reproducibility of scientific software.
+
+Conducting a large scale research on building scientific codes reproducibly is quite infeasible due to how many different software there is.
+Therefore, we conduct a case study, focusing on preCICE\cite{preCICEv2}, its adapters and solvers.
+One of the goals of the case study is to get a first hand experience and estimate on the complexity and possible challenges when packaging scientific software.
+Also, consistent portability of building the software on three different systems is verified.
 
@@
 
+With Nix' dependency management, we can easily copy the package definitions we just wrote onto different systems, e.g., entirely different Linux distributions, and execute the build of the closures there.
+Building the same closures yields the same outputs, this will also be described in section \ref{sec:nix-on-hpc}.
+
 \section{Comparision}
@@
 
-\section{Nix on HPC}
+\section{Nix on HPC}\label{sec:nix-on-hpc}
 
@@
 In our repository, we provide a setup bash script that simplifies download, configuration and installation of the static Nix binary which can then be used to easily build and access the solvers adapters and the preCICE VM.
+
+After installing Nix on HPC, we were able to produce the same binary output for the adapters and solvers as we built on our local machines as well as within the preCICE VM\footnote{The sha256 hashes of the resulting outputs were compared and matched.}
 

### 2023-11-04 paper/paper.tex
paper: Explain why preCICE is a good fit for the case study

@@
 Also, consistent portability of building the software on three different systems is verified.
+preCICE is a well fit for such a case study, as its official adapters and solvers vary in their programming languages, project sizes and other factors, providing a quite manifold set of scientific simulation software while still being able to verify if the software is still correct by running preCICE simulations across several different solver binaries.
 

### 2024-02-14 paper/paper.tex
feat: overview of precice distribution

@@
 The preCICE distribution consists out of the preCICE core library, as well as some tools, bindings, adapters, tutorials, the website and documentation, and the preCICE VM\footnote{\url{https://precice.org/installation-distribution.html}}.
+Table \ref{table:label-distribution} shows an overview over all packages provided by the preCICE distribution and their respected versions.
+
+\begin{figure*}[!t]
+  \normalsize
+  \caption{Overview of preCICE distribution}
+  \label{table:label-distribution}
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    \bfseries Package & \bfseries Type & \bfseries Build System & \bfseries Version \\
+    \hline
+    precice & library & cmake & 2.5.0 \\
+    \hline
+    fortran-module & bindings & make & 9e3f405 \\
+    \hline
+    julia & bindings & julia & v2.5.0 \\
+    \hline
+    matlab & bindings & matlab build script & v2.5.0.0 \\
+    \hline
+    pyprecice & bindings & setup.py & 2.5.0.1 \\
+    \hline
+    config-visualizer & tooling & setup.py & 60f2165 \\
+    \hline
+    precice-aste & adapter & cmake & 3.0.0 \\
+    \hline
+    dealii & solver & cmake & 9.4.1 \\
+    \hline
+    precice-dealii-adapter & adapter & cmake & dbb25bea \\
+    \hline
+    precice-calculix-adapter & adapter & makefile & 2.20 \\
+    \hline
+    fenics & solver & cmake + setup.py & 2019.1.0 \\
+    \hline
+    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\
+    \hline
+    precice-dune & adapter & scripts + cmake & 2.8.0 \\
+    \hline
+    Dune-fem & library & setup.py & 2.8.0.0 \\
+    \hline
+    code-aster & solver & custom build system in python & 14.6.0 \\
+    \hline
+    openfoam & solver & wmake/Allwmake & 2206 \\
+    \hline
+    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\
+    \hline
+    Precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\
+    \hline
+    Nutils & library & setup.py & 7.0.0 \\
+    \hline
+  \end{tabular}
+\end{figure*}
 

### 2024-02-14 paper/paper.tex
feat: apply changes made to abstract for deRSE24

@@
   Using this approach, we demonstrate how to create a reproducible and self-contained environment for preCICE and highlight the benefits of using Nix and NixOS for managing software and system configurations, resulting in improved reproducibility.
-  In addition, we compare the performance and usability of Nix with two widely used high-performance computing (HPC) solutions, Spack and EasyBuild.
+  In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
   This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.

### 2024-02-14 paper/paper.tex
feat: Add paragraph about docker

@@
 Problems arise from dependencies, library versions, and system configurations, leading to inconsistencies across different computing environments.
-Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are time-consuming, and prone to errors.
-It is not scalable and inefficient to make every researcher to tediously recreate the exact same environment with traditional approaches.
+Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are prone to errors and time-consuming at best.
+It is not scalable and highly inefficient to make every researcher to tediously recreate the exact same environment with traditional approaches.
 This calls for a more efficient and automated solution.
+
+Attempts were made to solve this situation by using Docker\footnote{\url{https://www.docker.com/}}, a software which describes software environments with the help of text files.
+Those text files are made up of imperative commands which are run inside of containers, one layer at a time.
+The result are several different layers which are all combines into a single output image that can be instantiated to a running container.
+Docker images can be copied to different hosts and should then provide the same environment on different machines.
+An issue that arises here, is that usually those images are based on one of the official Docker images\footnote{\url{https://docs.docker.com/trusted-content/official-images/}} all of which bring traditional package managers like \texttt{apt}.
+When a user specifies to install the \texttt{python3} package with a traditional package manager today, the package manager could yield python3.8, whereas running the same command two months from now could yield python in the version 3.9.
+This could potentially lead to different results when a user wants to only add a single dependency but has to rebuild the whole image, thus rendering it useless in terms of reproducibility.
 

### 2024-04-19 paper/paper.tex
feat: port paper to eceasst template

@@
-\documentclass[conference,final,a4paper]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
+\documentclass{eceasst}
+
+% \usepackage{cite}
+% \usepackage{algorithmic}
+% \usepackage{textcomp}
 \usepackage{minted}
-\usepackage{multicol}
-\usepackage{microtype}
-\usepackage[hyphens]{url}
+% \usepackage{multicol}
+% \usepackage{microtype}
+% \usepackage[hyphens]{url}
+\usepackage{subfig}
+\input{frontmatter}
 
@@
 
-\begin{document}
+\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
+\short{TODO}
+\author{
+Max Hausch\autref{1},
+Simon Hauser\autref{2}}
 
-\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
+\institute{
+\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
+\autlabel{2} \email{st148883@stud.uni-stuttgart.de}}
 
-\author{\IEEEauthorblockN{Max Hausch}
-\IEEEauthorblockA{Stuttgart, Germany \\
-st175425@stud.uni-stuttgart.de}
-\and
-\IEEEauthorblockN{Simon Hauser}
-\IEEEauthorblockA{Stuttgart, Germany \\
-st148883@stud.uni-stuttgart.de}
-}
-
-\maketitle
-
-\begin{abstract}
+\abstract{
 Ensuring the reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
@@
 In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
-  This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.
-\end{abstract}
+This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.}
 
-\begin{IEEEkeywords}
-  Reproducibility, Nix, NixOS, Spack, EasyBuild, preCICE, HPC.
-\end{IEEEkeywords}
+\keywords{Reproducibility, Nix, NixOS, Spack, EasyBuild, preCICE, HPC}
+
+\begin{document}
+\maketitle
 
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{coreutils.png}
+    \includegraphics[width=0.4\textwidth]{figures/coreutils.png}
     \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -ocoreutils.png}}
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{precice-figure.png}
+    \includegraphics[width=0.5\textwidth]{figures/precice.png}
     \caption{A visual representation on how preCICE works in principle, see: \url{https://precice.org}}
@@
 \begin{figure*}
-    \begin{multicols}{2}
     \normalsize
@@
     \end{minted}
-    \end{multicols}
     \caption{A sample of Nix code used to package the dealii-adapter, see: \url{https://github.com/precice/dealii-adapter/}}
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{pm-logos.png}
+    \includegraphics[width=0.45\textwidth]{figures/pm-logos.png}
     \caption{Logos of (left to right) Nix\cite{nix-logo}, Spack\cite{spack-logo} and EasyBuild\cite{easybuild-logo}.}
@@
     \centering
-    \includegraphics[width=0.4\textwidth]{nix-tree.png}
+    \includegraphics[width=0.4\textwidth]{figures/nix-tree.png}
     \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --tree \$(nix build nixpkgs\#coreutils --print-out-paths)}}
@@
 
-\bibliographystyle{IEEEtran}
+\nocite{*}
+\bibliographystyle{eceasst}
 \bibliography{paper}

### 2024-04-19 paper/paper.tex
fix: drop logos
chore: cleanup some citations

@@
 
-% \usepackage{cite}
-% \usepackage{algorithmic}
-% \usepackage{textcomp}
 \usepackage{minted}
-% \usepackage{multicol}
-% \usepackage{microtype}
-% \usepackage[hyphens]{url}
 \usepackage{subfig}
@@
 
-In the past years, the Nix package manager\cite{Dolstra_2004} and NixOS\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising solutions to address these challenges.
+In the past years, the Nix package manager~\cite{Dolstra_2004} and NixOS~\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising solutions to address these challenges.
 This paper explores the use of Nix/NixOS to improve the reproducibility of scientific software.
@@
 Conducting a large scale research on building scientific codes reproducibly is quite infeasible due to how many different software there is.
-Therefore, we conduct a case study, focusing on preCICE\cite{preCICEv2}, its adapters and solvers.
+Therefore, we conduct a case study, focusing on preCICE~\cite{preCICEv2}, its adapters and solvers.
 One of the goals of the case study is to get a first hand experience and estimate on the complexity and possible challenges when packaging scientific software.
@@
 By using Nix, researchers and practitioners in HPC could easily reproduce computational experiments, ensuring that the same software stack, libraries, and configurations are utilized consistently.
-Reproducibility of scientific results can be achieved using the Nix package manager as described by Devresse et al.\cite{Devresse_2015}
+Reproducibility of scientific results can be achieved using the Nix package manager as described by Devresse et al~\cite{Devresse_2015}.
 This not only streamlines the deployment process but also facilitates collaboration and sharing of software environments, making it easier for researchers to validate and build upon each other's work.
@@
 There are several other papers which discuss and highlight the importance of reproducibly of scientific experiments.
-In their paper, Dalle lists technical factors that lead to issues regarding reproducibly of in silico simulation experiments.\cite{Dalle_2012}
+In their paper, Dalle lists technical factors that lead to issues regarding reproducibly of in silico simulation experiments~\cite{Dalle_2012}.
 \begin{enumerate}
@@
 
-One suggestion to enhance reproducibility by using a virtual machine (VM) running on a hypervisor software, e.g. VirtualBox\footnote{\url{https://www.virtualbox.org/}} is made by Kim.\cite{Kim_2019}
+One suggestion to enhance reproducibility by using a virtual machine (VM) running on a hypervisor software, e.g. VirtualBox\footnote{\url{https://www.virtualbox.org/}} is made by Kim~\cite{Kim_2019}.
 A major advantage compared to experiments with a non-reproducible ad-hoc environment is, that other researchers can simply copy the VirtualBox image containing the whole operating system, including the software stack and all dependencies.
@@
 The authors of the xSDK\footnote{\url{https://xsdk.info/}} came to a very similar conclusion.
-Their goal is to ``identify, adapt, and adopt best practices in software engineering''\cite{xsdk-website} and ``to provide the foundation of this extensible scientific software ecosystem''\cite{xsdk-website}.
-To achieve this they provide a list of package policies~\cite{xSDK2023}, which for example specifies as mandatory policy that packages ``must support portable installation through Spack''\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''\cite{xSDK2023}, which includes CMake and autoconf as examples.
+Their goal is to ``identify, adapt, and adopt best practices in software engineering''~\cite{xsdk-website} and ``to provide the foundation of this extensible scientific software ecosystem''~\cite{xsdk-website}.
+To achieve this they provide a list of package policies~\cite{xSDK2023}, which for example specifies as mandatory policy that packages ``must support portable installation through Spack''~\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''~\cite{xSDK2023}, which includes CMake and autoconf as examples.
 If a package supports all these policies, they can be added to the growing list of packages in the xSDK, which currently among other things include PETSc, deal.ii and preCICE.
 These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes like nix store paths, and common build systems like CMake and autoconf.
-\begin{figure}[htbp]
-  \centerline{\includegraphics{figures/xsdk.png}}
-  \caption{xSDK Logo}
-  \label{fig}
-\end{figure}
 
@@
 This chapter defines metrics and uses these metrics to compare the Nix package manager~\cite{Dolstra_2004} with widely used HPC package managers Spack~\cite{Gamblin_2015} and EasyBuild~\cite{Geimer_2014}.
-The logos of the different package managers can be seen in Fig. \ref{fig:pm-logos}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.45\textwidth]{figures/pm-logos.png}
-    \caption{Logos of (left to right) Nix\cite{nix-logo}, Spack\cite{spack-logo} and EasyBuild\cite{easybuild-logo}.}
-    \label{fig:pm-logos}
-\end{figure}
 

### 2024-04-20 paper/paper.tex
feat(paper): Add citations where possible

@@
 
-We also compare Nix, EasyBuild, and Spack as package managers in the realm of scientific software and if Nix is feasible for being used in a scientific context at all.
+We also compare Nix, EasyBuild~\cite{easybuil6495863}, and Spack~\cite{spack7832814} as package managers in the realm of scientific software and if Nix is feasible for being used in a scientific context at all.
 
@@
 
-To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution, which is currently preCICE \texttt{v2211.0}.
+To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution~\cite{preciceDistribution}, which is currently preCICE \texttt{v2211.0}.
 The preCICE distribution consists out of the preCICE core library, as well as some tools, bindings, adapters, tutorials, the website and documentation, and the preCICE VM\footnote{\url{https://precice.org/installation-distribution.html}}.
@@
 
-ASTE needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
+ASTE\footnote{\url{https://github.com/precice/aste}} needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
 However, this feature is needed for ASTE.
@@
 
-The precice-config-visualizer package can be used to generate images that represent a specified preCICE config file to check graphically what is configured in the file.
+The precice-config-visualizer\footnote{\url{https://github.com/precice/config-visualizer}} package can be used to generate images that represent a specified preCICE config file to check graphically what is configured in the file.
 
@@
 
-Julia support in NixOS is currently still in its early stages and cannot be declaratively be defined by Nix\footnote{\url{https://github.com/NixOS/nixpkgs/issues/20649}}.
+Julia~\cite{bezanson2017julia} support in NixOS is currently still in its early stages and cannot be declaratively be defined by Nix\footnote{\url{https://github.com/NixOS/nixpkgs/issues/20649}}.
 
@@
 A preCICE distribution comes bundled with several adapters, which are currently not in the upstream nixpkgs repository.
-Many of the adapters use solvers that aren't packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.\\
+Many of the adapters use solvers that aren't packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.
 
@@
 
-The preCICE adapter for CSM code CalculiX, requires the original CalculiX source code and provides a new Makefile in the adapter repository.
+The preCICE adapter for CSM code CalculiX~\cite{Uekermann2017_Adapters}, requires the original CalculiX source code and provides a new Makefile in the adapter repository.
 By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it's possible to override this location with a make variable.
@@
 
-The code\_aster preCICE adapter is a python file that needs to be placed in the code-aster lib directory.
+The code\_aster preCICE adapter~\cite{Uekermann2017_Adapters} is a python file that needs to be placed in the code-aster lib directory.
 For this, the code\_aster solver needs to be packaged at the latest stable version 14.6.
@@
 
-In order to package this adapter, we first need to package deal.II, which uses CMake as build system and works without any adjustments.
+In order to package this adapter, we first need to package deal.II~\cite{dealII95}, which uses CMake as build system and works without any adjustments.
 The same applies for the adapter, which also uses CMake and needs preCICE as a dependency.
@@
 
-This package uses a combination of CMake files and custom build scripts which makes the build process quite tedious.
+The DUNE~\cite{bastian2020dune} package uses a combination of CMake files and custom build scripts which makes the build process quite tedious.
 This might be an issue because of the ongoing migration to CMake from autotools and might be resolved in the future.
@@
 
-We are explicitly referring to the legacy version of FEniCS here, not the newer and still experimental FEniCSx.
-This adapter uses the open source computing platform FEniCS, which is already packaged within nixpkgs, to provide preCICE support.
+We are explicitly referring to the legacy version of FEniCS~\cite{fenics} here, not the newer and still experimental FEniCSx.
+This adapter~\cite{Rodenberg2021} uses the open source computing platform FEniCS, which is already packaged within nixpkgs, to provide preCICE support.
 
@@
 
-As the preCICE OpenFOAM-adapter is an OpenFOAM module, the derivation of the adapter needs OpenFOAM and wmake as build inputs.
+As the preCICE OpenFOAM-adapter~\cite{OpenFOAMpreCICE} is an OpenFOAM module, the derivation of the adapter needs OpenFOAM and wmake as build inputs.
 The correct environment variables have to be set again, but this is rather easy now, as we have the \texttt{set-openfoam-vars} script that we can simply source during the buildPhase.
@@
 
-This package is a preCICE adapter for the CFD code SU2 and extends the original SU2 source code by patching the files.
+This package is a preCICE adapter~\cite{Uekermann2017_Adapters} for the CFD code SU2 and extends the original SU2 source code by patching the files.
 For this, the adapter provides a custom script that needs to be executed before the build phase.

### 2024-04-20 paper/paper.tex
feat(paper): Add note about success of the case study

@@
 
-In conclusion, the case study has shown that packaging of scientific software isn't as trivial as expected.
+In conclusion, the case study was a success, as it has shown that packaging of scientific software with Nix is possible, but it isn't as trivial as expected.
+Once the software is packaged, it is quite easy to be used and combined in environments with other packaged softwares without conflicts.
+The packaged adapters and solvers work well together with the preCICE library and simulations can be run successfully, so this specific case study has shown the potential of Nix in the domain of scientific software.
+More research has to be done on the topic, yet we are positive about possible future results in Nix abilities and its flexibility in similar domains.
+
 The majority of the packages required some solution for a fundamental issue within the scientific software or its build system that isn't solvable for none experienced Nix users, which makes it hard for people to use it even for their own software.

### 2024-05-02 paper/paper.tex
feat(paper): Add Benjamin as author and state equal contribution
feat(paper): Add Flemisch paper citation and mention xSDK in the intro

@@
 \author{
-Max Hausch\autref{1},
-Simon Hauser\autref{2}}
+Max Hausch\autref{1+},
+Simon Hauser\autref{2+},
+Benjamin Uekermann\autref{3}}
 
@@
 \autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
-\autlabel{2} \email{st148883@stud.uni-stuttgart.de}}
+\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
+\autlabel{3} \email{Benjamin.Uekermann@ipvs.uni-stuttgart.de}\\
+\autlabel{+} These authors contributed equally to this work.}
 
@@
 
-Attempts were made to solve this situation by using Docker\footnote{\url{https://www.docker.com/}}, a software which describes software environments with the help of text files.
+Attempts were made by scientists like Koch et al.~\cite{koch2023sustainable} to solve this situation by using Docker\footnote{\url{https://www.docker.com/}}, a software which describes software environments with the help of text files.
 Those text files are made up of imperative commands which are run inside of containers, one layer at a time.
@@
 This could potentially lead to different results when a user wants to only add a single dependency but has to rebuild the whole image, thus rendering it useless in terms of reproducibility.
+
+Another ongoing effort, to at least make it easier to build scientific software, is made by the xSDK community~\cite{xSDK2023}.
+The xSDK declares policies that a scientific software framework or library has to satisfy to be officially part of the xSDK ecosystem.
 

### 2024-05-03 paper/paper.tex
feat: short title
chore: more friendly wording in intro

@@
 \title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
-\short{TODO}
+\short{preCICE on Nix: A case study}
 \author{
@@
 When a user specifies to install the \texttt{python3} package with a traditional package manager today, the package manager could yield python3.8, whereas running the same command two months from now could yield python in the version 3.9.
-This could potentially lead to different results when a user wants to only add a single dependency but has to rebuild the whole image, thus rendering it useless in terms of reproducibility.
+This could potentially lead to different results when a user wants to add a single dependency but has to rebuild the whole image, thus rendering it inpractical in terms of reproducibility.
 

### 2024-05-03 paper/paper.tex
chore: cleanup distribution table

@@
     \hline
-    \bfseries Package & \bfseries Type & \bfseries Build System & \bfseries Version \\
-    \hline
-    precice & library & cmake & 2.5.0 \\
-    \hline
-    fortran-module & bindings & make & 9e3f405 \\
-    \hline
-    julia & bindings & julia & v2.5.0 \\
-    \hline
-    matlab & bindings & matlab build script & v2.5.0.0 \\
-    \hline
-    pyprecice & bindings & setup.py & 2.5.0.1 \\
-    \hline
-    config-visualizer & tooling & setup.py & 60f2165 \\
-    \hline
-    precice-aste & adapter & cmake & 3.0.0 \\
-    \hline
-    dealii & solver & cmake & 9.4.1 \\
-    \hline
-    precice-dealii-adapter & adapter & cmake & dbb25bea \\
-    \hline
-    precice-calculix-adapter & adapter & makefile & 2.20 \\
-    \hline
-    fenics & solver & cmake + setup.py & 2019.1.0 \\
-    \hline
-    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\
-    \hline
-    precice-dune & adapter & scripts + cmake & 2.8.0 \\
-    \hline
-    Dune-fem & library & setup.py & 2.8.0.0 \\
-    \hline
-    code-aster & solver & custom build system in python & 14.6.0 \\
-    \hline
-    openfoam & solver & wmake/Allwmake & 2206 \\
-    \hline
-    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\
-    \hline
-    Precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\
-    \hline
-    Nutils & library & setup.py & 7.0.0 \\
-    \hline
+    \bfseries Package & \bfseries Type & \bfseries Build System & \bfseries Version \\ \hline
+    precice & library & cmake & 2.5.0 \\ \hline
+    fortran-module & bindings & make & 9e3f405 \\ \hline
+    julia & bindings & julia & v2.5.0 \\ \hline
+    matlab & bindings & matlab build script & v2.5.0.0 \\ \hline
+    pyprecice & bindings & setup.py & 2.5.0.1 \\ \hline
+    config-visualizer & tooling & setup.py & 60f2165 \\ \hline
+    precice-aste & adapter & cmake & 3.0.0 \\ \hline
+    dealii & solver & cmake & 9.4.1 \\ \hline
+    precice-dealii-adapter & adapter & cmake & dbb25bea \\ \hline
+    precice-calculix-adapter & adapter & makefile & 2.20 \\ \hline
+    fenics & solver & cmake + setup.py & 2019.1.0 \\ \hline
+    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\ \hline
+    precice-dune & adapter & scripts + cmake & 2.8.0 \\ \hline
+    Dune-fem & library & setup.py & 2.8.0.0 \\ \hline
+    code-aster & solver & custom build system in python & 14.6.0 \\ \hline
+    openfoam & solver & wmake/Allwmake & 2206 \\ \hline
+    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
+    Precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
+    Nutils & library & setup.py & 7.0.0 \\ \hline
   \end{tabular}

### 2024-05-03 paper/paper.tex
feat: improve comparision with overview tables, ref to fosdem presentation and conclusion

@@
 
-\section{Comparision}
+\section{Comparison}
 
-This chapter defines metrics and uses these metrics to compare the Nix package manager~\cite{Dolstra_2004} with widely used HPC package managers Spack~\cite{Gamblin_2015} and EasyBuild~\cite{Geimer_2014}.
+This chapter compares Nix with two widely used solution in HPC environments, Spack and EasyBuild.
+There was a similar comparison done at FOSDEM 2018~\cite{comparision-fosdem} which also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
+The comparison first takes a high level look on some statistics and how they compare and after that we will discuss, usability, reproducibly and performance in more detail.
 
-\subsection{Metrics definition}
+\subsection{Overview}
 
-In this section, we define all metrics used in the comparison of the selected package managers.
-The metrics are grouped with related metrics.\\
+The first table~\ref{table:comparison-overview} shows some basic stats about the amount of packages the solutions provide out of the box and how they are implemented.
+This is important because the higher the number the more likely it is that your package is already packaged and doesn't require additional work.
+It's also important to look how packages are implemented because, if a package needs to be packaged by for example researches, domain specific languages have the downside that, researches now have to get familiar with this DSL, while general purpose languages like python don't have this downside because they are already used in their field.
+Although this is a downside, DSL usually are design with a purpose in mind and can get you faster to the desired outcome.
 
-\subsubsection{Overview}
-The first few metrics are grouped as Overview, where we present some basic stats about each package manager.
-First, we look at the number of available packages, then at the number of unique packages, meaning all packages minus packages that are available at different versions and last we compare how packages can be implemented.\\
+\begin{figure*}
+  \normalsize
+  \caption{Overview of package count and how packages are implemented}
+  \label{table:comparison-overview}
+  \centering
+  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
+    \hline
+    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
+    Packages~\footnote{\label{foot:repology}\url{https://repology.org/repositories/statistics/total}} & 91800 & 7910 & 19500 \\ \hline
+    Unique packages & 73200 & 6610 & 3470 \\ \hline
+    Package implementation & Nix Domain Specific Language & Python & Python \\ \hline
+  \end{tabular}
+\end{figure*}
 
-\subsubsection{Installation}
-In this section, we look at the installation of all 3 package managers, how they can be updated and if root permissions are needed to either install it or later install packages.\\
+The installing and updating process is documented in the second table~\ref{table:comparison-installation} and differs widely, from custom install scripts over to using pip or just using git clone.
 
-\subsubsection{Usage}
-In Usage, we present basic usages of the package managers.
-We show how you can search for packages, install software, install software with a specific version, install software at a specific point in time, specify the compiler that is used to install a piece of software and last we show how to build optimized software for your architecture.\\
+Spack's way of installing has some problems, with using \texttt{git clone} on the latest default branch, in their case \texttt{develop}, this has the implication, that you aren't using the latest stable release of the software but rather a random commit, that might have broken something recently.
+It also has the problem that you can't easily update the package repository independent from the package manager, this requires additional work from the users if they require this.
 
-\subsubsection{Reproducibility}
-Here we look into reproducibility of build software.
-We first look into how the reproducibility is achieved, then if the architecture and cpu features are considered and latestly we look at the dependencies can affect reproducibility.\\
+It's also worth mentioning that all 3 solutions don't require root for installation but Nix has the option for a system wide installation of the Nix daemon, which then can be used to install packages locally into the users PATH, which no longer requires root.
 
-\subsubsection{Performance}
-We will take a short look at the performance of the package manager and the resulting packages\\
+\begin{figure*}
+  \normalsize
+  \caption{How the package manager are being installed and how are they updated}
+  \label{table:comparison-installation}
+  \centering
+  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
+    \hline
+    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
+    Installation way & install script~\footnote{\url{https://nixos.org/download.html\#download-nix}} & \texttt{git clone}~\footnote{\url{https://spack.readthedocs.io/en/latest/getting_started.html\#installation}} & \texttt{pip install} or afterwards self install using \texttt{eb} executable~\footnote{\url{https://docs.easybuild.io/installation/}} \\ \hline
+    Installation mode & single user/multi user & single user & single user \\ \hline
+    Updating self & \texttt{nix-channel} / nix flake for pinning & \texttt{git pull} & using the \texttt{eb} executable \\ \hline
+  \end{tabular}
+\end{figure*}
 
-\subsubsection{Testing}
-Finally, we look into tests and see what features the package managers provide for validating that a package works as expected.
+The basic usage of all three tools is pretty similar and they support searching and installing package in a simple manner, all this is show in table~\ref{table:comparison-usage}.
+The difference in installing software comes down to the unique way that Nix installs software, it downloads/builds the paths into its local store, which can either be located in the user's home directory or in the path \texttt{/nix/store} depending on the installation.
+\texttt{nix-shell} later only puts these downloaded paths into the \texttt{\$PATH} environment variable.
+Spack and EasyBuild on the other hand use the widely used \texttt{module} system that usually is already found on HPC clusters, it allows loading software into the current environment.
 
-\subsection{Results}
-
-This section presents the results we collected.
-First we will provide a description and afterwards we will discuss the gathered results.\\
-
-\subsubsection{Overview}
-
-For the first set of metrics, we observed that in our comparison Nix provides the most packages with 83000 packages, 67000 of them unique, followed by Spack with around 7000 packages, 6000 of them unique.\footnote{\url{https://repology.org/repositories/statistics/total}}
-Regarding EasyBuild, there are around 17000 and 3000 unique packages.\footnote{\url{https://docs.easybuild.io/release-notes/}}
-
-While Nix has the most amount of packages, it is a general purpose package manager not specifically made for scientific software.
-This means, that most of the packages are not focused on scientific software, but it still comes with important packages like HDF5 and mostly offers mpi support for those packages.
-
-Spack and EasyBuild lay their focus more on the HPC/Simulation environment, so the main purpose of these package managers is to provide scientific software, but that also means that sometimes common packages, like openssl3.0 simply aren't available.
-
-Both, Spack and EasyBuild, use python scripts to implement their packages, Nix on the other hand, uses with its own DSL.
-Using a common language to define packages makes it easier for new people to get started packaging their own software because they might already be familiar with the language and generally, there is more documentation available for general purpose language, especially python.
-The Nix DSL is only useful inside Nix and people looking to package their software for Nix, need to learn a new language that is only applicable within Nix.\\
-
-\subsubsection{Installation}
-
-Nix, the package manager, can be installed with a custom installation script.\footnote{\url{https://nixos.org/download.html\#download-nix}}
-There are multiple ways to install it, with the Nix daemon, making it a multi-user installation, this installation requires root privileges for installation and installs it in \texttt{/nix}, after that normal users can use the Nix binary to communicate with a running daemon to download and install packages, this does not require root permissions.
-The single user installation allows to setup the Nix package manager inside your home directory, this method does not require root for installation and usage.
-The third way is to install Nix as part of NixOS and requires a new installation of the operating system, requires root for installation but not for usage.
-
-Spack is being installed with git and can be cloned into any directory, which means it does not require root for installation and usage.\footnote{\url{https://spack.readthedocs.io/en/latest/getting_started.html\#installation}}
-
-EasyBuild also has multiple ways to install it.\footnote{\url{https://docs.easybuild.io/installation/}}
-It can be installed with \texttt{pip install} (or virtualenv/pipenv) and after that can be loaded with \texttt{module load EasyBuild}.
-It can also be installed into a temporary directory using \texttt{pip} and afterwards properly installed with using the temporary installation to bootstrap an install into the latest release.
-EasyBuild installation requires modifying \texttt{PATH}, \texttt{PYTHONPATH} and potentially \texttt{MODULEPATH} depending on the location, it is being installed in.
-It also requires Tcl/C environment-modules or Tcl-only variant of environment modules or Lmod being installed and configured.
-None the less EasyBuild can be installed without root privileges and generally can also be used without root, but EasyBuild has a package option called \texttt{osdependencies} with requires dependencies being installed on the operating system.
-For example CMake (till 3.18) requires \texttt{openssl-devel}, \texttt{libssl-dev} and \texttt{libopenssl-devel} making it impossible to install for a normal user, if one of the above packages is not installed.
-This also means that users can't install any package that requires the CMake build system without updating the package definitions.
-
-Last but not least, updating the tools and their packages, is also widely different.
-Nix has something that is called channels that need to be advanced upstream and after that users can update their local channels with \texttt{nix-channel --update}, or use the newer Nix flake technology which pins the inputs so users can choose freely when to update to which commit of the upstream nixpkgs repository.
-
-Spack is being updated by fetching a new version of the local repository, which then can lead to merge conflicts if users change the provided package definitions.
-
-EasyBuild can be updated using the \texttt{--install-latest-eb-release} the tool provides.\\
-
-\subsubsection{Usage}
-
-The basic usage of each tool is pretty similar.
-All 3 package managers allow a user to query for a package than then can be installed.
-
-Nix allows for downloading and installing software within the current shell session\footnote{\url{https://nix.dev/tutorials/first-steps/}}, that means that an only accessible within the shell and not actually installed on the system in the classical sense of making it available to multiple users of the system.
-A user or an administrator can at any point \texttt{collect-garbage} and all packages not specified as installed system packages are being removed that includes packages installed in the current shell, as long as the shell session isn't still open.
-If a user wants to install a specific software at a given version there are two main options made available by Nix.
-Either there is an old package version definition provided by the packages, like \texttt{boost159} or a user can opt into overriding the package by overriding the \texttt{version} and the \texttt{src} \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, that will build the package given the specific source, which can be a remote tarball, revision in a repository or even a local directory.
-With these options, users can override any aspect of the package definition, including the dependencies and even providing custom patches.
-Each change to the package will trigger a new build because the input for the build was changed.
-
-Spack and EasyBuild on the other hand, keep the old software versions in the specifications or keep the whole specification, so you can reinstall an old version by specifying the available version, for example, \texttt{spack install gcc@6.4.0}.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}}
+Installing old software also differs for all three solutions, Nix on the one hand keeps some old, widely used software packages around and maintains those, examples are \texttt{boost} which is available in different versions or e.g.
+\texttt{hdf5} which is available in the latest version and version \texttt{1.10}.
+The other way of installing older software on nix is using something that is called \texttt{overrideAttrs}, it can be used to change anything of a package, like the used sources, meaning it can be used to replace it with an older/newer version or it can be used to add an patch, add a new dependency or install additional files into the \texttt{\$out} directory.
+An example would be the following \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, it replaces the precice version with 2.4.0 and adds the old source, which will then trigger a new build.
+Spack and EasyBuild on the other hand, keep the old software versions in the specifications or keep the whole specification, so you can reinstall an old version by specifying the available version, for example, \texttt{spack install gcc@6.4.0}
 If the required version is not available, you have to edit the current package definition and add the required version.
-There is no elegant way of overriding the package in a command call and Spack even provides a command \texttt{spack edit} which does it in tree.
-This can be an issue if you later want to update the git revision, as this might lead to merge conflicts as soon as you checkout a new revision, as described earlier.
 
@@
 
-Next, we look at how you can build a package with a specific compiler version.
+Next, we look at how you can build a package with a specific compiler version using specific compiler flags.
 The flexibility of Nix overrides allows different levels of rebuilds of packages with a specific compiler, you can either just rebuild the requested package with specific compiler, this for example rebuilds curl with clang8 \texttt{nix-shell -p '(curl.override \{ stdenv = clang8Stdenv;\})'} or you can rebuild the package and all required dependencies with a specific compiler by overriding the \texttt{stdenv} on nixpkgs level, \texttt{nix-build -E 'with (import <nixpkgs> \{ config.replaceStdenv = \{ pkgs \}: pkgs.clang8Stdenv; \}); curl'}.
@@
 By default, there are some useful helpers provided, like \texttt{makeStaticLibraries} which allows compiling any library as a static library, or \texttt{enableDebugging} which sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
-
-Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}}
+Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.
 On this note, Spack tooling and interface to override compilers, flags and more is way more user friendly and readable than using and invoking the Nix DSL within a one line command call.
-
 EasyBuild has a whole different system, where packages have defined toolchains.\footnote{\url{https://docs.easybuild.io/common-toolchains/}}
@@
 
-\subsubsection{Reproducibility}
+\begin{figure*}
+  \normalsize
+  \caption{Usage of the package manager}
+  \label{table:comparison-usage}
+  \centering
+  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
+    \hline
+    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
+    Search & \texttt{nix-env -qa 'mpi'}~\footnote{\url{https://nix.dev/tutorials/first-steps/}} & \texttt{spack list 'query'}~\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}} & \texttt{eb -search mpi}~\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}} \\ \hline
+    Installing software & \texttt{nix-shell -p precice} & \texttt{spack install precice \&\& module load ...} & \texttt{eb package.eb -robot \&\& module load ...} \\ \hline
+    Installing specific version of a software & \texttt{nix-shell -p boost159} or \texttt{overrideAttrs} & \texttt{spack install gcc@6.4.0} & find old easybuild config and install that one \\ \hline
+    Installing software at a point in time & \texttt{nix-shell -p pan -I nixpkgs=\$URL} & \texttt{git checkout sha \&\& spack install precice} & \texttt{pip install} older version  \\ \hline
+    Installing software using a specific compiler & \texttt{stdenv} overlay & \texttt{spack install precice \%clang} & \texttt{eb -toolchain=...} \\ \hline
+    Setting compiler options & \texttt{stdenv} overlay & \texttt{spack install ... cflags='-O3 -march=native'} & \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS} and respective fortran env vars \\ \hline
+  \end{tabular}
+\end{figure*}
+
+\subsection{Reproducibility}
 
@@
 It also does something that Nix, currently doesn't support, which is it storing the architecture and cpu features that are used in this build, that way it can provide binary packages that are optimized for specific CPUs.
-None the less it has one problem that Nix doesn't have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies.
-\begin{itemize}
-  \item libm.so
-  \item libc.so
-  \item libdl.so
-  \item libnl-3.so
-  \item libnl-route-3.so
-  \item librt.so
-  \item libutil.so
-  \item libpthread.so
-\end{itemize}
+None the less it has one problem that Nix doesn't have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies \texttt{libm.so}, \texttt{libc.so}, \texttt{libdl.so}, \texttt{libnl-3.so}, \texttt{libnl-route-3.so}, \texttt{librt.so}, \texttt{libutil.so}, \texttt{libpthread.so} as well as the dynamic loader.
 Most of these dependencies are stable, and usually don't change on a system update, but it also means that it simply can't guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
+It is also worth mentioning that the Spack developers are aware of this problem and are currently working on building these system libraries themselves~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
 
@@
 
-\subsubsection{Performance}
+\subsection{Performance}
 
@@
 
-\subsubsection{Testing}
+\subsection{Testing}
 
@@
 The tests are pythons scripts, and it is possible to wait until systemd targets are running, until ports are available or wait until commands are completed and compare the output and status codes.
+
+\subsection{Conclusion}
+
+To summarize the comparison, while Nix has the biggest software library, it lacks some scientific software compared to Spack and EasyBuild.
+On the other hand, the build sandbox that Nix offers, guarantees a level of reproducibility that Spack and EasyBuild currently can't provide, because it disallows random linking.
+Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, like changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it's possible to change the compile flags for all packages in the dependency graph which allows some additional performance gains, if any package is compiled against the specific hardware.
 

### 2024-05-06 paper/paper.tex
feat(paper): Add section about codeocean and wandb

@@
 The xSDK declares policies that a scientific software framework or library has to satisfy to be officially part of the xSDK ecosystem.
+
+There are commercial, domain specifig solutions, e.g.
+CodeOcean\footnote{\url{https://codeocean.com/}} for bioinformatics or Weights and Biases\footnote{\url{https://wandb.ai/site}} for machine learning.
+With these archiving platforms, experiments can be run using technologies like docker so they can be verified by other scientists.
+Those commercial platforms are closed source, so the source code cannot be reviewed nor adjusted as also discussed by~\cite{koch2023sustainable}.
 

### 2024-05-08 paper/paper.tex
Fix affiliation (add more emails if you want)

@@
 \author{
-Max Hausch\autref{1+},
-Simon Hauser\autref{2+},
-Benjamin Uekermann\autref{3}}
+Max Hausch\autref{1}\autref{*},
+Simon Hauser\autref{1}\autref{*},
+Benjamin Uekermann\autref{1}}
 
 \institute{
-\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
-\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
-\autlabel{3} \email{Benjamin.Uekermann@ipvs.uni-stuttgart.de}\\
-\autlabel{+} These authors contributed equally to this work.}
+\autlabel{1} Institute for Parallel and Distributed Systems\\ University of Stuttgart\\ \email{benjamin.uekermann@ipvs.uni-stuttgart.de}\\
+%\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
+%\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
+\autlabel{*} These authors contributed equally to this work.}
 

### 2024-05-08 paper/paper.tex
Tweak and shorten abstract

@@
 
-\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
+\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem}
 \short{preCICE on Nix: A case study}
@@
 \abstract{
-Ensuring the reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
-However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations and software environments.
+Ensuring reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
+However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations, and software environments.
 In this paper, we present a possible solution for these challenges by utilizing Nix and NixOS.
-Nix is a package manager and functional language that allows to mitigate these problems by guaranteeing that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
-NixOS is a purely functional Linux distribution, built on top of Nix that enables the build of reproducible systems including configuration files, packages and their dependencies.
-We present a case study on improving the reproducibility of preCICE, an open-source coupling library, and some of its main adapters using Nix and NixOS.
-Using this approach, we demonstrate how to create a reproducible and self-contained environment for preCICE and highlight the benefits of using Nix and NixOS for managing software and system configurations, resulting in improved reproducibility.
-In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
-This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.}
+Nix is a package manager and functional language, which guarantees that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
+NixOS is a purely functional Linux distribution, built on top of Nix, which enables the build of reproducible systems including configuration files, packages, and their dependencies.
+We study the potential of Nix and NixOS by a case study on the reproducibility of the preCICE ecosystem.
+preCICE is a coupling library for partitioned multiphysics simulations.
+The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
+We demonstrate, how to create a reproducible and self-contained environment for this ecosystem and highlight the benefits of using Nix and NixOS.
+%In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
+%This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.
+}
 

### 2024-05-08 paper/paper.tex
Smooth and shorten introduction

@@
 
-In scientific research, it's crucial to be able to reproduce and verify results.
+In scientific research, it is crucial to be able to reproduce and verify results.
 Reproducibility ensures that experiments can be repeated and findings can be validated, which is essential for reliable and credible research.
-Being able to replicate experiments and computations is important for verifying scientific claims.
-However, achieving reproducibility in scientific software has been a challenge due to complex dependencies, conflicting software environments, and changing software systems.
+However, achieving reproducibility in scientific software has been a challenge due to complex dependencies, conflicting software environments, and changing software systems~\cite{Dalle_2012}.
 Problems arise from dependencies, library versions, and system configurations, leading to inconsistencies across different computing environments.
 Traditional approaches to reproducibility, such as manual setup instructions or virtualization techniques, are prone to errors and time-consuming at best.
-It is not scalable and highly inefficient to make every researcher to tediously recreate the exact same environment with traditional approaches.
-This calls for a more efficient and automated solution.
 
-Attempts were made by scientists like Koch et al.~\cite{koch2023sustainable} to solve this situation by using Docker\footnote{\url{https://www.docker.com/}}, a software which describes software environments with the help of text files.
+Many scientists aim to solve this situation using Docker\footnote{\url{https://www.docker.com/}} (e.g.,\cite{koch2023sustainable}),
+a software which describes software environments with the help of text files.
 Those text files are made up of imperative commands which are run inside of containers, one layer at a time.
-The result are several different layers which are all combines into a single output image that can be instantiated to a running container.
+The result is several different layers combined into a single output image, which can be instantiated into a running container.
 Docker images can be copied to different hosts and should then provide the same environment on different machines.
-An issue that arises here, is that usually those images are based on one of the official Docker images\footnote{\url{https://docs.docker.com/trusted-content/official-images/}} all of which bring traditional package managers like \texttt{apt}.
-When a user specifies to install the \texttt{python3} package with a traditional package manager today, the package manager could yield python3.8, whereas running the same command two months from now could yield python in the version 3.9.
-This could potentially lead to different results when a user wants to add a single dependency but has to rebuild the whole image, thus rendering it inpractical in terms of reproducibility.
+Those images are usually based on one of the official Docker images\footnote{\url{https://docs.docker.com/trusted-content/official-images/}}, however, which use traditional package managers, such as apt.
+When a user then specifies to install the python3 package, for instance, a traditional package manager could yield version 3.8 today, but version 3.9 in a few months.
+Full reproducibility can, thus, only be achieved by storing the complete image. Altering a single dependency (e.g., a bugfix) causes a rebuild of the image and, thus, destroys reproducibility.
 
-Another ongoing effort, to at least make it easier to build scientific software, is made by the xSDK community~\cite{xSDK2023}.
-The xSDK declares policies that a scientific software framework or library has to satisfy to be officially part of the xSDK ecosystem.
+There are, moreover, commercial, domain specifig solutions to achieve reproducibility, e.g., CodeOcean\footnote{\url{https://codeocean.com/}} mainly for bioinformatics or Weights and Biases\footnote{\url{https://wandb.ai/site}} for machine learning.
+With these archiving platforms, experiments can be rerun using technologies such as Docker.
+These platforms are closed source, however, such that the source code cannot be reviewed nor adjusted~\cite{koch2023sustainable}.
 
-There are commercial, domain specifig solutions, e.g.
-CodeOcean\footnote{\url{https://codeocean.com/}} for bioinformatics or Weights and Biases\footnote{\url{https://wandb.ai/site}} for machine learning.
-With these archiving platforms, experiments can be run using technologies like docker so they can be verified by other scientists.
-Those commercial platforms are closed source, so the source code cannot be reviewed nor adjusted as also discussed by~\cite{koch2023sustainable}.
+In past years, the Nix package manager~\cite{Dolstra_2004} and NixOS~\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising alternatives~\cite{Devresse_2015}.
+Nix allows functional descriptions of dependencies up to fixed versions, thus avoiding the issue described above.
+Similar ideas are followed by EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
+All have in common that they rely on scientific software following best practices concerning building and packaging. Unfortunately, most legacy software projects do not do this.
+This is why, for example, the xSDK community~\cite{xSDK2023} tries to set a standard for policies for math software.
 
-In the past years, the Nix package manager~\cite{Dolstra_2004} and NixOS~\cite{Dolstra_2010}, a Linux distribution built around it, have emerged as promising solutions to address these challenges.
-This paper explores the use of Nix/NixOS to improve the reproducibility of scientific software.
-
-Conducting a large scale research on building scientific codes reproducibly is quite infeasible due to how many different software there is.
-Therefore, we conduct a case study, focusing on preCICE~\cite{preCICEv2}, its adapters and solvers.
-One of the goals of the case study is to get a first hand experience and estimate on the complexity and possible challenges when packaging scientific software.
-Also, consistent portability of building the software on three different systems is verified.
-preCICE is a well fit for such a case study, as its official adapters and solvers vary in their programming languages, project sizes and other factors, providing a quite manifold set of scientific simulation software while still being able to verify if the software is still correct by running preCICE simulations across several different solver binaries.
-
-In high-performance computing (HPC), where performance and efficiency are critical, managing software dependencies and configurations becomes even more challenging.
-By using Nix, researchers and practitioners in HPC could easily reproduce computational experiments, ensuring that the same software stack, libraries, and configurations are utilized consistently.
-Reproducibility of scientific results can be achieved using the Nix package manager as described by Devresse et al~\cite{Devresse_2015}.
-This not only streamlines the deployment process but also facilitates collaboration and sharing of software environments, making it easier for researchers to validate and build upon each other's work.
-Additionally, because packages in Nix are highly customizable, optimized builds for specific HPC clusters can be realized.
-
-We also compare Nix, EasyBuild~\cite{easybuil6495863}, and Spack~\cite{spack7832814} as package managers in the realm of scientific software and if Nix is feasible for being used in a scientific context at all.
-
-There are several other papers which discuss and highlight the importance of reproducibly of scientific experiments.
-In their paper, Dalle lists technical factors that lead to issues regarding reproducibly of in silico simulation experiments~\cite{Dalle_2012}.
-\begin{enumerate}
-    \item software bugs: if bugs get fixed in future versions of the same software, they may distort results
-    \item software availability: loosing track of old software versions may lead to unreproducible experiments
-    \item floating points numbers: due to interpretation errors of floating point numbers different simulators could produce different results
-    \item computer and operating systems evolutions: evolving software stacks and dependencies could lead to researchers not being able to reproduce results of an experiment
-\end{enumerate}
-Nix could probably solve three out of these four factors:
-\begin{enumerate}
-    \item using Nix, software --- including its bugs --- can be built reproducibly
-    \item using Nix, even old software can be rebuild and used\footnote{\url{https://blinry.org/nix-time-travel/}}
-    \item Nix cannot really help with floating point issues that affect different codes
-    \item using NixOS, entire operating systems can be reproducibly built, including complete dependencies and kernels
-\end{enumerate}
-
-One suggestion to enhance reproducibility by using a virtual machine (VM) running on a hypervisor software, e.g. VirtualBox\footnote{\url{https://www.virtualbox.org/}} is made by Kim~\cite{Kim_2019}.
-A major advantage compared to experiments with a non-reproducible ad-hoc environment is, that other researchers can simply copy the VirtualBox image containing the whole operating system, including the software stack and all dependencies.
-However, when using the original VM image to build on top of simulations and expanding research, practitioners often have to install several other pieces of software which could cause altering dependencies of existing packages which in turn could lead to falsification of the initial results.
-In this scenario, using Nix and NixOS could address this challenge, because here, dependencies of software stacks are always complete and self-contained.
-So even if researchers want to expand research, new software can be installed alongside existing packages without interference.
+In this paper, we analyze how well Nix and NixOS can improve reproducibility of scientific software.
+To this end, we study the preCICE ecosystem~\cite{preCICEv2} as an example.
+preCICE is a coupling library for partitioned multiphysics simulations.
+The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
+We try to build the complete ecosystem using Nix and verify portability on three different systems.
 

### 2024-05-08 paper/paper.tex
Shorten abstract a bit more

@@
 In this paper, we present a possible solution for these challenges by utilizing Nix and NixOS.
-Nix is a package manager and functional language, which guarantees that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
+Nix is a package manager and functional language, which guarantees that a package and all its dependencies can be built reproducibly.
 NixOS is a purely functional Linux distribution, built on top of Nix, which enables the build of reproducible systems including configuration files, packages, and their dependencies.

### 2024-05-10 paper/paper.tex
fix(paper): it's -> it is / don't -> do not

--- 2024-05-08 paper/paper.tex
@@
 
-\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem}
+\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
 \short{preCICE on Nix: A case study}
 \author{
-Max Hausch\autref{1}\autref{*},
-Simon Hauser\autref{1}\autref{*},
-Benjamin Uekermann\autref{1}}
+Max Hausch\autref{1+},
+Simon Hauser\autref{2+},
+Benjamin Uekermann\autref{3}}
 
 \institute{
-\autlabel{1} Institute for Parallel and Distributed Systems\\ University of Stuttgart\\ \email{benjamin.uekermann@ipvs.uni-stuttgart.de}\\
-%\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
-%\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
-\autlabel{*} These authors contributed equally to this work.}
+\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
+\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
+\autlabel{3} \email{Benjamin.Uekermann@ipvs.uni-stuttgart.de}\\
+\autlabel{+} These authors contributed equally to this work.}
 
 \abstract{
-Ensuring reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
-However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations, and software environments.
+Ensuring the reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
+However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations and software environments.
 In this paper, we present a possible solution for these challenges by utilizing Nix and NixOS.
-Nix is a package manager and functional language, which guarantees that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
-NixOS is a purely functional Linux distribution, built on top of Nix, which enables the build of reproducible systems including configuration files, packages, and their dependencies.
-We study the potential of Nix and NixOS by a case study on the reproducibility of the preCICE ecosystem.
-preCICE is a coupling library for partitioned multiphysics simulations.
-The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
-We demonstrate, how to create a reproducible and self-contained environment for this ecosystem and highlight the benefits of using Nix and NixOS.
-%In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
-%This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.
-}
+Nix is a package manager and functional language that allows to mitigate these problems by guaranteeing that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
+NixOS is a purely functional Linux distribution, built on top of Nix that enables the build of reproducible systems including configuration files, packages and their dependencies.
+We present a case study on improving the reproducibility of preCICE, an open-source coupling library, and some of its main adapters using Nix and NixOS.
+Using this approach, we demonstrate how to create a reproducible and self-contained environment for preCICE and highlight the benefits of using Nix and NixOS for managing software and system configurations, resulting in improved reproducibility.
+In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
+This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.}
 
@@
 Nix enables users to have multiple versions of the same software installed at the same time without the softwares raising conflicts.
-As each build output references their complete dependency graphs inside the Nix store, they don't interfere with other build outputs.
+As each build output references their complete dependency graphs inside the Nix store, they do not interfere with other build outputs.
 An example of a graphical representation of a dependency graph can be seen in Fig. \ref{fig:nix-graph}.
@@
 
-The preCICE library itself is already available in the nixpkgs repository, so we don't have to package the software.
+The preCICE library itself is already available in the nixpkgs repository, so we do not have to package the software.
 Looking at the package definition and the preCICE source code, the library is quite easy to build and package as it uses CMake as a build system.
@@
 The preCICE adapter for CSM code CalculiX~\cite{Uekermann2017_Adapters}, requires the original CalculiX source code and provides a new Makefile in the adapter repository.
-By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it's possible to override this location with a make variable.
+By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
 It then builds the original code and the adapter code together, which results in a combined binary.
@@
 
-Another dependency, which is also provided by code\_aster is metis, which partially builds, so we do not replace this dependency, but we need to fix the second part of the build, which includes the metis programs like gpmetis, ndmetis, mpmetis and more.
+Another dependency, which is also provided by code\_aster is metis, which partially builds, so we do not replace this dependency, but we need to fix the second part of the build, which includes the metis programs such as gpmetis, ndmetis, mpmetis and more.
 The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \texttt{/home/karypis/local/lib}.
@@
 
-Mumps builds using a Makefile and can be configured using Makefile.inc, by default it provides a couple of these configurations for systems like Ubuntu, so we use one of these provided configuration files to write our own custom config file.
-It defines the location for dependencies like scotch, metis, parmetis, liblapack, blas, scalapack and blacs.
+Mumps builds using a Makefile and can be configured using Makefile.inc, by default it provides a couple of these configurations for systems such as Ubuntu, so we use one of these provided configuration files to write our own custom config file.
+It defines the location for dependencies such as scotch, metis, parmetis, liblapack, blas, scalapack and blacs.
 Almost all the dependencies, are available upstream, but we need to recompile scotch with additional build flags, \texttt{scotch ptscotch esmumps ptesmumps} using Nix overrideAttrs feature.
@@
 
-It's also a minor issue that the dune-project is a collection of repositories rather than a monorepository, because users have to correctly clone all repositories so the build system finds all relevant information.
+It is also a minor issue that the dune-project is a collection of repositories rather than a monorepository, because users have to correctly clone all repositories so the build system finds all relevant information.
 We clone all of the required dune repos into a directory and additionally clone the adapter into the same directory.
@@
 We used Nix to define a VM image with NixOS as a base, that can be built reproducibly.
-The image contains all the adapters and solvers from the case study and some additional custom tools that can only be seen in the official, like the \texttt{preciceToPNG} command.
+The image contains all the adapters and solvers from the case study and some additional custom tools that can only be seen in the official VM, such as the \texttt{preciceToPNG} command.
 Also, we generate an iso and a Vagrant VirtualBox file of the VM as additional outputs.
 
-Most of the functionality that can be found in the official VM is also available in our NixOS VM and preCICE tutorials, like the \textit{perpendicular-flap} can be run successfully.
+Most of the functionality that can be found in the official VM is also available in our NixOS VM and preCICE tutorials, such as the \textit{perpendicular-flap} can be run successfully.
 
@@
 It is also worth mentioning that writing your own custom build system, requires additional time to maintain this build system, which could have been spent working on the software itself.
-Packages that use common, industry proven package managers, like CMake and autotools make it easier for developers of the software to build their software as well as package managers to package it.
+Packages that use common, industry proven package managers, such as CMake and autotools make it easier for developers of the software to build their software as well as package managers to package it.
 These build tools, generally also provide interfaces for dependency management which can be used to tell users and package maintainers which dependencies are needed, and it is usually better than only documenting required dependencies, because documentation needs to be kept in sync with the underlying code.
@@
 If a package supports all these policies, they can be added to the growing list of packages in the xSDK, which currently among other things include PETSc, deal.ii and preCICE.
-These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes like nix store paths, and common build systems like CMake and autoconf.
+These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes such as nix store paths, and common build systems such as CMake and autoconf.
 
@@
 Additionally, hard-coding libraries to \texttt{/usr/lib} and \texttt{/usr/include} might work on one distribution, but might make it hard or even impossible to install a package on a different distribution or even a later version of the same distribution.
-Solutions like \texttt{pkg-config} that can be configured, so it finds libraries outside the common paths, should always be preferred and work out of the box with the common paths like \texttt{/usr/lib}.
+Solutions such as \texttt{pkg-config} that can be configured, so it finds libraries outside the common paths, should always be preferred and work out of the box with the common paths such as \texttt{/usr/lib}.
 
@@
 This is important because the higher the number the more likely it is that your package is already packaged and doesn't require additional work.
-It's also important to look how packages are implemented because, if a package needs to be packaged by for example researches, domain specific languages have the downside that, researches now have to get familiar with this DSL, while general purpose languages like python don't have this downside because they are already used in their field.
+It is also important to look how packages are implemented because, if a package needs to be packaged by for example researches, domain specific languages have the downside that, researches now have to get familiar with this DSL, while general purpose languages such as python do not have this downside because they are already used in their field.
 Although this is a downside, DSL usually are design with a purpose in mind and can get you faster to the desired outcome.
@@
 
-It's also worth mentioning that all 3 solutions don't require root for installation but Nix has the option for a system wide installation of the Nix daemon, which then can be used to install packages locally into the users PATH, which no longer requires root.
+It is also worth mentioning that all 3 solutions do not require root for installation but Nix has the option for a system wide installation of the Nix daemon, which then can be used to install packages locally into the users PATH, which no longer requires root.
 
@@
 \texttt{hdf5} which is available in the latest version and version \texttt{1.10}.
-The other way of installing older software on nix is using something that is called \texttt{overrideAttrs}, it can be used to change anything of a package, like the used sources, meaning it can be used to replace it with an older/newer version or it can be used to add an patch, add a new dependency or install additional files into the \texttt{\$out} directory.
+The other way of installing older software on nix is using something that is called \texttt{overrideAttrs}, it can be used to change anything of a package, such as the used sources, meaning it can be used to replace it with an older/newer version or it can be used to add an patch, add a new dependency or install additional files into the \texttt{\$out} directory.
 An example would be the following \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, it replaces the precice version with 2.4.0 and adds the old source, which will then trigger a new build.
@@
 This is possible because every c/c++/\ldots package, is build using the \texttt{stdenv} which defines the compiler and default compile flags.
-While Nix provides a default stdenv, which is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, like the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
-By default, there are some useful helpers provided, like \texttt{makeStaticLibraries} which allows compiling any library as a static library, or \texttt{enableDebugging} which sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
+While Nix provides a default stdenv, which is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, such as the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
+By default, there are some useful helpers provided, such as \texttt{makeStaticLibraries} which allows compiling any library as a static library, or \texttt{enableDebugging} which sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
 Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.
@@
 A toolchain is a set of packages defined at a specific version and users can create their own toolchains.
-For example, EasyBuild provides a foss toolchain which contains packages like, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version like foss-2022b.
+For example, EasyBuild provides a foss toolchain which contains packages such as, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version such as foss-2022b.
 Packages are then provided with toolchains in mind and are never deleted, so installing a package with a different compiler can be achieved by selecting a different eb script with a different toolchain version.
@@
 
-As previously mentioned, Nix computes an output hash over all its inputs that is unique for this input, so if anything in the inputs changes, like the version of a dependency, the output hash changes resulting in a different software build.
-This guarantees that given the same set of inputs, we are able to compile the same software, including it's dependency tree as you can see in Fig. \ref{fig:nix-tree}.
-One thing to mention is, that Nix currently only builds generic architecture build, so the inputs of a derivation don't include any information about the architecture and/or compiler features.
+As previously mentioned, Nix computes an output hash over all its inputs that is unique for this input, so if anything in the inputs changes, such as the version of a dependency, the output hash changes resulting in a different software build.
+This guarantees that given the same set of inputs, we are able to compile the same software, including it is dependency tree as you can see in Fig. \ref{fig:nix-tree}.
+One thing to mention is, that Nix currently only builds generic architecture build, so the inputs of a derivation do not include any information about the architecture and/or compiler features.
 Another thing to mention is, that all software in the \texttt{/nix/store} can only link back to the \texttt{/nix/store} itself.
@@
 None the less it has one problem that Nix doesn't have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies \texttt{libm.so}, \texttt{libc.so}, \texttt{libdl.so}, \texttt{libnl-3.so}, \texttt{libnl-route-3.so}, \texttt{librt.so}, \texttt{libutil.so}, \texttt{libpthread.so} as well as the dynamic loader.
-Most of these dependencies are stable, and usually don't change on a system update, but it also means that it simply can't guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
+Most of these dependencies are stable, and usually do not change on a system update, but it also means that it simply can't guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
 It is also worth mentioning that the Spack developers are aware of this problem and are currently working on building these system libraries themselves~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
@@
 This leads to the problem of binaries linking back to the operating system \texttt{/usr/lib}, that Spack also has.
-For example a build of curl, uses libssl, libcrypto, libz, libm and libc, all of which are provided by the operating system, and for EasyBuild it includes libraries that change more frequently like libssl, libcrypto and libz.\\
+For example a build of curl, uses libssl, libcrypto, libz, libm and libc, all of which are provided by the operating system, and for EasyBuild it includes libraries that change more frequently such as libssl, libcrypto and libz.\\
 
@@
 The advantage of overriding the stdenv is that every package in a dependency tree can be optimized which might lead to better overall performance of packages, but it also has the downside that every package needs to be rebuild which might take several days, for example preCICE required a rebuild of 1500 packages, which took 2 days on an high-performance computing (HPC) cluster (2x AMD EPYC 7702).
-So it might be sufficient to optimize only the package, which is easily possible with \texttt{(package.override)} like shown before or additionally optimize its direct dependencies, which is possible but more time-consuming to do, because one has to manually override each dependency.
+So it might be sufficient to optimize only the package, which is easily possible with \texttt{(package.override)} such as shown before or additionally optimize its direct dependencies, which is possible but more time-consuming to do, because one has to manually override each dependency.
 
@@
 
-Like previously mentioned, EasyBuild applies \texttt{-march=native} by default, but you can further customize your compiler flags using \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS}, \texttt{\$FFLAGS}, \texttt{\$F90FLAGS}.
+EasyBuild applies \texttt{-march=native} by default, but you can further customize your compiler flags using \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS}, \texttt{\$FFLAGS}, \texttt{\$F90FLAGS}.
 It also has the same problem as Spack that it might have unavoidable unoptimized system package which might impact performance negatively.\\
@@
 Nix has something that is called a checkPhase which can be run for each package after being built.
-There are sane defaults for all available build systems, like \texttt{make test} or \texttt{ctest} for makefiles and CMake, \texttt{cargo check} and \texttt{cargo test} for rust packages, \texttt{go test} for go packages and \texttt{pytest} for python packages and more.
-By default, this checkPhase is not enabled for every build system, like the \texttt{stdenv} build system, which is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
+There are sane defaults for all available build systems, such as \texttt{make test} or \texttt{ctest} for makefiles and CMake, \texttt{cargo check} and \texttt{cargo test} for rust packages, \texttt{go test} for go packages and \texttt{pytest} for python packages and more.
+By default, this checkPhase is not enabled for every build system, such as the \texttt{stdenv} build system, which is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
 For these build systems, package maintainers have to manually disable the check phase or they can override the check phase if the tests aren't working out of the box and make the needed adjustments to resolve the issues.
@@
 On the other hand, the build sandbox that Nix offers, guarantees a level of reproducibility that Spack and EasyBuild currently can't provide, because it disallows random linking.
-Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, like changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it's possible to change the compile flags for all packages in the dependency graph which allows some additional performance gains, if any package is compiled against the specific hardware.
+Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, such as changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it is possible to change the compile flags for all packages in the dependency graph which allows some additional performance gains, if any package is compiled against the specific hardware.
 
@@
 We'd need administrator rights to be allowed to setup the Nix store globally on the cluster and to install Nix globally for all users.
-However, Nix supports a Linux kernel feature called user namespaces, which is a form of kernel based isolation of resources and other attributes like user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.
+However, Nix supports a Linux kernel feature called user namespaces, which is a form of kernel based isolation of resources and other attributes such as user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.
 
@@
 
-Lastly we looked at Nix on HPC and have seen that it's possible to get it running, and we are able to run simulations on one node, so we currently assume that with the correct networking setup it should be possible to run a simulation across multiple nodes.
+Lastly we looked at Nix on HPC and have seen that it is possible to get it running, and we are able to run simulations on one node, so we currently assume that with the correct networking setup it should be possible to run a simulation across multiple nodes.
 It is currently not ideally solved on getting Nix running on HPC as a single user installation, so this can be improved in the future.
@@
 While this case study and this comparison has provided valuable insights into the topic of reproducible software, there are still numerous unanswered questions that merit further investigation.
-This includes a full evaluation if there are performance improvements using a full impure system rebuild, with \texttt{-march=native} compared to only rebuilding the package at hand like we suspect.
+This includes a full evaluation if there are performance improvements using a full impure system rebuild, with \texttt{-march=native} compared to only rebuilding the package at hand.
 It would also be interesting to look at trying to reproduce a nixpkgs commit from 10 years ago, and if there might be issues and incompatibilities with newer Nix versions, which currently get released in a 6 weeks time windows.

### 2024-05-10 paper/paper.tex
fix(paper): Expand all short formulations of ' and replace "which"

@@
 
-preCICE is an open-source project, which means its source code is freely available and can be modified and distributed by the community.
+preCICE is an open-source project, that means its source code is freely available and can be modified and distributed by the community.
 This fosters collaboration and encourages the development of new features and improvements by researchers and engineers worldwide.
@@
 
-To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution~\cite{preciceDistribution}, which is currently preCICE \texttt{v2211.0}.
+To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution~\cite{preciceDistribution}, currently preCICE \texttt{v2211.0}.
 The preCICE distribution consists out of the preCICE core library, as well as some tools, bindings, adapters, tutorials, the website and documentation, and the preCICE VM\footnote{\url{https://precice.org/installation-distribution.html}}.
@@
 
-This package is a collection of tools which help in development of new adapters.
+This package is a collection of tools that help in development of new adapters.
 It stands for Artificial Solver Testing Environment, so it allows for debugging adapter mapping setups and other similar tasks.
 
-ASTE\footnote{\url{https://github.com/precice/aste}} needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
+ASTE\footnote{\url{https://github.com/precice/aste}} needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, that is built without python support in the nixpkgs repository to reduce compilation time.
 However, this feature is needed for ASTE.
@@
 
-A preCICE distribution comes bundled with several adapters, which are currently not in the upstream nixpkgs repository.
-Many of the adapters use solvers that aren't packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.
+A preCICE distribution comes bundled with several adapters, that are currently not in the upstream nixpkgs repository.
+Many of the adapters use solvers that are not packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.
 
@@
 By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
-It then builds the original code and the adapter code together, which results in a combined binary.
+It then builds the original code and the adapter code together, resulting in a combined binary.
 
-The Makefile also has hard-coded dependency location for \texttt{SPOOLES}, \texttt{ARPACK} and \texttt{YAML} so these need to be replaced with the equivalent \texttt{pkg-config} calls, because Nix doesn't have these dependencies at the hard-coded locations, and generally only finds dependencies with \texttt{pkg-config}.
+The Makefile also has hard-coded dependency location for \texttt{SPOOLES}, \texttt{ARPACK} and \texttt{YAML} so these need to be replaced with the equivalent \texttt{pkg-config} calls, because Nix does not have these dependencies at the hard-coded locations, and generally only finds dependencies with \texttt{pkg-config}.
 Additionally, there is no \texttt{install} target provided by the Makefile so the resulting binary needs to be installed manually by copying it to \texttt{\$out}.\\
@@
 
-With this approach, we disable HDF5, which is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
+With this approach, we disable HDF5, as it is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
 HDF5 is packaged with multiple outputs.
 One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, this is, because most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
-Additionally, code\_aster requires medfile and scotch, which are both available upstream and can be used as a dependency.
+Additionally, code\_aster requires medfile and scotch, that are both available upstream and can be used as a dependency.
 
-Another dependency, which is also provided by code\_aster is metis, which partially builds, so we do not replace this dependency, but we need to fix the second part of the build, which includes the metis programs such as gpmetis, ndmetis, mpmetis and more.
+Another dependency, that is also provided by code\_aster is metis, that partially builds, so we do not replace this dependency, but we need to fix the second part of the build, including the metis programs such as gpmetis, ndmetis, mpmetis and more.
 The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \texttt{/home/karypis/local/lib}.
@@
 Almost all the dependencies, are available upstream, but we need to recompile scotch with additional build flags, \texttt{scotch ptscotch esmumps ptesmumps} using Nix overrideAttrs feature.
-For packaging blacs, which also uses a Makefile, we configure the build using a custom \texttt{Bmake.inc} file, which is used to set compiler flags, the path to the mpi and bash installation.
+For packaging blacs, that also uses a Makefile, we configure the build using a custom \texttt{Bmake.inc} file, that is used to set compiler flags, the path to the mpi and bash installation.
 
-The last missing dependency for code\_aster, tfel, uses CMake as a build system which makes packaging it trivial with Nix and we only have to set a couple of CMake flags.
+The last missing dependency for code\_aster, tfel, uses CMake as a build system making it trivial to package it with Nix and we only have to set a couple of CMake flags.
 
-After successfully packaging all dependencies, the build phase of code\_aster completes, but the installation part has another issue, that also affects new versions of Ubuntu and other distributions which use a python version greater than 3.9.
+After successfully packaging all dependencies, the build phase of code\_aster completes, but the installation part has another issue, that also affects new versions of Ubuntu and other distributions that use a python version greater than 3.9.
 There is a forum post without any resolution from the developers or the community, so we need to manually patch the bug.
 
-The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, which works if your version is 3.9.x but not with 3.10.x because then the code calculates \texttt{lib64/python3.1/site-packages} rather than the expected \texttt{lib64/python3.10/site-packages}.
+The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, that works if your version is 3.9.x but not with 3.10.x because then the code calculates \texttt{lib64/python3.1/site-packages} rather than the expected \texttt{lib64/python3.10/site-packages}.
 
@@
 
-Last but not least, we can copy the code\_aster adapter at the required location which concludes this adapter installation.\\
+Last but not least, we can copy the code\_aster adapter at the required location.\\
 
@@
 
-In order to package this adapter, we first need to package deal.II~\cite{dealII95}, which uses CMake as build system and works without any adjustments.
-The same applies for the adapter, which also uses CMake and needs preCICE as a dependency.
+In order to package this adapter, we first need to package deal.II~\cite{dealII95}, that uses CMake as build system and works without any adjustments.
+The same applies for the adapter, that also uses CMake and needs preCICE as a dependency.
 When supplied with the inputs, the adapter builds without any problems.
@@
 
-The DUNE~\cite{bastian2020dune} package uses a combination of CMake files and custom build scripts which makes the build process quite tedious.
+The DUNE~\cite{bastian2020dune} package uses a combination of CMake files and custom build scripts making the build process quite tedious.
 This might be an issue because of the ongoing migration to CMake from autotools and might be resolved in the future.
@@
 For the build and install, process we need to manually set \texttt{\$DUNE\_CONTROL\_PATH} and \texttt{\$DUNE\_PY\_DIR} environment variables, both variables also need to be set at runtime.
-Additionally, we need to patch the python install process because the current CMake file that handles this install tries to access the internet with \texttt{pip install} which isn't allowed in Nix' sandboxed builds.
+Additionally, we need to patch the python install process because the current CMake file that handles this install tries to access the internet with \texttt{pip install}.
+This is not allowed in Nix' sandboxed builds.
 Therefore, we need to patch this \texttt{pip} command with the options \texttt{--no-build-isolation --no-cache-dir --no-index --no-deps} and we need to remove the \texttt{--upgrade} option.\\
@@
 We are explicitly referring to the legacy version of FEniCS~\cite{fenics} here, not the newer and still experimental FEniCSx.
-This adapter~\cite{Rodenberg2021} uses the open source computing platform FEniCS, which is already packaged within nixpkgs, to provide preCICE support.
+This adapter~\cite{Rodenberg2021} uses the open source computing platform FEniCS, that is already packaged within nixpkgs, to provide preCICE support.
 
 Although it is available, not all features the adapter needs are enabled, including PETSc support.
-To fix this, we need to package PETSc4py, which provides python bindings for PETSc and is not present in the upstream nixpkgs repository.
+To fix this, we need to package PETSc4py, that provides python bindings for PETSc and is not present in the upstream nixpkgs repository.
 The build process is possible without any issues, because it uses the internal \texttt{buildPythonApplication} build tool, but we need to add the additional build optional \texttt{build\_src --force} because the cython code needs to be rebuilt.
-We also aren't able to enable tests for this package because tests depend on OpenMPI and the network, which isn't available within the Nix build sandbox.
+We also are not able to enable tests for this package because tests depend on OpenMPI and the network, that is not available within the Nix build sandbox.
 
@@
 
-Building the adapter with Nix is now possible with \texttt{buildPythonPackage} and doesn't require any special workarounds.
-A majority of the \texttt{pytest} ran successfully, but two tests failed inside the Nix sandbox, which means all tests needed to be disabled.\\
+Building the adapter with Nix is now possible with \texttt{buildPythonPackage} and does not require any special workarounds.
+A majority of the \texttt{pytest} ran successfully, but two tests failed inside the Nix sandbox, that means all tests needed to be disabled.\\
 
@@
 
-OpenFOAM uses its own custom build system called \texttt{wmake} which is sometimes called with a wrapper script called \texttt{Allwmake}.
+OpenFOAM uses its own custom build system called \texttt{wmake} that is sometimes called with a wrapper script called \texttt{Allwmake}.
 The build system sets 36 environment variables that are set by wmake, one step at a time by checking several parameters, e.g.
@@
 During runtime, before running any OpenFOAM commands, a small wrapper script has to be started that sets all relevant environment variables, or a user has to manually source a file inside the installation directory of OpenFOAM.
-OpenFOAM uses modules for extensibility and flexibility which need the source code of OpenFOAM and also require wmake as a build tool.
+OpenFOAM uses modules for extensibility and flexibility that need the source code of OpenFOAM and also require wmake as a build tool.
 
@@
 During the buildPhase, the rendered script will be sourced to set all variables, e.g.
-so the \texttt{OPENFOAM\_SRC\_PATH} points to \texttt{/build/openfoam} which is the default location for Nix builds inside the sandbox.
+so the \texttt{OPENFOAM\_SRC\_PATH} points to \texttt{/build/openfoam} being the default location for Nix builds inside the sandbox.
 After sourcing the script, the single command \texttt{./Allwmake -j -q} is sufficient to start the build.
@@
 
-OpenFOAM is available on Spack and EasyBuild which both patch the OpenFOAM source files as those tools also have to set environment variables to make a build compatible with wmake.\\
+OpenFOAM is available on Spack and EasyBuild that both patch the OpenFOAM source files as those tools also have to set environment variables to make a build compatible with wmake.\\
 
@@
 For this, the adapter provides a custom script that needs to be executed before the build phase.
-Nix offers the patchPhase which allows us to run the script in this phase and patch the original source, after that the \texttt{stdenv} automatically recognizes that the repository uses autotools for building the software and runs the configurePhase.
+Nix offers the patchPhase that allows us to run the script in this phase and patch the original source, after that the \texttt{stdenv} automatically recognizes that the repository uses autotools for building the software and runs the configurePhase.
 
@@
 OpenFOAM and Allwmake are prime examples of these issues because now package managers have to understand a whole new build system for exactly one package in order to successfully package it and make it easily available to users.
-It is also worth mentioning that writing your own custom build system, requires additional time to maintain this build system, which could have been spent working on the software itself.
+It is also worth mentioning that writing your own custom build system, requires additional time to maintain this build system, that could have been spent working on the software itself.
 Packages that use common, industry proven package managers, such as CMake and autotools make it easier for developers of the software to build their software as well as package managers to package it.
-These build tools, generally also provide interfaces for dependency management which can be used to tell users and package maintainers which dependencies are needed, and it is usually better than only documenting required dependencies, because documentation needs to be kept in sync with the underlying code.
+These build tools, generally also provide interfaces for dependency management that can be used to tell users and package maintainers what dependencies are needed, and it is usually better than only documenting required dependencies, because documentation needs to be kept in sync with the underlying code.
 They also generally already provide proven interfaces for configuring packages with additional features, and these also might be self documenting.
@@
 Their goal is to ``identify, adapt, and adopt best practices in software engineering''~\cite{xsdk-website} and ``to provide the foundation of this extensible scientific software ecosystem''~\cite{xsdk-website}.
-To achieve this they provide a list of package policies~\cite{xSDK2023}, which for example specifies as mandatory policy that packages ``must support portable installation through Spack''~\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''~\cite{xSDK2023}, which includes CMake and autoconf as examples.
-If a package supports all these policies, they can be added to the growing list of packages in the xSDK, which currently among other things include PETSc, deal.ii and preCICE.
+To achieve this they provide a list of package policies~\cite{xSDK2023}, that for example specifies as mandatory policy that packages ``must support portable installation through Spack''~\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''~\cite{xSDK2023}, that includes CMake and autoconf as examples.
+If a package supports all these policies, they can be added to the growing list of packages in the xSDK, that currently among other things include PETSc, deal.ii and preCICE.
 These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes such as nix store paths, and common build systems such as CMake and autoconf.
@@
 This chapter compares Nix with two widely used solution in HPC environments, Spack and EasyBuild.
-There was a similar comparison done at FOSDEM 2018~\cite{comparision-fosdem} which also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
+There was a similar comparison done at FOSDEM 2018~\cite{comparision-fosdem} that also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
 The comparison first takes a high level look on some statistics and how they compare and after that we will discuss, usability, reproducibly and performance in more detail.
@@
 The first table~\ref{table:comparison-overview} shows some basic stats about the amount of packages the solutions provide out of the box and how they are implemented.
-This is important because the higher the number the more likely it is that your package is already packaged and doesn't require additional work.
+This is important because the higher the number the more likely it is that your package is already packaged and does not require additional work.
 It is also important to look how packages are implemented because, if a package needs to be packaged by for example researches, domain specific languages have the downside that, researches now have to get familiar with this DSL, while general purpose languages such as python do not have this downside because they are already used in their field.
@@
 
-Spack's way of installing has some problems, with using \texttt{git clone} on the latest default branch, in their case \texttt{develop}, this has the implication, that you aren't using the latest stable release of the software but rather a random commit, that might have broken something recently.
-It also has the problem that you can't easily update the package repository independent from the package manager, this requires additional work from the users if they require this.
+Spack's way of installing has some problems, with using \texttt{git clone} on the latest default branch, in their case \texttt{develop}, this has the implication, that you are not using the latest stable release of the software but rather a random commit, that might have broken something recently.
+It also has the problem that you cannot easily update the package repository independent from the package manager, this requires additional work from the users if they require this.
 
-It is also worth mentioning that all 3 solutions do not require root for installation but Nix has the option for a system wide installation of the Nix daemon, which then can be used to install packages locally into the users PATH, which no longer requires root.
+It is also worth mentioning that all 3 solutions do not require root for installation but Nix has the option for a system wide installation of the Nix daemon, that can then be used to install packages locally into the users PATH, without requiring root permissions.
 
@@
 The basic usage of all three tools is pretty similar and they support searching and installing package in a simple manner, all this is show in table~\ref{table:comparison-usage}.
-The difference in installing software comes down to the unique way that Nix installs software, it downloads/builds the paths into its local store, which can either be located in the user's home directory or in the path \texttt{/nix/store} depending on the installation.
+The difference in installing software comes down to the unique way that Nix installs software, it downloads/builds the paths into its local store, that can either be located in the user's home directory or in the path \texttt{/nix/store} depending on the installation.
 \texttt{nix-shell} later only puts these downloaded paths into the \texttt{\$PATH} environment variable.
@@
 
-Installing old software also differs for all three solutions, Nix on the one hand keeps some old, widely used software packages around and maintains those, examples are \texttt{boost} which is available in different versions or e.g.
-\texttt{hdf5} which is available in the latest version and version \texttt{1.10}.
+Installing old software also differs for all three solutions, Nix on the one hand keeps some old, widely used software packages around and maintains those, examples are \texttt{boost} that is available in different versions or e.g.
+\texttt{hdf5} that is available in the latest version and version \texttt{1.10}.
 The other way of installing older software on nix is using something that is called \texttt{overrideAttrs}, it can be used to change anything of a package, such as the used sources, meaning it can be used to replace it with an older/newer version or it can be used to add an patch, add a new dependency or install additional files into the \texttt{\$out} directory.
-An example would be the following \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, it replaces the precice version with 2.4.0 and adds the old source, which will then trigger a new build.
+An example would be the following \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, it replaces the preCICE version with 2.4.0 and adds the old source, that will then trigger a new build.
 Spack and EasyBuild on the other hand, keep the old software versions in the specifications or keep the whole specification, so you can reinstall an old version by specifying the available version, for example, \texttt{spack install gcc@6.4.0}
@@
 Spack, can achieve something similar by checking out your Spack version at a previous revision and then installing the package requested package, but this approach has a downside that it affects the whole Spack installation, and now also work with an older Spack tool.
-On the other hand, EasyBuild, could be installed using pip at an earlier version, but it also offers something different called toolchains, which we will look at shortly.
+On the other hand, EasyBuild, could be installed using pip at an earlier version, but it also offers something different called toolchains, we will look at shortly.
 
@@
 The flexibility of Nix overrides allows different levels of rebuilds of packages with a specific compiler, you can either just rebuild the requested package with specific compiler, this for example rebuilds curl with clang8 \texttt{nix-shell -p '(curl.override \{ stdenv = clang8Stdenv;\})'} or you can rebuild the package and all required dependencies with a specific compiler by overriding the \texttt{stdenv} on nixpkgs level, \texttt{nix-build -E 'with (import <nixpkgs> \{ config.replaceStdenv = \{ pkgs \}: pkgs.clang8Stdenv; \}); curl'}.
-This is possible because every c/c++/\ldots package, is build using the \texttt{stdenv} which defines the compiler and default compile flags.
-While Nix provides a default stdenv, which is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, such as the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
-By default, there are some useful helpers provided, such as \texttt{makeStaticLibraries} which allows compiling any library as a static library, or \texttt{enableDebugging} which sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
+This is possible because every c/c++/\ldots package, is build using the \texttt{stdenv} that defines the compiler and default compile flags.
+While Nix provides a default stdenv, that is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, such as the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
+By default, there are some useful helpers provided, such as \texttt{makeStaticLibraries} that allows compiling any library as a static library, or \texttt{enableDebugging} that sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
 Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.
@@
 A toolchain is a set of packages defined at a specific version and users can create their own toolchains.
-For example, EasyBuild provides a foss toolchain which contains packages such as, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version such as foss-2022b.
+For example, EasyBuild provides a foss toolchain that contains packages such as, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version such as foss-2022b.
 Packages are then provided with toolchains in mind and are never deleted, so installing a package with a different compiler can be achieved by selecting a different eb script with a different toolchain version.
-On the other hand this results in a less elegant user flow, because if a user wants to install a package with a toolchain that isn't provided they have to fork a package definition and update the toolchain.\\
+On the other hand this results in a less elegant user flow, because if a user wants to install a package with a toolchain that is not provided they have to fork a package definition and update the toolchain.\\
 
@@
 
-Spack does something similar to Nix where it adds hashes to each installation path, but also stores additional files inside the package installation which include the complete spec of the package than can later be used to reproducibly build the same package.~\cite{Gamblin_2015}.
-It also does something that Nix, currently doesn't support, which is it storing the architecture and cpu features that are used in this build, that way it can provide binary packages that are optimized for specific CPUs.
-None the less it has one problem that Nix doesn't have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies \texttt{libm.so}, \texttt{libc.so}, \texttt{libdl.so}, \texttt{libnl-3.so}, \texttt{libnl-route-3.so}, \texttt{librt.so}, \texttt{libutil.so}, \texttt{libpthread.so} as well as the dynamic loader.
-Most of these dependencies are stable, and usually do not change on a system update, but it also means that it simply can't guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
+Spack does something similar to Nix where it adds hashes to each installation path, but also stores additional files inside the package installation including the complete spec of the package than can later be used to reproducibly build the same package.~\cite{Gamblin_2015}.
+It also does something that Nix, currently lacks support for, namely storing the architecture and cpu features that are used in this build.
+That way it can provide binary packages that are optimized for specific CPUs.
+None the less it has one problem that Nix does not have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies \texttt{libm.so}, \texttt{libc.so}, \texttt{libdl.so}, \texttt{libnl-3.so}, \texttt{libnl-route-3.so}, \texttt{librt.so}, \texttt{libutil.so}, \texttt{libpthread.so} as well as the dynamic loader.
+Most of these dependencies are stable, and usually do not change on a system update, but it also means that it simply cannot guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
 It is also worth mentioning that the Spack developers are aware of this problem and are currently working on building these system libraries themselves~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
@@
 This leads to the problem of binaries linking back to the operating system \texttt{/usr/lib}, that Spack also has.
-For example a build of curl, uses libssl, libcrypto, libz, libm and libc, all of which are provided by the operating system, and for EasyBuild it includes libraries that change more frequently such as libssl, libcrypto and libz.\\
+For instance, a curl build, uses the libraries libssl, libcrypto, libz, libm and libc, all of them being provided by the operating system, and for EasyBuild it includes libraries that change more frequently such as libssl, libcrypto and libz.\\
 
@@
 With Nix' possibility to override stdenv it is also possible to set custom compiler flags and Nix provides an option \texttt{impureUseNativeOptimizations} to build packages optimized for your hardware, using the \texttt{-march=native} compiler flag, it can then be applied to the package, that is being installed, or every dependency in the dependency graph.~\footnote{\url{https://nixos.wiki/wiki/Build_flags}}
-Nix can guarantee these compiler flags for every dependency because Nix wraps its compiler binaries in scripts, which then means it has more control over them and packages can not remove any flags by clearing environment variables.
-The advantage of overriding the stdenv is that every package in a dependency tree can be optimized which might lead to better overall performance of packages, but it also has the downside that every package needs to be rebuild which might take several days, for example preCICE required a rebuild of 1500 packages, which took 2 days on an high-performance computing (HPC) cluster (2x AMD EPYC 7702).
-So it might be sufficient to optimize only the package, which is easily possible with \texttt{(package.override)} such as shown before or additionally optimize its direct dependencies, which is possible but more time-consuming to do, because one has to manually override each dependency.
+Nix can guarantee these compiler flags for every dependency because Nix wraps its compiler binaries in scripts, meaning it has more control over them and packages can not remove any flags by clearing environment variables.
+The advantage of overriding the stdenv is that every package in a dependency tree can be optimized, potentially leading to better overall performance of packages.
+But it also has the downside that every package needs to be rebuild potentially taking several days, for example preCICE requires a rebuild of 1500 packages, running for two days on an high-performance computing (HPC) cluster (2x AMD EPYC 7702).
+So it might be sufficient to optimize only the package, that is easily possible with \texttt{(package.override)} such as shown before or additionally optimize its direct dependencies, that is possible but more time-consuming to do, because one has to manually override each dependency.
 
@@
 Additionally, there is a way to define flags for a specific registered compiler, but these are only respected by CMake and autotools packages.
-Also, a package might also have unavoidable system dependencies and are not optimized, which might impact performance negatively.
+Also, a package might also have unavoidable system dependencies and are not optimized, potentially impacting performance negatively.
 
 EasyBuild applies \texttt{-march=native} by default, but you can further customize your compiler flags using \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS}, \texttt{\$FFLAGS}, \texttt{\$F90FLAGS}.
-It also has the same problem as Spack that it might have unavoidable unoptimized system package which might impact performance negatively.\\
+It also has the same problem as Spack that it might have unavoidable unoptimized system package that might impact performance negatively.\\
 
@@
 We also shortly look at the capabilities of the package managers to run test and evaluate the built packages.
-Nix has something that is called a checkPhase which can be run for each package after being built.
+Nix has something that is called a checkPhase that can be run for each package after being built.
 There are sane defaults for all available build systems, such as \texttt{make test} or \texttt{ctest} for makefiles and CMake, \texttt{cargo check} and \texttt{cargo test} for rust packages, \texttt{go test} for go packages and \texttt{pytest} for python packages and more.
-By default, this checkPhase is not enabled for every build system, such as the \texttt{stdenv} build system, which is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
-For these build systems, package maintainers have to manually disable the check phase or they can override the check phase if the tests aren't working out of the box and make the needed adjustments to resolve the issues.
+By default, this checkPhase is not enabled for every build system, such as the \texttt{stdenv} build system, that is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
+For these build systems, package maintainers have to manually disable the check phase or they can override the check phase if the tests are not working out of the box and make the needed adjustments to resolve the issues.
 
@@
 To summarize the comparison, while Nix has the biggest software library, it lacks some scientific software compared to Spack and EasyBuild.
-On the other hand, the build sandbox that Nix offers, guarantees a level of reproducibility that Spack and EasyBuild currently can't provide, because it disallows random linking.
-Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, such as changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it is possible to change the compile flags for all packages in the dependency graph which allows some additional performance gains, if any package is compiled against the specific hardware.
+On the other hand, the build sandbox that Nix offers, guarantees a level of reproducibility that Spack and EasyBuild currently cannot provide, because it disallows random linking.
+Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, such as changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it is possible to change the compile flags for all packages in the dependency graph allowing for additional performance gains, if any package is compiled against the specific hardware.
 
@@
 
-As mentioned, Nix uses the Nix store, which lies in \texttt{/nix/store} by default and all software artifacts are linked against libraries inside the Nix store.
+As mentioned, Nix uses the Nix store, that lies in \texttt{/nix/store} by default and all software artifacts are linked against libraries inside the Nix store.
 So if there is no such path as \texttt{/nix/store} on the HPC clusters filesystem, all software would need to be recompiled as all links and paths inside the binaries would point to a non existent location.
-We'd need administrator rights to be allowed to setup the Nix store globally on the cluster and to install Nix globally for all users.
-However, Nix supports a Linux kernel feature called user namespaces, which is a form of kernel based isolation of resources and other attributes such as user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.
+We would need administrator rights to be allowed to setup the Nix store globally on the cluster and to install Nix globally for all users.
+However, Nix supports a Linux kernel feature called user namespaces.
+This is a form of kernel based isolation of resources and other attributes such as user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.
 
@@
 We can simply use a statically compiled version the Nix package manager, however, to circumvent this problem.
-In our repository, we provide a setup bash script that simplifies download, configuration and installation of the static Nix binary which can then be used to easily build and access the solvers adapters and the preCICE VM.
+In our repository, we provide a setup bash script that simplifies download, configuration and installation of the static Nix binary that can then be used to easily build and access the solvers adapters and the preCICE VM.
 
@@
 
-In conclusion, the case study was a success, as it has shown that packaging of scientific software with Nix is possible, but it isn't as trivial as expected.
+In conclusion, the case study was a success, as it has shown that packaging of scientific software with Nix is possible, but it is not as trivial as expected.
 Once the software is packaged, it is quite easy to be used and combined in environments with other packaged softwares without conflicts.
@@
 
-The majority of the packages required some solution for a fundamental issue within the scientific software or its build system that isn't solvable for none experienced Nix users, which makes it hard for people to use it even for their own software.
+The majority of the packages required some solution for a fundamental issue within the scientific software or its build system that is not solvable for none experienced Nix users, making it hard for people to use it even for their own software.
 Nix itself also is partly at fault, because of its unique nature, that every piece of software lies within \texttt{/nix/store} and that each store path is its own-isolated tree based on the Filesystem Hierarchy Standard (FHS).
@@
 The work of the xSDK people and their policy might be a step in the right direction but looking at the current list of software that is available within the latest xSDK release, 26 libraries, indicates that it is an uphill battle that might take several more years, especially if we consider that the first release happened in 2016.
-We have also seen, that the current HPC solutions, Spack and EasyBuild make it somewhat possible to reproducibly build software but not to the degree in which Nix allows it.
+We have also seen, that the current HPC solutions, Spack and EasyBuild make it somewhat possible to reproducibly build software but not to the degree that Nix allows it.
 That is because they focus on flexibility of installing packages.
@@
 This includes a full evaluation if there are performance improvements using a full impure system rebuild, with \texttt{-march=native} compared to only rebuilding the package at hand.
-It would also be interesting to look at trying to reproduce a nixpkgs commit from 10 years ago, and if there might be issues and incompatibilities with newer Nix versions, which currently get released in a 6 weeks time windows.
+It would also be interesting to look at trying to reproduce a nixpkgs commit from 10 years ago, and if there might be issues and incompatibilities with newer Nix versions, that currently get released in a 6 weeks time windows.
 

### 2024-05-10 paper/paper.tex

@@
 
-Nix is a general purpose build tool and purely functional package manager that provides some features that help to build software derivations reproducibly.
+Nix is a purely functional package manager that provides some features to build software derivations reproducibly.
 It does so, by recursively calculating a hash over all the inputs of a derivation and its dependencies to ensure the completeness of the whole derivation.
@@
 
+As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
+Inputs to those functions are parameters such as the package version or the location of the source code.
+Evaluating and realizing those functions with the same inputs, yields the same outputs.
+This is a critical factor where Nix's reproducibility comes from.
+Outputs, i.e. the build artifacts, never change after being built once.
+Nix builds are sandboxed, meaning that there is no internet access possible during a build.
+
 All contents, including source files and resulting build artifacts, are stored inside the Nix store.
-Per default, the Nix store resides in the path \texttt{/nix/store} on the file system and all file- and directory names stored inside the Nix store are prefixed with a cryptographic hash of the inputs of the specific file/directory followed by a human-readable name and a version number.
-This eases checking for integrity the Nix store and ensures that builds using the same inputs are performed only once.
-
-These are two examples for possible directory names for the preCICE 2.5.0 package:
+Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
+The naming scheme with the preCICE 2.5.0 package as an example is
 
@@
 
-Hashing the inputs and storing them in the manner described above provides several features that other package managers cannot provide.
-Nix enables users to have multiple versions of the same software installed at the same time without the softwares raising conflicts.
-As each build output references their complete dependency graphs inside the Nix store, they do not interfere with other build outputs.
+This eases checking the Nix store for integrity and ensures that builds using the same inputs are performed only once.
+
+As each build output references their complete dependency graphs inside the Nix store, they do not interfere with other build outputs, enabling users to have multiple versions of the same software installed at the same time without raising conflicts.
 An example of a graphical representation of a dependency graph can be seen in Fig. \ref{fig:nix-graph}.
-Nix modifies environment variables, such as the \texttt{PATH} variable, to make calling the binaries more convenient to users.
-This fact allows multiple users to safely install software without the need of elevated permissions, as users cannot exchange the binaries of other users.
 
@@
 
-Package derivations for the Nix package manager are written in a domain specific language which is also called Nix.
-Derivations are the ``build plans'' on how Nix should build a specific software, which resulting build artifacts should be stored where, what the inputs are and so forth.
-There are several different phases during a build, which contain shell code that is being run one phase after the other.
-Those phases default to a given behavior, depending on the build system used.
-All phases can be overridden freely to give users a maximum of flexibility when building and packaging software for Nix.
-The most important phases users will override are the \texttt{patchPhase}, \texttt{buildPhase}, \texttt{installPhase} and \texttt{checkPhase}.
-During the build, a special bash variable \texttt{\$out} is set that contains the path to the target directory inside the Nix store.
-
-As the Nix language is purely functional, evaluating and building the derivation with the same inputs, yields the same outputs.
-This is a critical factor where Nix's reproducibility comes from.
-Also, Nix builds are sandboxed so they can reproducibly be run, meaning that there is no internet access possible during a build.
-
-Another benefit of the reproducibility are remote builds, so a build cluster can build packages and serve its Nix store to other machines running Nix.
-The upstream build cluster is called ``hydra'' and Nix defaults to using the upstream hydra as a substituter, i.e. fetching all packages from the hydras binary cache.
-When evaluating a derivation, Nix checks if the hashed path is existent in the local Nix store.
-If it's not, then it queries all known binary caches, which per default is only the upstream binary cache, and fetches the path and all its dependent paths until the full closure is copied into the local Nix store.
-
 Software patches are quite easy to apply in Nix.
-One can simply provide a list of \texttt{.patch} files inside the Nix file describing a derivation, Nix will then include the patch files in the calculation of the hash.
-This new store path can probably not be found in the upstream binary cache, so Nix will fall back to building from source.
-Nix also provides overriding all inputs freely, so every package can be flexibly adjusted to the users needs.
+One can simply provide a list of \texttt{.patch} files as inputs, Nix will then include the patch files in the build and the calculation of the hash.
+Users can override all inputs freely, so every package can be flexibly adjusted to the users needs.
 

### 2024-05-10 paper/paper.tex

@@
 
-\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on preCICE adapters and solvers}
+\title{Improving reproducibility of scientific software using Nix/NixOS: A case study on the preCICE ecosystem}
 \short{preCICE on Nix: A case study}
 \author{
-Max Hausch\autref{1+},
-Simon Hauser\autref{2+},
-Benjamin Uekermann\autref{3}}
+Max Hausch\autref{1}\autref{*},
+Simon Hauser\autref{1}\autref{*},
+Benjamin Uekermann\autref{1}}
 
 \institute{
-\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
-\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
-\autlabel{3} \email{Benjamin.Uekermann@ipvs.uni-stuttgart.de}\\
-\autlabel{+} These authors contributed equally to this work.}
+\autlabel{1} Institute for Parallel and Distributed Systems\\ University of Stuttgart\\ \email{benjamin.uekermann@ipvs.uni-stuttgart.de}\\
+%\autlabel{1} \email{st175425@stud.uni-stuttgart.de}\\
+%\autlabel{2} \email{st148883@stud.uni-stuttgart.de}\\
+\autlabel{*} These authors contributed equally to this work.}
 
 \abstract{
-Ensuring the reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
-However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations and software environments.
+Ensuring reproducibility of scientific software is crucial for the advancement of research and the validation of scientific findings.
+However, achieving reproducibility in software-intensive scientific projects is often challenging due to dependencies, system configurations, and software environments.
 In this paper, we present a possible solution for these challenges by utilizing Nix and NixOS.
-Nix is a package manager and functional language that allows to mitigate these problems by guaranteeing that a package and all its dependencies can be built reproducibly as long as there is a build plan at the desired time.
-NixOS is a purely functional Linux distribution, built on top of Nix that enables the build of reproducible systems including configuration files, packages and their dependencies.
-We present a case study on improving the reproducibility of preCICE, an open-source coupling library, and some of its main adapters using Nix and NixOS.
-Using this approach, we demonstrate how to create a reproducible and self-contained environment for preCICE and highlight the benefits of using Nix and NixOS for managing software and system configurations, resulting in improved reproducibility.
-In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
-This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.}
+Nix is a package manager and functional language, which guarantees that a package and all its dependencies can be built reproducibly.
+NixOS is a purely functional Linux distribution, built on top of Nix, which enables the build of reproducible systems including configuration files, packages, and their dependencies.
+We study the potential of Nix and NixOS by a case study on the reproducibility of the preCICE ecosystem.
+preCICE is a coupling library for partitioned multiphysics simulations.
+The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
+We demonstrate, how to create a reproducible and self-contained environment for this ecosystem and highlight the benefits of using Nix and NixOS.
+%In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
+%This evaluation enables us to assess the advantages and disadvantages of employing Nix to improve reproducibility in scientific software development within an HPC context.
+}
 

### 2024-05-10 paper/paper.tex
chore: split long lines on sentence end for better diffing
feat(paper): Make Discussion its own chapter
drop(paper): Chapter on HPC

@@
 Similar ideas are followed by EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
-All have in common that they rely on scientific software following best practices concerning building and packaging. Unfortunately, most legacy software projects do not do this.
+All have in common that they rely on scientific software following best practices concerning building and packaging.
+Unfortunately, most legacy software projects do not do this.
 This is why, for example, the xSDK community~\cite{xSDK2023} tries to set a standard for policies for math software.
@@
 
-\subsection{Discussion}
+\section{Discussion}
 
@@
 
-\section{Nix on HPC}\label{sec:nix-on-hpc}
-
-High-performance computing refers to the utilization of powerful computational resources to tackle computationally intensive scientific problems.
-It involves the parallel execution of tasks across multiple computing units, also called nodes in this context, enabling researchers to perform complex simulations, data analysis, and modeling at large scales.
-HPC systems, such as clusters, supercomputers, and cloud-based infrastructures, usually offer immense processing power, storage capacity, and specialized hardware, supporting scientific research across various domains\footnote{\url{https://www.ibm.com/topics/hpc}}.
-
-In many HPC environments, the \texttt{module} system is being used to load different pieces of software from a local repository of available softwares, such as compilers and different message passing interfaces (MPI).
-With the module command, different users can use different software or different versions of the same software at the same time without interfering with each other.
-One major drawback here is, that without further configuration, users can only load software modules that were previously installed into the cluster by the clusters administrators, so any other software that is needed for a simulation, including its dependencies, have to be compiled by the users that want to conduct an experiment.
-Sometimes, this can be quite hard, as some software expects libraries or other software artifacts at some hard-coded location as \texttt{/usr/lib/\ldots}\footnote{\url{https://hpc-wiki.info/hpc/Modules}}.
-
-As mentioned, Nix uses the Nix store, that lies in \texttt{/nix/store} by default and all software artifacts are linked against libraries inside the Nix store.
-So if there is no such path as \texttt{/nix/store} on the HPC clusters filesystem, all software would need to be recompiled as all links and paths inside the binaries would point to a non existent location.
-We would need administrator rights to be allowed to setup the Nix store globally on the cluster and to install Nix globally for all users.
-However, Nix supports a Linux kernel feature called user namespaces.
-This is a form of kernel based isolation of resources and other attributes such as user ids\footnote{\url{https://www.man7.org/linux/man-pages/man7/user_namespaces.7.html}}.
-
-Using namespaces, it is possible for the Nix package manager to mount an arbitrary user-writable location to \texttt{/nix/store} inside a user namespace, so all programs running in that namespace can access the Nix store as if it would actually be installed under the default store path.
-Another challenge is to get Nix to work without the needed libraries that can usually be found inside the Nix store, leading to a bootstrapping problem.
-We can simply use a statically compiled version the Nix package manager, however, to circumvent this problem.
-In our repository, we provide a setup bash script that simplifies download, configuration and installation of the static Nix binary that can then be used to easily build and access the solvers adapters and the preCICE VM.
-
-After installing Nix on HPC, we were able to produce the same binary output for the adapters and solvers as we built on our local machines as well as within the preCICE VM\footnote{The sha256 hashes of the resulting outputs were compared and matched.}
-
 \section{Conclusion and Outlook}

### 2024-05-10 paper/paper.tex
reformat(paper): Case study -> Results

@@
 
-\section{Case study}
+\section{Results}
 
@@
 The authors of the xSDK\footnote{\url{https://xsdk.info/}} came to a very similar conclusion.
-Their goal is to ``identify, adapt, and adopt best practices in software engineering''~\cite{xsdk-website} and ``to provide the foundation of this extensible scientific software ecosystem''~\cite{xsdk-website}.
+Their goal is to ``identify, adapt, and adopt best practices in software engineering'' and ``to provide the foundation of this extensible scientific software ecosystem''~\footnote{\url{https://xsdk.info/}}.
 To achieve this they provide a list of package policies~\cite{xSDK2023}, that for example specifies as mandatory policy that packages ``must support portable installation through Spack''~\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''~\cite{xSDK2023}, that includes CMake and autoconf as examples.
@@
 This chapter compares Nix with two widely used solution in HPC environments, Spack and EasyBuild.
-There was a similar comparison done at FOSDEM 2018~\cite{comparision-fosdem} that also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
+There was a similar comparison done at FOSDEM 2018~\footnote{\url{https://archive.fosdem.org/2018/schedule/event/installing_software_for_scientists/}} that also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
 The comparison first takes a high level look on some statistics and how they compare and after that we will discuss, usability, reproducibly and performance in more detail.

### 2024-05-10 paper/paper.tex
feat: cleanup bibtex and move bibtex to footnotes
feat(paper): Make dot output use less space

@@
     \centering
-    \includegraphics[width=0.4\textwidth]{figures/coreutils.png}
-    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -ocoreutils.png}}
+    \includegraphics[width=1\textwidth]{figures/coreutils.png}
+    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -Grankdir=LR -ocoreutils.png}}
     \label{fig:nix-graph}

### 2024-05-10 paper/paper.tex
drop(paper): Example precice nix paths
chore: drop comparison

@@
 Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
-The naming scheme with the preCICE 2.5.0 package as an example is
-
-\begin{itemize}
-  \item \texttt{/nix/store/0a5gw3l\ldots-precice-2.5.0}
-  \item \texttt{/nix/store/9gpyg37\ldots-precice-2.5.0}
-\end{itemize}
-
+The naming scheme for the preCICE 2.5.0 package as an example is \texttt{/nix/store/0a5gw3l\ldots-precice-2.5.0}.
 This eases checking the Nix store for integrity and ensures that builds using the same inputs are performed only once.
@@
 
-\section{Comparison}
-
-This chapter compares Nix with two widely used solution in HPC environments, Spack and EasyBuild.
-There was a similar comparison done at FOSDEM 2018~\footnote{\url{https://archive.fosdem.org/2018/schedule/event/installing_software_for_scientists/}} that also included conda and Guix, but this was by the release manager of EasyBuild and we want to provide an update with this comparison.
-The comparison first takes a high level look on some statistics and how they compare and after that we will discuss, usability, reproducibly and performance in more detail.
-
-\subsection{Overview}
-
-The first table~\ref{table:comparison-overview} shows some basic stats about the amount of packages the solutions provide out of the box and how they are implemented.
-This is important because the higher the number the more likely it is that your package is already packaged and does not require additional work.
-It is also important to look how packages are implemented because, if a package needs to be packaged by for example researches, domain specific languages have the downside that, researches now have to get familiar with this DSL, while general purpose languages such as python do not have this downside because they are already used in their field.
-Although this is a downside, DSL usually are design with a purpose in mind and can get you faster to the desired outcome.
-
-\begin{figure*}
-  \normalsize
-  \caption{Overview of package count and how packages are implemented}
-  \label{table:comparison-overview}
-  \centering
-  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
-    \hline
-    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
-    Packages~\footnote{\label{foot:repology}\url{https://repology.org/repositories/statistics/total}} & 91800 & 7910 & 19500 \\ \hline
-    Unique packages & 73200 & 6610 & 3470 \\ \hline
-    Package implementation & Nix Domain Specific Language & Python & Python \\ \hline
-  \end{tabular}
-\end{figure*}
-
-The installing and updating process is documented in the second table~\ref{table:comparison-installation} and differs widely, from custom install scripts over to using pip or just using git clone.
-
-Spack's way of installing has some problems, with using \texttt{git clone} on the latest default branch, in their case \texttt{develop}, this has the implication, that you are not using the latest stable release of the software but rather a random commit, that might have broken something recently.
-It also has the problem that you cannot easily update the package repository independent from the package manager, this requires additional work from the users if they require this.
-
-It is also worth mentioning that all 3 solutions do not require root for installation but Nix has the option for a system wide installation of the Nix daemon, that can then be used to install packages locally into the users PATH, without requiring root permissions.
-
-\begin{figure*}
-  \normalsize
-  \caption{How the package manager are being installed and how are they updated}
-  \label{table:comparison-installation}
-  \centering
-  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
-    \hline
-    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
-    Installation way & install script~\footnote{\url{https://nixos.org/download.html\#download-nix}} & \texttt{git clone}~\footnote{\url{https://spack.readthedocs.io/en/latest/getting_started.html\#installation}} & \texttt{pip install} or afterwards self install using \texttt{eb} executable~\footnote{\url{https://docs.easybuild.io/installation/}} \\ \hline
-    Installation mode & single user/multi user & single user & single user \\ \hline
-    Updating self & \texttt{nix-channel} / nix flake for pinning & \texttt{git pull} & using the \texttt{eb} executable \\ \hline
-  \end{tabular}
-\end{figure*}
-
-The basic usage of all three tools is pretty similar and they support searching and installing package in a simple manner, all this is show in table~\ref{table:comparison-usage}.
-The difference in installing software comes down to the unique way that Nix installs software, it downloads/builds the paths into its local store, that can either be located in the user's home directory or in the path \texttt{/nix/store} depending on the installation.
-\texttt{nix-shell} later only puts these downloaded paths into the \texttt{\$PATH} environment variable.
-Spack and EasyBuild on the other hand use the widely used \texttt{module} system that usually is already found on HPC clusters, it allows loading software into the current environment.
-
-Installing old software also differs for all three solutions, Nix on the one hand keeps some old, widely used software packages around and maintains those, examples are \texttt{boost} that is available in different versions or e.g.
-\texttt{hdf5} that is available in the latest version and version \texttt{1.10}.
-The other way of installing older software on nix is using something that is called \texttt{overrideAttrs}, it can be used to change anything of a package, such as the used sources, meaning it can be used to replace it with an older/newer version or it can be used to add an patch, add a new dependency or install additional files into the \texttt{\$out} directory.
-An example would be the following \texttt{nix-shell -p 'precice.overrideAttrs(\_: \{ version = "2.4.0"; src = \ldots \} )'}, it replaces the preCICE version with 2.4.0 and adds the old source, that will then trigger a new build.
-Spack and EasyBuild on the other hand, keep the old software versions in the specifications or keep the whole specification, so you can reinstall an old version by specifying the available version, for example, \texttt{spack install gcc@6.4.0}
-If the required version is not available, you have to edit the current package definition and add the required version.
-
-To achieve reproducibility, it is necessary that we can install software from a previous revision of the packages.
-Nix allows us to simply specify the channel revision we want to use to build our software, it then evaluates the requested packages and builds it.
-This can easily be done without affecting the current system by specifying it in the shell command \texttt{nix-shell -p pan -I nixpkgs=pathToRevision/urlToTarball}, it now only uses this channel for building the requested package.
-Spack, can achieve something similar by checking out your Spack version at a previous revision and then installing the package requested package, but this approach has a downside that it affects the whole Spack installation, and now also work with an older Spack tool.
-On the other hand, EasyBuild, could be installed using pip at an earlier version, but it also offers something different called toolchains, we will look at shortly.
-
-Next, we look at how you can build a package with a specific compiler version using specific compiler flags.
-The flexibility of Nix overrides allows different levels of rebuilds of packages with a specific compiler, you can either just rebuild the requested package with specific compiler, this for example rebuilds curl with clang8 \texttt{nix-shell -p '(curl.override \{ stdenv = clang8Stdenv;\})'} or you can rebuild the package and all required dependencies with a specific compiler by overriding the \texttt{stdenv} on nixpkgs level, \texttt{nix-build -E 'with (import <nixpkgs> \{ config.replaceStdenv = \{ pkgs \}: pkgs.clang8Stdenv; \}); curl'}.
-This is possible because every c/c++/\ldots package, is build using the \texttt{stdenv} that defines the compiler and default compile flags.
-While Nix provides a default stdenv, that is used to compile all packages, it is possible to override and replace the package with your another stdenv provided by nixpkgs, such as the clang8Stdenv, you can write your own stdenv, or you can extend the current stdenv with flags.
-By default, there are some useful helpers provided, such as \texttt{makeStaticLibraries} that allows compiling any library as a static library, or \texttt{enableDebugging} that sets the compiler flags \texttt{-ggdb -Og} and can then be enabled on the package plus all dependencies.
-Spack provides an easy way to specify the compiler using \texttt{spack install precice \%clang} and only requires a registered compiler, it can be a system compiler, or the compiler can be installed using Spack.
-On this note, Spack tooling and interface to override compilers, flags and more is way more user friendly and readable than using and invoking the Nix DSL within a one line command call.
-EasyBuild has a whole different system, where packages have defined toolchains.\footnote{\url{https://docs.easybuild.io/common-toolchains/}}
-A toolchain is a set of packages defined at a specific version and users can create their own toolchains.
-For example, EasyBuild provides a foss toolchain that contains packages such as, binutils, gcc, OpenMPI, openblas and more and packages are provided against a specific toolchain version such as foss-2022b.
-Packages are then provided with toolchains in mind and are never deleted, so installing a package with a different compiler can be achieved by selecting a different eb script with a different toolchain version.
-On the other hand this results in a less elegant user flow, because if a user wants to install a package with a toolchain that is not provided they have to fork a package definition and update the toolchain.\\
-
-\begin{figure*}
-  \normalsize
-  \caption{Usage of the package manager}
-  \label{table:comparison-usage}
-  \centering
-  \begin{tabular}{|p{2cm}|p{3.5cm}|p{3.5cm}|p{3.5cm}|}
-    \hline
-    \textbf{Metrics} & \textbf{Nix} & \textbf{Spack} & \textbf{EasyBuild} \\ \hline
-    Search & \texttt{nix-env -qa 'mpi'}~\footnote{\url{https://nix.dev/tutorials/first-steps/}} & \texttt{spack list 'query'}~\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}} & \texttt{eb -search mpi}~\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html}} \\ \hline
-    Installing software & \texttt{nix-shell -p precice} & \texttt{spack install precice \&\& module load ...} & \texttt{eb package.eb -robot \&\& module load ...} \\ \hline
-    Installing specific version of a software & \texttt{nix-shell -p boost159} or \texttt{overrideAttrs} & \texttt{spack install gcc@6.4.0} & find old easybuild config and install that one \\ \hline
-    Installing software at a point in time & \texttt{nix-shell -p pan -I nixpkgs=\$URL} & \texttt{git checkout sha \&\& spack install precice} & \texttt{pip install} older version  \\ \hline
-    Installing software using a specific compiler & \texttt{stdenv} overlay & \texttt{spack install precice \%clang} & \texttt{eb -toolchain=...} \\ \hline
-    Setting compiler options & \texttt{stdenv} overlay & \texttt{spack install ... cflags='-O3 -march=native'} & \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS} and respective fortran env vars \\ \hline
-  \end{tabular}
-\end{figure*}
-
-\subsection{Reproducibility}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.4\textwidth]{figures/nix-tree.png}
-    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --tree \$(nix build nixpkgs\#coreutils --print-out-paths)}}
-    \label{fig:nix-tree}
-\end{figure}
-
-As previously mentioned, Nix computes an output hash over all its inputs that is unique for this input, so if anything in the inputs changes, such as the version of a dependency, the output hash changes resulting in a different software build.
-This guarantees that given the same set of inputs, we are able to compile the same software, including it is dependency tree as you can see in Fig. \ref{fig:nix-tree}.
-One thing to mention is, that Nix currently only builds generic architecture build, so the inputs of a derivation do not include any information about the architecture and/or compiler features.
-Another thing to mention is, that all software in the \texttt{/nix/store} can only link back to the \texttt{/nix/store} itself.
-Nix validates this for every package and breaks links than allowing links back to a file in the home directory or the directory that was used for building the software.
-This is required, to guarantee the best possible reproducibly.
-
-Spack does something similar to Nix where it adds hashes to each installation path, but also stores additional files inside the package installation including the complete spec of the package than can later be used to reproducibly build the same package.~\cite{Gamblin_2015}.
-It also does something that Nix, currently lacks support for, namely storing the architecture and cpu features that are used in this build.
-That way it can provide binary packages that are optimized for specific CPUs.
-None the less it has one problem that Nix does not have: it allows packages to link back into system dependencies, for example, while building hpl-2.3 using a gcc 12.2.0 compiled with Spack we see the usage of the following system dependencies \texttt{libm.so}, \texttt{libc.so}, \texttt{libdl.so}, \texttt{libnl-3.so}, \texttt{libnl-route-3.so}, \texttt{librt.so}, \texttt{libutil.so}, \texttt{libpthread.so} as well as the dynamic loader.
-Most of these dependencies are stable, and usually do not change on a system update, but it also means that it simply cannot guarantee the best possible reproducibility and it gets worse if the system compiler was used rather than a compiler build from scratch using Spack.
-It is also worth mentioning that the Spack developers are aware of this problem and are currently working on building these system libraries themselves~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
-
-EasyBuild does achieve reproducibility with its toolchains and the approach to keep old package definitions, meaning that they also keep old dependency definitions and each dependency is pinned at a specific version~\cite{Geimer_2014}.
-Regarding, CPU architecture and features, EasyBuild does not do generic builds by default but rather prioritizes the idea of optimizing for a specific system, meaning defaulting to \texttt{-march=native} for all packages and it has to manually be disabled~\footnote{\url{https://docs.easybuild.io/controlling-compiler-optimization-flags/\#controlling\_compiler\_optimization\_flags\_optarch\_default}}.
-This leads to inherently impure and irreproducible builds as building the same easyconfig on two different host/architectures results in different binaries.
-On the other hand, the already mentioned option for \texttt{osdependencies} goes against the reproducibility idea, because any os dependency is not managed by the package manager and might change at any point.
-This leads to the problem of binaries linking back to the operating system \texttt{/usr/lib}, that Spack also has.
-For instance, a curl build, uses the libraries libssl, libcrypto, libz, libm and libc, all of them being provided by the operating system, and for EasyBuild it includes libraries that change more frequently such as libssl, libcrypto and libz.\\
-
-\subsection{Performance}
-
-With Nix' possibility to override stdenv it is also possible to set custom compiler flags and Nix provides an option \texttt{impureUseNativeOptimizations} to build packages optimized for your hardware, using the \texttt{-march=native} compiler flag, it can then be applied to the package, that is being installed, or every dependency in the dependency graph.~\footnote{\url{https://nixos.wiki/wiki/Build_flags}}
-Nix can guarantee these compiler flags for every dependency because Nix wraps its compiler binaries in scripts, meaning it has more control over them and packages can not remove any flags by clearing environment variables.
-The advantage of overriding the stdenv is that every package in a dependency tree can be optimized, potentially leading to better overall performance of packages.
-But it also has the downside that every package needs to be rebuild potentially taking several days, for example preCICE requires a rebuild of 1500 packages, running for two days on an high-performance computing (HPC) cluster (2x AMD EPYC 7702).
-So it might be sufficient to optimize only the package, that is easily possible with \texttt{(package.override)} such as shown before or additionally optimize its direct dependencies, that is possible but more time-consuming to do, because one has to manually override each dependency.
-
-Spack also allows to override cflags, using \texttt{spack install \ldots cflags='-march=native'}, so the package is then being optimized.
-By setting cflags with \texttt{==}, the cflags are also applied to the dependencies, causing them to be rebuilt.\footnote{\url{https://spack.readthedocs.io/en/latest/basic_usage.html\#compiler-flags}}
-Additionally, there is a way to define flags for a specific registered compiler, but these are only respected by CMake and autotools packages.
-Also, a package might also have unavoidable system dependencies and are not optimized, potentially impacting performance negatively.
-
-EasyBuild applies \texttt{-march=native} by default, but you can further customize your compiler flags using \texttt{\$CFLAGS}, \texttt{\$CXXFLAGS}, \texttt{\$FFLAGS}, \texttt{\$F90FLAGS}.
-It also has the same problem as Spack that it might have unavoidable unoptimized system package that might impact performance negatively.\\
-
-\subsection{Testing}
-
-We also shortly look at the capabilities of the package managers to run test and evaluate the built packages.
-Nix has something that is called a checkPhase that can be run for each package after being built.
-There are sane defaults for all available build systems, such as \texttt{make test} or \texttt{ctest} for makefiles and CMake, \texttt{cargo check} and \texttt{cargo test} for rust packages, \texttt{go test} for go packages and \texttt{pytest} for python packages and more.
-By default, this checkPhase is not enabled for every build system, such as the \texttt{stdenv} build system, that is responsible for make, CMake and similar, does not enable it but the \texttt{go}, \texttt{rust} and \texttt{python} build system does.
-For these build systems, package maintainers have to manually disable the check phase or they can override the check phase if the tests are not working out of the box and make the needed adjustments to resolve the issues.
-
-Spack has something similar, but its not enabled by default, and users, not package maintainers, has to manually specify that they want to run test on installing the package with \texttt{--test=root}.
-
-EasyBuild on the other hand does not offer this feature but has another feature called sanity checks, where EasyBuild ensures that paths and or binaries, specified by a package maintainer inside the package definition, are successfully installed.
-
-Additionally, Nix offers a way for testing multiple packages in combination and or with a specific network setup by providing a way to write your own tests.
-These tests are then running inside isolated qemu virtual machines and users are able to start as many VMs as they need and define the set of commands that need to run on each of the so-called nodes.
-The tests are pythons scripts, and it is possible to wait until systemd targets are running, until ports are available or wait until commands are completed and compare the output and status codes.
-
-\subsection{Conclusion}
-
-To summarize the comparison, while Nix has the biggest software library, it lacks some scientific software compared to Spack and EasyBuild.
-On the other hand, the build sandbox that Nix offers, guarantees a level of reproducibility that Spack and EasyBuild currently cannot provide, because it disallows random linking.
-Spack and EasyBuild on the other side both provide a way better usability to end user with their command line interfaces, Nix still is able to do the same things, such as changing the compiler but generally requires knowledge of the Nix language to make this achievable, but if end user have this knowledge it is possible to change the compile flags for all packages in the dependency graph allowing for additional performance gains, if any package is compiled against the specific hardware.
-
 \section{Conclusion and Outlook}

### 2024-05-10 paper/paper.tex
fix(paper): Use \path for long paths

@@
 Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
-The naming scheme for the preCICE 2.5.0 package as an example is \texttt{/nix/store/0a5gw3l\ldots-precice-2.5.0}.
+The naming scheme for the preCICE 2.5.0 package as an example is \path{/nix/store/0a5gw3l\ldots-precice-2.5.0}.
 This eases checking the Nix store for integrity and ensures that builds using the same inputs are performed only once.
@@
 Another dependency, that is also provided by code\_aster is metis, that partially builds, so we do not replace this dependency, but we need to fix the second part of the build, including the metis programs such as gpmetis, ndmetis, mpmetis and more.
-The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \texttt{/home/karypis/local/lib}.
+The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \path{/home/karypis/local/lib}.
 After removing this line, it also installs, so to resolve this issue in the buildPhase, we unpack the metis tarball, remove this line with \texttt{sed} and then repack the directory.
@@
 
-The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, that works if your version is 3.9.x but not with 3.10.x because then the code calculates \texttt{lib64/python3.1/site-packages} rather than the expected \texttt{lib64/python3.10/site-packages}.
+The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, that works if your version is 3.9.x but not with 3.10.x because then the code calculates \path{lib64/python3.1/site-packages} rather than the expected \path{lib64/python3.10/site-packages}.
 

### 2024-05-10 paper/paper.tex
feat(paper): Write about the not working tutorial cases

@@
 
-Most of the functionality that can be found in the official VM is also available in our NixOS VM and preCICE tutorials, such as the \textit{perpendicular-flap} can be run successfully.
+Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing and \texttt{turek-hron-fsi3} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package.
 

### 2024-05-10 paper/paper.tex
feat: fixup conclusion after section drop and link original report

@@
 With Nix' dependency management, we can easily copy the package definitions we just wrote onto different systems, e.g., entirely different Linux distributions, and execute the build of the closures there.
-Building the same closures yields the same outputs, this will also be described in section \ref{sec:nix-on-hpc}.
 
@@
 The work of the xSDK people and their policy might be a step in the right direction but looking at the current list of software that is available within the latest xSDK release, 26 libraries, indicates that it is an uphill battle that might take several more years, especially if we consider that the first release happened in 2016.
-We have also seen, that the current HPC solutions, Spack and EasyBuild make it somewhat possible to reproducibly build software but not to the degree that Nix allows it.
-That is because they focus on flexibility of installing packages.
-Spacks way of allowing users to change the compiler of a package, easily enable additional features on installation, override cflags and more advanced features, for example regarding mpi, shows that the Nix tooling needs more work or a layer on top that strips away the uniqueness of the DSL for something more user- and beginner-friendly.
 
-Lastly we looked at Nix on HPC and have seen that it is possible to get it running, and we are able to run simulations on one node, so we currently assume that with the correct networking setup it should be possible to run a simulation across multiple nodes.
-It is currently not ideally solved on getting Nix running on HPC as a single user installation, so this can be improved in the future.
-
-The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/}.
-Also the \texttt{setup.sh} and \texttt{clean.sh} scripts used to install Nix on HPC can be found there.\\
+The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more indetail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
 

### 2024-05-10 paper/paper.tex
fix: one page overflow

@@
 This is not allowed in Nix' sandboxed builds.
-Therefore, we need to patch this \texttt{pip} command with the options \texttt{--no-build-isolation --no-cache-dir --no-index --no-deps} and we need to remove the \texttt{--upgrade} option.\\
+Therefore, we need to patch the \texttt{pip} command with the options \texttt{--no-cache-dir --no-index --no-deps --no-build-isolation} and remove the \texttt{--upgrade} option.\\
 

### 2024-05-10 paper/paper.tex
feat(paper): Re-add some HPC

@@
 Nix allows functional descriptions of dependencies up to fixed versions, thus avoiding the issue described above.
-Similar ideas are followed by EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
+Similar ideas are followed by the popular high performance computing (HPC) package managers EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
 All have in common that they rely on scientific software following best practices concerning building and packaging.
@@
 
-\subsection{Nix}
+\subsection{Nix}\label{sec:nix}
 
@@
 With Nix' dependency management, we can easily copy the package definitions we just wrote onto different systems, e.g., entirely different Linux distributions, and execute the build of the closures there.
+Building the same closures yields the same outputs, as already described in section \ref{sec:nix}.
 
@@
 
-The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more indetail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
+The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
 

### 2024-05-10 paper/paper.tex
format(paper): Add sloppybar

@@
 
+\begin{sloppypar}
 Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing and \texttt{turek-hron-fsi3} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package.
+\end{sloppypar}
 

### 2024-05-10 paper/paper.tex
reformat(paper): Reformat sentence so it looks better

--- 2024-05-10 paper/paper.tex
@@
 
-Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing and \texttt{turek-hron-fsi3} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package.
+\begin{sloppypar}
+Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{turek-hron-fsi2} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package, \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing.
+\end{sloppypar}
 

### 2024-05-13 paper/paper.tex
fix: make nix example smaller
Smooth section 2.1

@@
 a software which describes software environments with the help of text files.
-Those text files are made up of imperative commands which are run inside of containers, one layer at a time.
+These text files are made up of imperative commands, which are run inside of containers, one layer at a time.
 The result is several different layers combined into a single output image, which can be instantiated into a running container.
@@
 
-In our case study, we use Nix for building and packaging preCICE adapters and their respective solvers.
 This chapter gives a short introduction into Nix and preCICE.
@@
 
-Nix is a purely functional package manager that provides some features to build software derivations reproducibly.
-It does so, by recursively calculating a hash over all the inputs of a derivation and its dependencies to ensure the completeness of the whole derivation.
-If any of the inputs changes, all dependent derivations will have to be rebuilt.
-
-As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
-Inputs to those functions are parameters such as the package version or the location of the source code.
-Evaluating and realizing those functions with the same inputs, yields the same outputs.
-This is a critical factor where Nix's reproducibility comes from.
-Outputs, i.e. the build artifacts, never change after being built once.
-Nix builds are sandboxed, meaning that there is no internet access possible during a build.
-
-All contents, including source files and resulting build artifacts, are stored inside the Nix store.
-Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
-The naming scheme for the preCICE 2.5.0 package as an example is \path{/nix/store/0a5gw3l\ldots-precice-2.5.0}.
-This eases checking the Nix store for integrity and ensures that builds using the same inputs are performed only once.
-
-As each build output references their complete dependency graphs inside the Nix store, they do not interfere with other build outputs, enabling users to have multiple versions of the same software installed at the same time without raising conflicts.
-An example of a graphical representation of a dependency graph can be seen in Fig. \ref{fig:nix-graph}.
+Nix is a purely functional package manager, which provides features to build software derivations reproducibly.
+It does so, by recursively calculating a hash over all inputs of a derivation and its dependencies to ensure the completeness of the whole derivation.
+If any of the inputs changes, all dependent derivations have to be rebuilt.
+Fig.~\ref{fig:nix-graph} shows the dependency graph of the coreutils package as an example.
 
@@
 
-Software patches are quite easy to apply in Nix.
-One can simply provide a list of \texttt{.patch} files as inputs, Nix will then include the patch files in the build and the calculation of the hash.
-Users can override all inputs freely, so every package can be flexibly adjusted to the users needs.
+As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
+Inputs of these functions are parameters such as the package version or the location of the source code.
+Evaluating these functions with the same inputs yields the same outputs.
+This is a critical factor for Nix's reproducibility.
+Outputs, i.e. the build artifacts, never change after being built once.
+Nix builds are, moreover, sandboxed, meaning that there is no internet access possible during a build.
+
+All contents, including source files and resulting build artifacts, are stored inside the Nix store.
+Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
+The naming scheme for packages includes the above mentioned hash, for preCICE v2.5.0, for example, 
+\path{/nix/store/0a5gw3l...-precice-2.5.0}.
+This eases checking the Nix store for integrity and ensures that builds using the same inputs are performed only once.
+As all build outputs reference their complete dependency graphs inside the Nix store, they do not interfere with other build outputs, enabling installation of multiple versions of the same software without conflicts.
+
+Software patches are easy to apply in Nix.
+One can simply provide a list of \texttt{.patch} files as inputs, Nix then includes the patch files in the build and in the calculation of the hash.
+Users can override all inputs, so that every package can be flexibly adjusted.
 
@@
     \begin{minted}{nix}
-{ lib, stdenv, fetchFromGitHub, cmake,
-  precice, dealii, enable3d ? false }:
-
+{ lib, stdenv, fetchFromGitHub, cmake, precice,
+  dealii, enable3d ? false }:
 stdenv.mkDerivation rec {
@@
   src = fetchFromGitHub {
-    owner = "precice";
-    repo = "dealii-adapter";
-    rev = "dbb25...8367c";
-    sha256 = "sha256-pPQ2...2jlflgUE=";
+    owner = "precice"; repo = "dealii-adapter";
+    rev = "dbb25...8367c"; sha256 = "sha256-pPQ2...2jlflgUE=";
   };
@@
 
-  cmakeFlags = lib.optionals enable3d [
-    "-DDIM=3"
-  ];
+  cmakeFlags = lib.optionals enable3d [ "-DDIM=3" ];
 
   installPhase = ''
-    mkdir -p $out/bin
-    cp elasticity $out/bin/elasticity
+    mkdir -p $out/bin && cp elasticity $out/bin/elasticity
   '';
@@
     \label{lst:dealii-adapter-nix}
-    \hrulefill
-    \vspace*{4pt}
 \end{figure*}

### 2024-05-13 paper/paper.tex
Rewrite preCICE section 2.2 and intro of section 3

@@
 
-preCICE (Precise Code Interaction Coupling Environment) is an open-source software library designed to facilitate the coupling of different computational simulation softwares.
-It provides an interface that allows different simulation tools to exchange data and work together in a collaborative manner, enabling multi-physics as well as multi-scale simulations.
-
-Many scientific simulations require complex combinations of multiple solvers, each specialized in a particular aspect of the problem.
+preCICE (Precise Code Interaction Coupling Environment) is an open-source software library designed to facilitate coupling of different simulation software packages.
+It provides an API, which allows different simulation tools to exchange data and work together in a collaborative manner, enabling multi-physics as well as multi-scale simulations.
+In fact, many scientific simulations require complex combinations of multiple solvers, each specialized in a particular aspect of the problem.
 For example, in fluid-structure interaction problems, where fluid flow interacts with deformable structures, separate solvers can be used to model the fluid dynamics and structural mechanics.
-preCICE acts as a bridge between all solvers participating in the simulation, allowing them to exchange data and synchronize their computations.
-
-preCICE offers a flexible and generic framework for coupling simulations.
-It supports a wide range of solvers, including finite element, finite volume, and particle-based methods.
-The library provides methods for interpolating data between meshes, managing communication between solvers, and handling time stepping and convergence.
-It can handle different types of coupling scenarios, such as one-way or bidirectional couplings, loose or tight couplings, and static or dynamic meshes.
-The core principle of preCICE's functionality is shown visually in Fig. \ref{fig:precice}.
 
@@
     \includegraphics[width=0.5\textwidth]{figures/precice.png}
-    \caption{A visual representation on how preCICE works in principle, see: \url{https://precice.org}}
+    \caption{Software setup of preCICE from the preCICE website \url{https://precice.org}}
     \label{fig:precice}
@@
 
-One main idea behind preCICE is to enable the combination of existing simulation tools without requiring significant modifications to their original source code.
-Instead, preCICE provides an abstraction layer that sits between the individual solvers, allowing them to communicate and exchange data efficiently.
-This approach promotes code reusability and reduces the effort required for coupling simulations.
+preCICE supports a wide range of existing (Legacy) solvers.
+For each solver, there is typically a specific adapter: an either independent software package or simply a source code patch that integrates the preCICE API into the solver, see Fig.~\ref{fig:precice}.
+The preCICE library itself is implemented in C++, bindings to other languages are either supported through native bindings (C, Fortran) or through independent software packages (Python, Julia, Matlab, Fortran).
+Many official adapters, all bindings, the library itself, tutorials, additional tools, and the website including user documentation are collected and released in so-called preCICE distributions\footnote{\url{https://precice.org/installation-distribution.html}}.
+In our case study, we target the distribution v2211.0~\cite{preciceDistribution}.
+To execute simulations, such as the included tutorials, requires the actual solvers next to the distribution.
+These are included, for example, in the preCICE VM\footnote{\url{https://precice.org/installation-vm.html}} and also part of our case study.
+Table \ref{table:label-distribution} lists all software packages of the case study.
+We try to reproducibly build all of these packages with Nix and try to run all tutorial cases.
+The diverse languages and diverse solver packages using various build system makes the preCICE ecosystem a challenging testcase, which is, however, representative for current research software in our experience.
 
-preCICE is an open-source project, that means its source code is freely available and can be modified and distributed by the community.
-This fosters collaboration and encourages the development of new features and improvements by researchers and engineers worldwide.
-
-Overall, preCICE simplifies the process of coupling different simulation programs, enabling the simulation of complex multi-physics problems.
-It promotes interoperability, code reusability, and collaboration, making it a valuable tool in various engineering and scientific domains.
+\begin{table}[!t]
+  \normalsize
+  \caption{Software packages of the case study}
+  \label{table:label-distribution}
+  \centering
+  \begin{tabular}{|l|l|l|l|}
+    \hline
+    \bfseries Package & \bfseries Type & \bfseries Build system & \bfseries Version \\ \hline
+    precice & library & cmake & 2.5.0 \\ \hline
+    fortran-module & bindings & make & 9e3f405 \\ \hline
+    julia & bindings & julia & 2.5.0 \\ \hline
+    matlab & bindings & matlab build script & 2.5.0.0 \\ \hline
+    pyprecice & bindings & setup.py & 2.5.0.1 \\ \hline
+    config-visualizer & tooling & setup.py & 60f2165 \\ \hline
+    precice-aste & tooling & cmake & 3.0.0 \\ \hline
+    dealii & solver & cmake & 9.4.1 \\ \hline
+    precice-dealii-adapter & adapter & cmake & dbb25bea \\ \hline
+    precice-calculix-adapter & solver + adapter & makefile & 2.20 \\ \hline
+    fenics & solver library& cmake + setup.py & 2019.1.0 \\ \hline
+    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\ \hline
+    precice-dune & adapter & scripts + cmake & 2.8.0 \\ \hline
+    dune-fem & solver library & setup.py & 2.8.0.0 \\ \hline
+    code-aster & solver & custom build system in python & 14.6.0 \\ \hline
+    openfoam & solver & wmake/Allwmake & 2206 \\ \hline
+    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
+    precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
+    nutils & solver library & setup.py & 7.0.0 \\ \hline
+  \end{tabular}
+\end{table}
 
@@
 
-To see how well Nix can be used to build and package scientific software reliably and reproducibly, we conduct a case study in the context of the latest preCICE distribution~\cite{preciceDistribution}, currently preCICE \texttt{v2211.0}.
-The preCICE distribution consists out of the preCICE core library, as well as some tools, bindings, adapters, tutorials, the website and documentation, and the preCICE VM\footnote{\url{https://precice.org/installation-distribution.html}}.
-Table \ref{table:label-distribution} shows an overview over all packages provided by the preCICE distribution and their respected versions.
-
-\begin{figure*}[!t]
-  \normalsize
-  \caption{Overview of preCICE distribution}
-  \label{table:label-distribution}
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    \bfseries Package & \bfseries Type & \bfseries Build System & \bfseries Version \\ \hline
-    precice & library & cmake & 2.5.0 \\ \hline
-    fortran-module & bindings & make & 9e3f405 \\ \hline
-    julia & bindings & julia & v2.5.0 \\ \hline
-    matlab & bindings & matlab build script & v2.5.0.0 \\ \hline
-    pyprecice & bindings & setup.py & 2.5.0.1 \\ \hline
-    config-visualizer & tooling & setup.py & 60f2165 \\ \hline
-    precice-aste & adapter & cmake & 3.0.0 \\ \hline
-    dealii & solver & cmake & 9.4.1 \\ \hline
-    precice-dealii-adapter & adapter & cmake & dbb25bea \\ \hline
-    precice-calculix-adapter & adapter & makefile & 2.20 \\ \hline
-    fenics & solver & cmake + setup.py & 2019.1.0 \\ \hline
-    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\ \hline
-    precice-dune & adapter & scripts + cmake & 2.8.0 \\ \hline
-    Dune-fem & library & setup.py & 2.8.0.0 \\ \hline
-    code-aster & solver & custom build system in python & 14.6.0 \\ \hline
-    openfoam & solver & wmake/Allwmake & 2206 \\ \hline
-    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
-    Precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
-    Nutils & library & setup.py & 7.0.0 \\ \hline
-  \end{tabular}
-\end{figure*}
-
-We are mainly interested in the tools, adapters including their solvers and the preCICE VM.
-For each of these entities, we present possible challenges, how to circumvent those and also comment on how software can be designed to make it as easy as possible for other users to build and package the software in question.
+We present challenges, solutions, and workarounds for all software packages of the preCICE distribution.
 

### 2024-05-13 paper/paper.tex
Smooth Section 3 till incl Code_Aster

@@
 \usepackage{subfig}
+
+\usepackage{xcolor}
+%\newcommand\todo[1]{\textcolor{red}{TODO #1}}
+
 \input{frontmatter}
@@
 
-We present challenges, solutions, and workarounds for all software packages of the preCICE distribution.
+We present challenges, solutions, and workarounds for all software packages of the preCICE ecosystem.
+\todo{mention where and which results to find on GitHub}
 
@@
 
-The preCICE distribution comes with a few tools that ease working with preCICE.
-They are mainly meant for debugging and visualizing several aspects of preCICE and preCICE configuration files.
+The preCICE distribution includes tools around preCICE.
 
@@
 
-This package is a collection of tools that help in development of new adapters.
-It stands for Artificial Solver Testing Environment, so it allows for debugging adapter mapping setups and other similar tasks.
+ASTE\footnote{\url{https://github.com/precice/aste}} stands for Artificial Solver Testing Environment.
+It is a thin wrapper around the preCICE API, which allows, for instance, testing of data mapping with real geometries.
+ASTE requires VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
+This feature is needed for ASTE, however.
+As already mentioned, Nix allows overriding of inputs, so in this case, we can look at the package definition of VTK.
+The parameter \texttt{enablePython} can simply be set to \texttt{true}. We then also need to supply the version of python, which we set to python3.
 
-ASTE\footnote{\url{https://github.com/precice/aste}} needs VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, that is built without python support in the nixpkgs repository to reduce compilation time.
-However, this feature is needed for ASTE.
-As already mentioned, Nix allows easy overriding of inputs, so in this case, we can look at the package definition of VTK.
-The parameter \texttt{enablePython} can simply be set to \texttt{true}, then we also need to supply the version of python we want to compile against, so in our case \texttt{python = python3;}.
+\subsubsection{Config visualizer}
 
-After enabling python support in VTK, the package can be built without further modifications.
-
-Only for testing, we additionally need to replace the absolute paths to \texttt{/bin/bash} and \texttt{/usr/bin/time}, then all tests succeed without any problems.
-
-\subsubsection{config-visualizer}
-
-The precice-config-visualizer\footnote{\url{https://github.com/precice/config-visualizer}} package can be used to generate images that represent a specified preCICE config file to check graphically what is configured in the file.
-
-This tool is already packaged upstream, so we can just use it.
-As it is a python package, it is quite easy to build with Nix, so the package definition is straight forward.
+The preCICE config visualizer\footnote{\url{https://github.com/precice/config-visualizer}}, as the name suggests, visualizes preCICE configuration files.
+Similarly to preCICE itself, the tool is already packaged upstream.
+As a python package, it is easy to handle with Nix.
 
@@
 
-\subsubsection{python-bindings}
+\todo{matlab and fortran module missing}
 
-The \texttt{pyprecice} package is already available upstream, however it is broken and does not compile as it is not able to find the python module \texttt{pkgconfig}.
+\subsubsection{Python bindings}
+
+The \texttt{pyprecice} package is already available upstream, but it is currently broken: It does not compile as it is not able to find the python module \texttt{pkgconfig}.
 After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.
+
+\todo{fixed upstream?}
 
@@
 
-\subsection{Adapters}
-
-A preCICE distribution comes bundled with several adapters, that are currently not in the upstream nixpkgs repository.
-Many of the adapters use solvers that are not packaged in the upstream repo as well, so in order to get the adapters to work, these also have to be packaged and sometimes patched.
+\subsection{Solvers and adapters}
 
@@
 
-The preCICE adapter for CSM code CalculiX~\cite{Uekermann2017_Adapters}, requires the original CalculiX source code and provides a new Makefile in the adapter repository.
-By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
+The preCICE adapter for the structural mechanics code CalculiX~\cite{Uekermann2017_Adapters} requires the original CalculiX source code and provides a new makefile in the adapter repository.
+By default, this makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
 It then builds the original code and the adapter code together, resulting in a combined binary.
 
-The Makefile also has hard-coded dependency location for \texttt{SPOOLES}, \texttt{ARPACK} and \texttt{YAML} so these need to be replaced with the equivalent \texttt{pkg-config} calls, because Nix does not have these dependencies at the hard-coded locations, and generally only finds dependencies with \texttt{pkg-config}.
-Additionally, there is no \texttt{install} target provided by the Makefile so the resulting binary needs to be installed manually by copying it to \texttt{\$out}.\\
+The Makefile also has hard-coded locations its dependencies: \texttt{SPOOLES}, \texttt{ARPACK}, and \texttt{YAML}. These need to be replaced with the equivalent \texttt{pkg-config} calls.
+Additionally, there is no \texttt{install} target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.\\
 
-\subsubsection{code\_aster}
+\subsubsection{Code\_Aster}
 
-The code\_aster preCICE adapter~\cite{Uekermann2017_Adapters} is a python file that needs to be placed in the code-aster lib directory.
-For this, the code\_aster solver needs to be packaged at the latest stable version 14.6.
-It uses a custom build system based on a \texttt{setup.py} file, that invokes build and install phases for dependencies and in the end for the solver itself.
+The Code\_Aster-preCICE adapter~\cite{Uekermann2017_Adapters} is a python file, which needs to be placed in the Code\_Aster lib directory.
+For this, the Code\_Aster solver needs to be packaged at version 14.6.
+It uses a custom build system based on a \texttt{setup.py} file, which invokes build and install phases for dependencies and in the end for the solver itself.
 
-The stable package also distributes the dependencies as pinned tarballs.
-On Nix almost none of these dependencies, complete the build stage with the \texttt{setup.py}, so we need to pull them out, package them separately as Nix packages and configured the build system with a \texttt{setup.cfg}.
+The Code\_Aster package distributes its dependencies as pinned tarballs.
+On Nix, almost none of these dependencies complete the build stage. We, thus, need to pull them, package them separately as Nix packages, and configure the build system with a \texttt{setup.cfg}.
 In this file, we are able to disable the installation of the pinned dependencies and provide paths to the dependencies instead.
 
-With this approach, we disable HDF5, as it is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
+More precisely, we disable HDF5, as it is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
 HDF5 is packaged with multiple outputs.
 One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, this is, because most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
-Additionally, code\_aster requires medfile and scotch, that are both available upstream and can be used as a dependency.
+Additionally, Code\_Aster requires Medfile and Scotch, which are both available upstream and can be used as a dependency.
+Another dependency is Metis, which partially builds. Thus, we decide to not replace it, but to fix the remainder of the build.
+The issue is precisely that the \texttt{CMakeLists.txt} for the metis programs hard-codes \texttt{link\_directories} to \path{/home/karypis/local/lib}.
+We unpack the Metis tarball, remove this line with \texttt{sed}, and repack the directory.
 
-Another dependency, that is also provided by code\_aster is metis, that partially builds, so we do not replace this dependency, but we need to fix the second part of the build, including the metis programs such as gpmetis, ndmetis, mpmetis and more.
-The issue is that in the \texttt{CMakeLists.txt} for the metis programs \texttt{link\_directories} is hard-coded to \path{/home/karypis/local/lib}.
-After removing this line, it also installs, so to resolve this issue in the buildPhase, we unpack the metis tarball, remove this line with \texttt{sed} and then repack the directory.
+The remaining two dependencies, Mumps and Tfel are not available upstream and need to be packaged.
+Mumps uses a Makefile and can be configured using Makefile.inc.
+By default, it provides a few configurations for systems such as Ubuntu.
+We use one of these provided configuration files to write our own custom config file.
+It defines the locations for dependencies, such as Scotch, Metis, Parmetis, or Blacs.
+Almost all dependencies are available upstream, but we need to recompile Scotch with additional build flags.
+Moreover, for packaging Blacs, which also uses a Makefile, we configure the build using a custom \texttt{Bmake.inc} file, which is used to set compiler flags, the path to MPI, and the bash installation.
+Tfel, on the other hand, uses CMake as build system making it easy to package it with Nix.
 
-The last 2 dependencies, are not available and need to be packaged, this includes mumps and tfel, also known as mfront.
-
-Mumps builds using a Makefile and can be configured using Makefile.inc, by default it provides a couple of these configurations for systems such as Ubuntu, so we use one of these provided configuration files to write our own custom config file.
-It defines the location for dependencies such as scotch, metis, parmetis, liblapack, blas, scalapack and blacs.
-Almost all the dependencies, are available upstream, but we need to recompile scotch with additional build flags, \texttt{scotch ptscotch esmumps ptesmumps} using Nix overrideAttrs feature.
-For packaging blacs, that also uses a Makefile, we configure the build using a custom \texttt{Bmake.inc} file, that is used to set compiler flags, the path to the mpi and bash installation.
-
-The last missing dependency for code\_aster, tfel, uses CMake as a build system making it trivial to package it with Nix and we only have to set a couple of CMake flags.
-
-After successfully packaging all dependencies, the build phase of code\_aster completes, but the installation part has another issue, that also affects new versions of Ubuntu and other distributions that use a python version greater than 3.9.
-There is a forum post without any resolution from the developers or the community, so we need to manually patch the bug.
-
-The issue is that the custom build system does not correctly calculate the python site-packages directory because it unconditionally slices the first 3 chars from \texttt{sys.version}, that works if your version is 3.9.x but not with 3.10.x because then the code calculates \path{lib64/python3.1/site-packages} rather than the expected \path{lib64/python3.10/site-packages}.
-
-After patching this bug code\_aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib} and we provide symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/} as code\_aster expects configuration files in these directories.
-
-Last but not least, we can copy the code\_aster adapter at the required location.\\
+After successfully packaging all dependencies, the build phase of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
+There is a forum post without resolution, so we need to manually patch the bug. \todo{link post}
+The issue is precisely, that the custom build system does not correctly calculate the python site-packages directory.
+It unconditionally slices the first 3 chars from \texttt{sys.version}, which works for version 3.9.x, but not for version 3.10.x.
+After patching, Code\_Aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib}, providing symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/}.
 

### 2024-05-14 paper/paper.tex
Smooth deal.II till incl OpenFOAM

@@
 The Makefile also has hard-coded locations its dependencies: \texttt{SPOOLES}, \texttt{ARPACK}, and \texttt{YAML}. These need to be replaced with the equivalent \texttt{pkg-config} calls.
-Additionally, there is no \texttt{install} target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.\\
+Additionally, there is no \texttt{install} target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.
 
@@
 
-In order to package this adapter, we first need to package deal.II~\cite{dealII95}, that uses CMake as build system and works without any adjustments.
-The same applies for the adapter, that also uses CMake and needs preCICE as a dependency.
-When supplied with the inputs, the adapter builds without any problems.
-As an example for the Nix language, you can see the abbreviated Nix code for the dealii.II adapter in Fig. \ref{lst:dealii-adapter-nix}.\\
+To package the deal.II-preCICE adapter, we first need to package deal.II~\cite{dealII95} itself, which uses CMake as build system and works without any adjustments.
+The same then applies for the adapter, which also uses CMake.
+As an example for the Nix language, we provide the Nix code for the deal.II-preCICE adapter in Fig.~\ref{lst:dealii-adapter-nix}.
 
@@
     \end{minted}
-    \caption{A sample of Nix code used to package the dealii-adapter, see: \url{https://github.com/precice/dealii-adapter/}}
+    \caption{Nix code to package the deal.II-preCICE adapter}
     \label{lst:dealii-adapter-nix}
@@
 
-The DUNE~\cite{bastian2020dune} package uses a combination of CMake files and custom build scripts making the build process quite tedious.
-This might be an issue because of the ongoing migration to CMake from autotools and might be resolved in the future.
-
-It is also a minor issue that the dune-project is a collection of repositories rather than a monorepository, because users have to correctly clone all repositories so the build system finds all relevant information.
-We clone all of the required dune repos into a directory and additionally clone the adapter into the same directory.
-
-For the build and install, process we need to manually set \texttt{\$DUNE\_CONTROL\_PATH} and \texttt{\$DUNE\_PY\_DIR} environment variables, both variables also need to be set at runtime.
-Additionally, we need to patch the python install process because the current CMake file that handles this install tries to access the internet with \texttt{pip install}.
-This is not allowed in Nix' sandboxed builds.
-Therefore, we need to patch the \texttt{pip} command with the options \texttt{--no-cache-dir --no-index --no-deps --no-build-isolation} and remove the \texttt{--upgrade} option.\\
+DUNE~\cite{bastian2020dune} uses a combination of CMake files and custom build scripts making the build process in Nix tedious.
+There is an ongoing migration from autotools to CMake for DUNE, which might improve the situation in the future.
+Another minor issue is that the DUNE project is rather a collection of repositories than a monorepo.
+This means, users have to correctly clone all repositories such that the build system finds all relevant information.
+We clone all of the required DUNE repositories into a directory and additionally clone the DUNE-preCICE adapter into the same directory.
+For the build and install process, we need to manually set the \texttt{\$DUNE\_CONTROL\_PATH} and the \texttt{\$DUNE\_PY\_DIR} environment variables. Both variables also need to be set at runtime.
+Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix' sandboxed builds.
 
@@
 
-We are explicitly referring to the legacy version of FEniCS~\cite{fenics} here, not the newer and still experimental FEniCSx.
-This adapter~\cite{Rodenberg2021} uses the open source computing platform FEniCS, that is already packaged within nixpkgs, to provide preCICE support.
+FEniCS~\cite{fenics} is already packaged upstream, however, not will all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support.
+This is why, we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
+The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code.
+We could not enable tests for PETSc4py, however, because they depend on OpenMPI and the network, which is not available within Nix' build sandbox.
+Afterwards, we specify PETSc4py as dependency for FEniCS and enable the PETSc support.
 
-Although it is available, not all features the adapter needs are enabled, including PETSc support.
-To fix this, we need to package PETSc4py, that provides python bindings for PETSc and is not present in the upstream nixpkgs repository.
-The build process is possible without any issues, because it uses the internal \texttt{buildPythonApplication} build tool, but we need to add the additional build optional \texttt{build\_src --force} because the cython code needs to be rebuilt.
-We also are not able to enable tests for this package because tests depend on OpenMPI and the network, that is not available within the Nix build sandbox.
-
-After this, we can specify PETSc4py as dependency for FEniCS and can enable support for this feature.
-
-Building the adapter with Nix is now possible with \texttt{buildPythonPackage} and does not require any special workarounds.
-A majority of the \texttt{pytest} ran successfully, but two tests failed inside the Nix sandbox, that means all tests needed to be disabled.\\
+Building the FEniCS-preCICE adapter is now directly possible with \texttt{buildPythonPackage}.
+All but two of the adapter's tests run successfully, which means, however, that we need to disable the tests.
 
@@
 
-There are two flavors of OpenFOAM, we are referring to the OpenFOAM fork of OpenCFD Ltd\footnote{\url{https://www.openfoam.com/}}.
+The preCICE-OpenFOAM adapter~\cite{OpenFOAMpreCICE} supports multiple flavors of OpenFOAM. We restrict our analysis to the OpenFOAM fork of OpenCFD Ltd\footnote{\url{https://www.openfoam.com/}}.
 
-OpenFOAM uses its own custom build system called \texttt{wmake} that is sometimes called with a wrapper script called \texttt{Allwmake}.
-The build system sets 36 environment variables that are set by wmake, one step at a time by checking several parameters, e.g.
-the cpu architecture or the location of the source code, and concatenating some other environment variables together.
+OpenFOAM uses its own custom build system called \texttt{wmake}, which is typically called with a \texttt{Allwmake} wrapper script.
+The build system sets 36 environment variables, one step at a time by checking several parameters, e.g.
+the CPU architecture or the location of the source code.
+Also during runtime, these variables need to be set, either by a wrapper script or by manually sourcing a file inside the installation directory of OpenFOAM.
+For Nix, these properties are unfortunate. To compile OpenFOAM with Nix, we need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
+We also use a shell script called \texttt{set-openfoam-vars}, which exports the environment variables to the current shell session.
+For simplicity, we hard-code all parameters, such as the processor architecture.
+These could be parametrized, however, based on the Nix inputs to allow for optimized builds.
+We use the script during the build phase to source all variables such that, for example, \texttt{OPENFOAM\_SRC\_PATH} points to the default location \texttt{/build/openfoam}.
+Afterwards, \texttt{./Allwmake -j -q} is sufficient to start the build.
+The install phase then copies the necessary files and directories to \texttt{\$out}, replaces the mock \texttt{OPENFOAM\_SRC\_PATH} by the value of \texttt{\$out}, and creates a wrapper for the \texttt{openfoam} shell script.
 
-Allwmake and wmake rely on these environment variables to be set, as otherwise the tools will exit with a warning.
-During runtime, before running any OpenFOAM commands, a small wrapper script has to be started that sets all relevant environment variables, or a user has to manually source a file inside the installation directory of OpenFOAM.
-OpenFOAM uses modules for extensibility and flexibility that need the source code of OpenFOAM and also require wmake as a build tool.
-
-For Nix, these properties are rather unfortunate.
-To compile OpenFOAM with Nix, we first need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
-Secondly, we render a shell script we call \texttt{set-openfoam-vars} that exports all necessary environment variables to the current shell session.
-At the time of writing, all parameters such as the processor architecture are hard-coded for simplicity reasons but could be parametrized based on the Nix inputs to allow for optimized builds.
-During the buildPhase, the rendered script will be sourced to set all variables, e.g.
-so the \texttt{OPENFOAM\_SRC\_PATH} points to \texttt{/build/openfoam} being the default location for Nix builds inside the sandbox.
-After sourcing the script, the single command \texttt{./Allwmake -j -q} is sufficient to start the build.
-
-The installPhase then simply copies the necessary files and directories to \texttt{\$out}, replaces the mock \texttt{OPENFOAM\_SRC\_PATH} by the value of \texttt{\$out} and creates a wrapper for the \texttt{openfoam} shell script.
-
-As the preCICE OpenFOAM-adapter~\cite{OpenFOAMpreCICE} is an OpenFOAM module, the derivation of the adapter needs OpenFOAM and wmake as build inputs.
-The correct environment variables have to be set again, but this is rather easy now, as we have the \texttt{set-openfoam-vars} script that we can simply source during the buildPhase.
-The last thing we need to change before launching the build with Allwmake is to set the target directory for the adapter.
-A single shared library that can be copied to \texttt{\$out/lib/} is the only output for this build.
-
-Debug and timing modes for the adapter can simply be enabled by setting the \texttt{debugMode} and \texttt{enableTimings} inputs to the derivation to true.
-
-OpenFOAM is available on Spack and EasyBuild that both patch the OpenFOAM source files as those tools also have to set environment variables to make a build compatible with wmake.\\
+The preCICE-OpenFOAM adapter is an OpenFOAM function object, requiring OpenFOAM and wmake as build inputs.
+The environment variables have to be set again through the \texttt{set-openfoam-vars} script.
+Afterwards, we set the target directory for the adapter to \texttt{\$out/lib/} as the adapter is a shared library. Building then works successfully.
 

### 2024-05-14 paper/paper.tex
Tweak final sentence in intro (remove "3 systems")

@@
 preCICE is a coupling library for partitioned multiphysics simulations.
-The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
-We try to build the complete ecosystem using Nix and verify portability on three different systems.
+The ecosystem includes diverse legacy solvers, adapters, language bindings, and testcases besides the coupling library itself making it a challenging and representative testcase.
+We try to build the complete ecosystem using Nix and run all testcases.
 

### 2024-05-14 paper/paper.tex
Smooth SU2 section

@@
 
-This package is a preCICE adapter~\cite{Uekermann2017_Adapters} for the CFD code SU2 and extends the original SU2 source code by patching the files.
-For this, the adapter provides a custom script that needs to be executed before the build phase.
-Nix offers the patchPhase that allows us to run the script in this phase and patch the original source, after that the \texttt{stdenv} automatically recognizes that the repository uses autotools for building the software and runs the configurePhase.
-
-We additionally needed to add the two configureFlags \texttt{--with-include} and \texttt{--with-lib} to the preCICE install directory because otherwise autoconf does not find the preCICE installation directory.
+The SU2-preCICE adapter~\cite{Uekermann2017_Adapters} patches and extends the original CFD code SU2 \todo{cite SU2}.
+To this end, the adapter provides a script to run before the build phase.
+We do this in Nix' patch phase. Afterwards, \texttt{stdenv} automatically recognizes that SU2 uses autotools for building and runs the configure phase.
+Finally, to find the preCICE installation, we need to set the configuration flags \texttt{--with-include} and \texttt{--with-lib}.
 

### 2024-05-14 paper/paper.tex
Add funding acknowledgment
Rewrite discussion

@@
 
-After detailing the challenges of building and packaging each solver and their respective adapter, we want to discuss some common points between these scientific software packages.
-preCICE, as a multi-component research software, covers many different simulation codes and therefore provides a good overview over different kinds of packages, making this case study representative for multi-component research software.\\
+After detailing the challenges of building and packaging all components of the preCICE ecosystem and running the tutorials as testcases, we want to discuss common points between these scientific software packages.
+The required effort to package the individual components varies drastically. Major problems can be traced back to the build system.
+Custom build systems require an disproportional amount of additional work to make them runnable on any system.
+OpenFOAM and Allwmake are prime examples of this issue.
+Package managers have to understand a whole new build system for a single package -- a fact that is often ignored when considering the trade-off between maintaining a custom build system versus pivoting to an industry standard, such as CMake or autotools.
+Such standard build tools generally provide interfaces for dependency management and optional features.
+This is often better than only documenting these, since documentation needs to be kept in sync with the underlying code.
 
-Generally speaking, there are major gaps regarding the ability to successfully package the software between each different software solution that can be traced back to the build system.
-Custom build systems require an disproportional amount of additional work to make them runnable on any system.
-OpenFOAM and Allwmake are prime examples of these issues because now package managers have to understand a whole new build system for exactly one package in order to successfully package it and make it easily available to users.
-It is also worth mentioning that writing your own custom build system, requires additional time to maintain this build system, that could have been spent working on the software itself.
-Packages that use common, industry proven package managers, such as CMake and autotools make it easier for developers of the software to build their software as well as package managers to package it.
-These build tools, generally also provide interfaces for dependency management that can be used to tell users and package maintainers what dependencies are needed, and it is usually better than only documenting required dependencies, because documentation needs to be kept in sync with the underlying code.
-They also generally already provide proven interfaces for configuring packages with additional features, and these also might be self documenting.
+The authors of xSDK\footnote{\url{https://xsdk.info/}} come to a similar conclusion.
+They provide a list of package policies~\cite{xSDK2023}, which, for example, specify that packages must have an appropriate build system, which includes CMake and autoconf as examples.
+If a package supports all policies, they can be added to the growing list of packages in xSDK.
+It is no surprise that xSDK packages (e.g., PETSc, deal.II, or preCICE) are also easy to package with Nix.
 
-The authors of the xSDK\footnote{\url{https://xsdk.info/}} came to a very similar conclusion.
-Their goal is to ``identify, adapt, and adopt best practices in software engineering'' and ``to provide the foundation of this extensible scientific software ecosystem''~\footnote{\url{https://xsdk.info/}}.
-To achieve this they provide a list of package policies~\cite{xSDK2023}, that for example specifies as mandatory policy that packages ``must support portable installation through Spack''~\cite{xSDK2023} and that all packages ``should have a build system that is appropriate for the language''~\cite{xSDK2023}, that includes CMake and autoconf as examples.
-If a package supports all these policies, they can be added to the growing list of packages in the xSDK, that currently among other things include PETSc, deal.ii and preCICE.
-These packages are already part of nixpkgs, or it is easy to provide a package for them because they support portable installation, meaning installation into custom prefixes such as nix store paths, and common build systems such as CMake and autoconf.
+The xSDK policies also requires portable installations.
+Several components in the preCICE ecosystem require installation into \texttt{\$HOME} or any required files to be located at specific locations.
+Both, the CalculiX-preCICE adapter and the SU2-preCICE adapter require the original source code to be present as they add new features on top or patch the source code.
+The current solution of cloning the code at the specific location works, but it might be better to look into Git submodules.
+This would moreover allow pinning the version of the original code compared to only documenting it.
 
-Another finding is, that is also addressed in xSDK through portable installations, is the requirement to install software to be installed into \texttt{\$HOME} or requires files to be located at a specific location on your file system.
-Most of the time, it is possible to work around this, but there are already available solutions for dependency management.
-Both, the CalculiX-adapter and SU2-adapter do require the original source code to be present as they add new features on top or patch the source code.
-The current solution of the user cloning the code at the specific location works, but it might be better to look into submodules because that also allows the developers to pin the version more easily rather than documenting it, as that could more easily lead to user error, when they try to package and install the adapters.
-
-Additionally, hard-coding libraries to \texttt{/usr/lib} and \texttt{/usr/include} might work on one distribution, but might make it hard or even impossible to install a package on a different distribution or even a later version of the same distribution.
-Solutions such as \texttt{pkg-config} that can be configured, so it finds libraries outside the common paths, should always be preferred and work out of the box with the common paths such as \texttt{/usr/lib}.
-
-With Nix' dependency management, we can easily copy the package definitions we just wrote onto different systems, e.g., entirely different Linux distributions, and execute the build of the closures there.
-Building the same closures yields the same outputs, as already described in section \ref{sec:nix}.
+Finally, hard-coding libraries to \texttt{/usr/lib} and \texttt{/usr/include} might work on one system, but might make it hard or even impossible to install a package on a different system.
+Configurable solutions, such as \texttt{pkg-config}, should always be preferred and work out of the box with common paths such as \texttt{/usr/lib}.
 
@@
 
+\acknowledge{
+We thankfully acknowledge the funding by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy EXC 2075 -- 390740016 and the support by the Stuttgart Center for Simulation Science (SimTech).
+}
+
 \nocite{*}

### 2024-05-14 paper/paper.tex
fix(paper): Trailing whitespace
feat(paper): Resolve some todos

@@
 We present challenges, solutions, and workarounds for all software packages of the preCICE ecosystem.
-\todo{mention where and which results to find on GitHub}
+The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
 
@@
 
-\todo{matlab and fortran module missing}
+\subsubsection{Matlab bindings}
+
+There were several attempts to package matlab\footnote{\url{https://de.mathworks.com/products/matlab.html}} for NixOS in the past\footnote{\url{https://github.com/NixOS/nixpkgs/issues/56887}}, yet there was no success so far.
+This might be due to the fact, that matlab needs to be installed by running an installation wizard that downloads files during the installation process and verifies the licence.
+Also, as matlab is not installed in the VM, we decided to not proceed in packaging the software.
+
+\subsubsection{Fortran bindings}
+
+preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the fortran programming language.
+We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
+The resulting binary could run the example that comes with the module.
 
@@
 
-The \texttt{pyprecice} package is already available upstream, but it is currently broken: It does not compile as it is not able to find the python module \texttt{pkgconfig}.
+The \texttt{pyprecice} package is already available upstream, but did not compile as it is not able to find the python module \texttt{pkgconfig}.
 After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.
+This solution was contributed to the upstream nixpkgs repository.
 
-\todo{fixed upstream?}
-
-\subsubsection{Julia-bindings}
+\subsubsection{Julia bindings}
 
@@
 After successfully packaging all dependencies, the build phase of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
-There is a forum post without resolution, so we need to manually patch the bug. \todo{link post}
+There is a forum post\footnote{\url{https://forum.code-aster.org/public/d/26475-problem-installing-code-aster-version-14-6/11}} without resolution, so we need to manually patch the bug.
 The issue is precisely, that the custom build system does not correctly calculate the python site-packages directory.
@@
 
-The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
-
 While this case study and this comparison has provided valuable insights into the topic of reproducible software, there are still numerous unanswered questions that merit further investigation.

### 2024-05-15 paper/paper.tex
feat(paper): Add su2 citation
feat: add 3 sentences to intro about spack, eb, nix comparison

@@
 Similar ideas are followed by the popular high performance computing (HPC) package managers EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
-All have in common that they rely on scientific software following best practices concerning building and packaging.
+At FOSDEM 2018, Spack, EasyBuild and Nix were compared with each other~\footnote{\url{https://archive.fosdem.org/2018/schedule/event/installing\_software\_for\_scientists/}} by the EasyBuild release manager.
+While looking into Spack and EasyBuild we found some deficits regarding reproducibility, that they link against core system libraries, that can break independently if those libraries get updated, but the Spack developers are aware of these shortcomings and are already working on improvements~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
+EasyBuild, on the other hand, even has an option for linking against system libraries called \texttt{osdependencies}~\footnote{\url{https://docs.easybuild.io/writing-easyconfig-files/\#dependency\_specs}}, so packages need to be vetted and potentially updated prior to usage that this option isn't used in any package of the dependency graph if the best possible reproducibility is desired.
+All three solutions have in common that they rely on scientific software following best practices concerning building and packaging.
 Unfortunately, most legacy software projects do not do this.
@@
 
-The SU2-preCICE adapter~\cite{Uekermann2017_Adapters} patches and extends the original CFD code SU2 \todo{cite SU2}.
+The SU2-preCICE adapter~\cite{Uekermann2017_Adapters} patches and extends the original CFD code SU2~\cite{su2}.
 To this end, the adapter provides a script to run before the build phase.

### 2024-05-15 paper/paper.tex
feat(paper): Add tutorials subsection

@@
 preCICE is a coupling library for partitioned multiphysics simulations.
-The ecosystem includes diverse legacy solvers, adapters, language bindings, and testcases besides the coupling library itself making it a challenging and representative testcase.
-We try to build the complete ecosystem using Nix and run all testcases.
+The ecosystem includes diverse legacy solvers, adapters, language bindings, and tutorials besides the coupling library itself making it a challenging and representative testcase.
+We try to build the complete ecosystem using Nix and run all tutorials.
 
@@
 
-preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the fortran programming language.
+preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the Fortran programming language.
 We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
@@
 
-\begin{sloppypar}
-Most of the functionality that can be found in the official VM is also available in our NixOS VM and all preCICE tutorials could be run successfully, except \texttt{turek-hron-fsi2} as the tutorial needs \texttt{swak4foam}, an add-on to OpenFOAM, we were not able to package, \texttt{flow-over-heated-plate-steady-state} as the manual steps involved were unclear, \texttt{partitioned-heat-conduction-complex} and \texttt{partitioned-heat-conduction-direct} as the \texttt{mshr} python dependency was missing.
-\end{sloppypar}
+\subsection{preCICE tutorials}
+
+The preCICE tutorials included in the distribution can not only be used as reference examples on how to couple different adapters, but also to verify functionality of adapter and solver installations by executing them.
+
+All v202211.0 test can be run successfully inside our NixOS VM and also in a \texttt{nix develop} environment, except \texttt{turek-hron-fsi2} and \texttt{flow-over-heated-plate-steady-state}.
+
+We could not manage to build \texttt{swak4foam}, an add-on to OpenFOAM that is needed in the \texttt{turek-hron-fsi2} tutorial, from source.
+The official preCICE VM uses a prebuilt version of swak4foam, however also patching the prebuilt binary did not work.
+
+The second simulation we were not able to run is \texttt{flow-over-heated-plate-steady-state}.
+This testcase needs code-aster to work, that we managed to package as already mentioned.
+However, when running the simulation after executing the manual steps as written in the documentation, there occurs a segmentation fault, crashing code-aster when the python code tries to call a Fortran subroutine.
 

### 2024-05-16 paper/paper.tex
Smooth VM and tutorials section

@@
 
-\subsubsection{su2-adapter}
+\subsubsection{SU2}
 
@@
 
-The current preCICE VM is built on top of Vagrant\footnote{\url{https://www.vagrantup.com/}} using Ubuntu as a base image.
-When it is run for the first time, it further installs software, compiles programs and the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
-This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the ``latest'' version, e.g. the \texttt{main} branch of the preCICE repository.
+The current preCICE VM\footnote{\url{https://precice.org/installation-vm.html}} is built on top of Vagrant\footnote{\url{https://www.vagrantup.com/}} using Ubuntu as a base image.
+When it is run for the first time, it further installs software, compiles programs, and the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
+This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the \textit{latest} version, e.g.
+the \texttt{main} branch of the preCICE repository.
 
 Nix comes with the built-in functionality of producing qemu\footnote{\url{https://www.qemu.org/}} VM images.
-We used Nix to define a VM image with NixOS as a base, that can be built reproducibly.
-The image contains all the adapters and solvers from the case study and some additional custom tools that can only be seen in the official VM, such as the \texttt{preciceToPNG} command.
+We use Nix to define a VM image with NixOS as a base, which can be built reproducibly.
+The image contains all components of the case study and some additional custom tools that can only be seen in the official VM, such as the \texttt{preciceToPNG} command.
 Also, we generate an iso and a Vagrant VirtualBox file of the VM as additional outputs.
@@
 
-The preCICE tutorials included in the distribution can not only be used as reference examples on how to couple different adapters, but also to verify functionality of adapter and solver installations by executing them.
-
-All v202211.0 test can be run successfully inside our NixOS VM and also in a \texttt{nix develop} environment, except \texttt{turek-hron-fsi2} and \texttt{flow-over-heated-plate-steady-state}.
-
-We could not manage to build \texttt{swak4foam}, an add-on to OpenFOAM that is needed in the \texttt{turek-hron-fsi2} tutorial, from source.
+The preCICE tutorials included in the distribution cannot only be used as reference examples on how to couple different solvers, but also to verify functionality of solver and adapter installations by executing them.
+To this end, we execute all 23 tutorials such that each participant option is at least run once. We do not compare results, but only test whether the cases complete.
+We test inside the NixOS VM and also in an ad-hoc shell environment provided by the \texttt{nix develop} command.
+All tests pass except the \textit{Turek-Hron FSI3} and the \textit{flow over heated plate steady state} cases.
+We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial.
 The official preCICE VM uses a prebuilt version of swak4foam, however also patching the prebuilt binary did not work.
-
-The second simulation we were not able to run is \texttt{flow-over-heated-plate-steady-state}.
-This testcase needs code-aster to work, that we managed to package as already mentioned.
-However, when running the simulation after executing the manual steps as written in the documentation, there occurs a segmentation fault, crashing code-aster when the python code tries to call a Fortran subroutine.
+The dependency is, however, no longer required in the v202404.0 release of the distribution.
+The flow over heated plate steady state tutorial requires Code\_Aster.
+Even though we manage to package the solver, we observe a segmentation fault at runtime when the python code tries to call a Fortran subroutine -- a problem we did not manage to debug.
 

### 2024-05-16 paper/paper.tex
Rewrite and shorten conclusion

@@
 
-\section{Conclusion and Outlook}
+\section{Conclusion}
 
-In conclusion, the case study was a success, as it has shown that packaging of scientific software with Nix is possible, but it is not as trivial as expected.
-Once the software is packaged, it is quite easy to be used and combined in environments with other packaged softwares without conflicts.
-The packaged adapters and solvers work well together with the preCICE library and simulations can be run successfully, so this specific case study has shown the potential of Nix in the domain of scientific software.
-More research has to be done on the topic, yet we are positive about possible future results in Nix abilities and its flexibility in similar domains.
+We investigated on whether Nix und NixOS are a potential solution to enable full reproducibility of research software environments.
+As a case study, we tried to package and test all components of the preCICE ecosystem -- a very heterogeneous set of legacy software packages, which is, however, representative for the state of research software today in our opinion.
 
-The majority of the packages required some solution for a fundamental issue within the scientific software or its build system that is not solvable for none experienced Nix users, making it hard for people to use it even for their own software.
-Nix itself also is partly at fault, because of its unique nature, that every piece of software lies within \texttt{/nix/store} and that each store path is its own-isolated tree based on the Filesystem Hierarchy Standard (FHS).
-But it is worth mentioning that not only Nix would benefit from more streamlined build systems and packaging, because both Spack and EasyBuild do something similar for their packages, where each package also lies in its own FHS.
-The work of the xSDK people and their policy might be a step in the right direction but looking at the current list of software that is available within the latest xSDK release, 26 libraries, indicates that it is an uphill battle that might take several more years, especially if we consider that the first release happened in 2016.
+We were able to package nearly all components and run nearly all test cases showing the potential of Nix in the domain of scientific software.
+Many packages required workarounds, however, which might be difficult to achieve for none-experienced Nix users.
+Nix also comes with a few peculiarities, which complicate workarounds.
+In fact, every piece of software has to lie within \texttt{/nix/store} and each path therein is its own-isolated tree based on the Filesystem Hierarchy Standard (FHS).
 
-While this case study and this comparison has provided valuable insights into the topic of reproducible software, there are still numerous unanswered questions that merit further investigation.
-This includes a full evaluation if there are performance improvements using a full impure system rebuild, with \texttt{-march=native} compared to only rebuilding the package at hand.
-It would also be interesting to look at trying to reproduce a nixpkgs commit from 10 years ago, and if there might be issues and incompatibilities with newer Nix versions, that currently get released in a 6 weeks time windows.
+Most problems, however, can be traced back to a lack of standardization in research software, especially regarding build systems. 
+Software packages that follow best practices for their programming language are also straightforward to package.
+The xSDK initiative defined such standardization policies for math software.
+Components of the preCICE ecosystem that already adhere to this standard, were among the easiest ones to package -- even though xSDK  ultimately targets Spack and not Nix.
 
-Last but not least, it would be interesting to investigate Nix and NixOS on HPC even more.
-It would be interesting to see how an HPC cluster completely build on NixOS could improve maintaining the HPC cluster, with slurm already being available as a nixpkgs package and module providing a systemd service.
-
-Finally, it would be interesting to investigate a nix store that is shared between users and how it might optimize the amount of storage needed for a cluster as each user is not required to rebuild the same package in their respective home directories.
-
-This summary shows, how diverse and complex Nix is, and how much more opportunities there are to be researched in the future.
+This paper is the result of a student research project and a streamlined version of its report~\cite{TODO}.
+Beyond this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
 

### 2024-05-16 paper/paper.tex
Add student report as ref
Remove dangling literature
Resort and tweak overview table

@@
 
+This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
+
 \section{Background}
@@
     \bfseries Package & \bfseries Type & \bfseries Build system & \bfseries Version \\ \hline
-    precice & library & cmake & 2.5.0 \\ \hline
-    fortran-module & bindings & make & 9e3f405 \\ \hline
-    julia & bindings & julia & 2.5.0 \\ \hline
-    matlab & bindings & matlab build script & 2.5.0.0 \\ \hline
-    pyprecice & bindings & setup.py & 2.5.0.1 \\ \hline
-    config-visualizer & tooling & setup.py & 60f2165 \\ \hline
-    precice-aste & tooling & cmake & 3.0.0 \\ \hline
-    dealii & solver & cmake & 9.4.1 \\ \hline
-    precice-dealii-adapter & adapter & cmake & dbb25bea \\ \hline
-    precice-calculix-adapter & solver + adapter & makefile & 2.20 \\ \hline
-    fenics & solver library& cmake + setup.py & 2019.1.0 \\ \hline
-    precice-fenics-adapter & adapter & setup.py & 1.4.0 \\ \hline
-    precice-dune & adapter & scripts + cmake & 2.8.0 \\ \hline
-    dune-fem & solver library & setup.py & 2.8.0.0 \\ \hline
-    code-aster & solver & custom build system in python & 14.6.0 \\ \hline
-    openfoam & solver & wmake/Allwmake & 2206 \\ \hline
-    openfoam-adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
-    precice-su2 & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
-    nutils & solver library & setup.py & 7.0.0 \\ \hline
+    preCICE & library & cmake & 2.5.0 \\ \hline
+    Fortran module & bindings & make & 9e3f405 \\ \hline
+    Julia bindings & bindings & julia & 2.5.0 \\ \hline
+    Matlab bindings & bindings & matlab build script & 2.5.0.0 \\ \hline
+    Python bindings & bindings & setup.py & 2.5.0.1 \\ \hline
+    ASTE & tooling & cmake & 3.0.0 \\ \hline
+    Config visualizer & tooling & setup.py & 60f2165 \\ \hline
+    preCICE-CalculiX adapter & solver + adapter & makefile & 2.20 \\ \hline
+    Code\_Aster & solver & custom system in python & 14.6.0 \\ \hline
+    Code\_Aster-preCICE adapter & adapter & custom system in python & ce995e0 \\ \hline
+    deal.II & solver & cmake & 9.4.1 \\ \hline
+    preCICE-deal.II adapter & adapter & cmake & dbb25bea \\ \hline
+    DUNE-FEM & solver library & setup.py & 2.8.0 \\ \hline
+    preCICE-DUNE adapter& adapter & scripts + cmake & 5f2364d \\ \hline
+    FEniCS & solver library& cmake + setup.py & 2019.1.0 \\ \hline
+    preCICE-FEniCS adapter & adapter & setup.py & 1.4.0 \\ \hline
+    Nutils & solver library & setup.py & 7.0.0 \\ \hline
+    OpenFOAM & solver & wmake/Allwmake & 2206 \\ \hline
+    preCICE-OpenFOAM adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
+    preCICE-SU2 adapter & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
   \end{tabular}
@@
 
-\subsubsection{Fortran bindings}
+\subsubsection{Fortran module}
 
@@
 
-This paper is the result of a student research project and a streamlined version of its report~\cite{TODO}.
+This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
 Beyond this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
@@
 
-\nocite{*}
 \bibliographystyle{eceasst}

### 2024-05-16 paper/paper.tex
Tweak matlab section

@@
     Julia bindings & bindings & julia & 2.5.0 \\ \hline
-    Matlab bindings & bindings & matlab build script & 2.5.0.0 \\ \hline
+    MATLAB bindings & bindings & matlab build script & 2.5.0.0 \\ \hline
     Python bindings & bindings & setup.py & 2.5.0.1 \\ \hline
@@
 
-\subsubsection{Matlab bindings}
+\subsubsection{MATLAB bindings}
 
-There were several attempts to package matlab\footnote{\url{https://de.mathworks.com/products/matlab.html}} for NixOS in the past\footnote{\url{https://github.com/NixOS/nixpkgs/issues/56887}}, yet there was no success so far.
-This might be due to the fact, that matlab needs to be installed by running an installation wizard that downloads files during the installation process and verifies the licence.
-Also, as matlab is not installed in the VM, we decided to not proceed in packaging the software.
+There were several attempts to package MATLAB\footnote{\url{https://de.mathworks.com/products/matlab.html}} for NixOS in the past\footnote{\url{https://github.com/NixOS/nixpkgs/issues/56887}}, yet there was no success so far.
+This might be due to the fact, that MATLAB needs to be installed by running an installation wizard that downloads files during the installation process and verifies the licence.
+Also, as MATLAB is not installed in the preCICE VM, we do not proceed in packaging the software.
 

### 2024-05-17 paper/paper.tex
fix(paper): Trailing whitespace
fix(paper): Fix meaning of FHS in Nix

@@
 
+% \begin{sloppypar}
+% \end{sloppypar}
+
 \section{Discussion}
@@
 Nix also comes with a few peculiarities, which complicate workarounds.
-In fact, every piece of software has to lie within \texttt{/nix/store} and each path therein is its own-isolated tree based on the Filesystem Hierarchy Standard (FHS).
+In fact, every piece of software has to lie within \texttt{/nix/store} and each path therein is its own-isolated tree inspired by the Filesystem Hierarchy Standard (FHS).
 

### 2024-05-17 paper/paper.tex

@@
 
-FEniCS~\cite{fenics} is already packaged upstream, however, not will all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support.
+FEniCS~\cite{fenics} is already packaged upstream, however, not with all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support and without mshr.
 This is why, we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
-The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code.
+The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code and pin the cython version to \texttt{0.29.34}~\footnote{\url{https://gitlab.com/petsc/petsc/-/issues/1359}}.
 We could not enable tests for PETSc4py, however, because they depend on OpenMPI and the network, which is not available within Nix' build sandbox.
 Afterwards, we specify PETSc4py as dependency for FEniCS and enable the PETSc support.
+Additionally, we need to recompile PETSc with additional features enabled that are not enabled by the upstream Nix package, these include parmetis, hypre and scalapack, this is done using \texttt{overrideAttrs}, a Nix feature that allows to change an existing package.
+Lastly, we also package FEniCS mshr, so all required features for the adapter package are now are available.
 

### 2024-05-17 paper/paper.tex
Final tweaks

@@
 The ecosystem includes diverse legacy solvers, adapters, and language bindings besides the coupling library itself making it a challenging and representative testcase.
-We demonstrate, how to create a reproducible and self-contained environment for this ecosystem and highlight the benefits of using Nix and NixOS.
+We demonstrate how to create a reproducible and self-contained environment for this ecosystem and discuss the benefits and limitations of using Nix and NixOS.
 %In addition, we compare the usability and reproducibility provided by Nix, in the context of preCICE, with two already established high-performance computing (HPC) solutions, Spack and EasyBuild.
@@
 When a user then specifies to install the python3 package, for instance, a traditional package manager could yield version 3.8 today, but version 3.9 in a few months.
-Full reproducibility can, thus, only be achieved by storing the complete image. Altering a single dependency (e.g., a bugfix) causes a rebuild of the image and, thus, destroys reproducibility.
+Full reproducibility can, thus, only be achieved by storing the complete image. Altering a single dependency (e.g., by a bugfix) causes a rebuild of the image and, thus, destroys reproducibility.
 
@@
 With these archiving platforms, experiments can be rerun using technologies such as Docker.
-These platforms are closed source, however, such that the source code cannot be reviewed nor adjusted~\cite{koch2023sustainable}.
+The platforms are closed source, however, such that the source code cannot be reviewed nor adjusted~\cite{koch2023sustainable}.
 
@@
 Similar ideas are followed by the popular high performance computing (HPC) package managers EasyBuild~\cite{easybuil6495863} and Spack~\cite{spack7832814}.
-At FOSDEM 2018, Spack, EasyBuild and Nix were compared with each other~\footnote{\url{https://archive.fosdem.org/2018/schedule/event/installing\_software\_for\_scientists/}} by the EasyBuild release manager.
-While looking into Spack and EasyBuild we found some deficits regarding reproducibility, that they link against core system libraries, that can break independently if those libraries get updated, but the Spack developers are aware of these shortcomings and are already working on improvements~\footnote{\url{https://github.com/spack/spack/issues/39560}}~\footnote{\url{https://github.com/spack/spack/pull/42082}}.
-EasyBuild, on the other hand, even has an option for linking against system libraries called \texttt{osdependencies}~\footnote{\url{https://docs.easybuild.io/writing-easyconfig-files/\#dependency\_specs}}, so packages need to be vetted and potentially updated prior to usage that this option isn't used in any package of the dependency graph if the best possible reproducibility is desired.
+At FOSDEM 2018, Kenneth Hoste compares Spack, EasyBuild, and Nix with each other~\cite{Hoste}.
+Minor deficits regarding reproducibility of Spack and EasyBuild are that they link against core system libraries, which can break independently.
+The Spack developers are aware of these shortcomings and are working on improvements\footnote{\url{https://github.com/spack/spack/issues/39560}}$^,$\footnote{\url{https://github.com/spack/spack/pull/42082}}.
+EasyBuild, on the other hand, even has an option for linking against system libraries called \texttt{osdependencies}\footnote{\url{https://docs.easybuild.io/writing-easyconfig-files/\#dependency\_specs}}.
+Thus, packages need to be vetted and potentially updated prior to usage.
 All three solutions have in common that they rely on scientific software following best practices concerning building and packaging.
@@
 We try to build the complete ecosystem using Nix and run all tutorials.
-
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
@@
 
-We present challenges, solutions, and workarounds for all software packages of the preCICE ecosystem.
-The Nix code for the packages for the case study on the preCICE adapters, solvers and the preCICE VM is available online in a git repository at \url{https://github.com/precice/nix-packages/} with a more in-detail report that also includes a comparison between Nix, Spack and EasyBuild and a evaluation of Nix on HPC~\footnote{\url{https://github.com/precice/nix-packages/releases/tag/initial-paper-release}}.
+We present challenges, solutions, and workarounds for packaging all components of the preCICE ecosystem.
+The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/}}.
 
@@
 
-The preCICE library itself is already available in the nixpkgs repository, so we do not have to package the software.
+The preCICE library itself is already available in the nixpkgs repository\footnote{\url{https://github.com/NixOS/nixpkgs}}, so we do not have to package the software.
 Looking at the package definition and the preCICE source code, the library is quite easy to build and package as it uses CMake as a build system.
@@
 
-The preCICE distribution includes tools around preCICE.
+The preCICE distribution includes several tools around preCICE.
 
@@
 
-preCICE provides bindings\footnote{\url{https://github.com/precice/fortran-module}} to the Fortran programming language.
-We were able to package these by providing custom phases for the \texttt{buildPhase} and the \texttt{installPhase}.
+We were able to package the Fortran module of preCICE by providing custom phases for the build phase and the install phase of Nix.
 The resulting binary could run the example that comes with the module.
@@
 
-The Makefile also has hard-coded locations its dependencies: \texttt{SPOOLES}, \texttt{ARPACK}, and \texttt{YAML}. These need to be replaced with the equivalent \texttt{pkg-config} calls.
-Additionally, there is no \texttt{install} target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.
+The Makefile also has hard-coded locations of its dependencies: SPOOLES, ARPACK, and yaml-cpp. These need to be replaced with the equivalent pkg-config calls.
+Additionally, there is no install target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.
 
@@
 HDF5 is packaged with multiple outputs.
-One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, this is, because most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
-Additionally, Code\_Aster requires Medfile and Scotch, which are both available upstream and can be used as a dependency.
-Another dependency is Metis, which partially builds. Thus, we decide to not replace it, but to fix the remainder of the build.
-The issue is precisely that the \texttt{CMakeLists.txt} for the metis programs hard-codes \texttt{link\_directories} to \path{/home/karypis/local/lib}.
-We unpack the Metis tarball, remove this line with \texttt{sed}, and repack the directory.
+One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, since most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
+Additionally, Code\_Aster requires Medfile and Scotch, which are both available upstream and can be used as dependencies.
+Another dependency is METIS, which partially builds. Thus, we decide to not replace it, but to fix the remainder of the build.
+The issue is precisely that the \texttt{CMakeLists.txt} for the METIS programs hard-codes \texttt{link\_directories} to \path{/home/karypis/local/lib}.
+We unpack the METIS tarball, remove this line with sed, and repack the directory.
 
@@
 We use one of these provided configuration files to write our own custom config file.
-It defines the locations for dependencies, such as Scotch, Metis, Parmetis, or Blacs.
+It defines the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
 Almost all dependencies are available upstream, but we need to recompile Scotch with additional build flags.
@@
 For the build and install process, we need to manually set the \texttt{\$DUNE\_CONTROL\_PATH} and the \texttt{\$DUNE\_PY\_DIR} environment variables. Both variables also need to be set at runtime.
-Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix' sandboxed builds.
+Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix's sandboxed builds.
 
@@
 
-FEniCS~\cite{fenics} is already packaged upstream, however, not with all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support and without mshr.
+FEniCS~\cite{fenics} is already packaged upstream, however, not with all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support and mshr, the FEniCS mesh generator.
 This is why, we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
 The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code and pin the cython version to \texttt{0.29.34}~\footnote{\url{https://gitlab.com/petsc/petsc/-/issues/1359}}.
-We could not enable tests for PETSc4py, however, because they depend on OpenMPI and the network, which is not available within Nix' build sandbox.
+We could not enable tests for PETSc4py, however, because they depend on OpenMPI and the network, which is not available within Nix's build sandbox.
 Afterwards, we specify PETSc4py as dependency for FEniCS and enable the PETSc support.
-Additionally, we need to recompile PETSc with additional features enabled that are not enabled by the upstream Nix package, these include parmetis, hypre and scalapack, this is done using \texttt{overrideAttrs}, a Nix feature that allows to change an existing package.
-Lastly, we also package FEniCS mshr, so all required features for the adapter package are now are available.
+Additionally, we need to recompile PETSc with additional features enabled that are not enabled by the upstream Nix package.
+These include ParMETIS, HYPRE and ScaLAPACK.
+We use \texttt{overrideAttrs}, a Nix feature that allows us to change an existing package.
+Lastly, we also package mshr, so all required features for the adapter package are now are available.
 
@@
 OpenFOAM uses its own custom build system called \texttt{wmake}, which is typically called with a \texttt{Allwmake} wrapper script.
-The build system sets 36 environment variables, one step at a time by checking several parameters, e.g.
-the CPU architecture or the location of the source code.
+The build system sets 36 environment variables, one step at a time by checking several parameters, e.g.~the CPU architecture or the location of the source code.
 Also during runtime, these variables need to be set, either by a wrapper script or by manually sourcing a file inside the installation directory of OpenFOAM.
 For Nix, these properties are unfortunate. To compile OpenFOAM with Nix, we need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
-We also use a shell script called \texttt{set-openfoam-vars}, which exports the environment variables to the current shell session.
+We also use a shell script, which exports the environment variables to the current shell session.
 For simplicity, we hard-code all parameters, such as the processor architecture.
@@
 The preCICE-OpenFOAM adapter is an OpenFOAM function object, requiring OpenFOAM and wmake as build inputs.
-The environment variables have to be set again through the \texttt{set-openfoam-vars} script.
+The environment variables have to be set again through the shell script mentioned above.
 Afterwards, we set the target directory for the adapter to \texttt{\$out/lib/} as the adapter is a shared library. Building then works successfully.
@@
 The current preCICE VM\footnote{\url{https://precice.org/installation-vm.html}} is built on top of Vagrant\footnote{\url{https://www.vagrantup.com/}} using Ubuntu as a base image.
-When it is run for the first time, it further installs software, compiles programs, and the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
+When run for the first time, it further installs software, compiles programs, and clones the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
 This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the \textit{latest} version, e.g.
@@
 The preCICE tutorials included in the distribution cannot only be used as reference examples on how to couple different solvers, but also to verify functionality of solver and adapter installations by executing them.
-To this end, we execute all 23 tutorials such that each participant option is at least run once. We do not compare results, but only test whether the cases complete.
+To this end, we execute all 23 tutorials of the distribution such that each participant option is at least run once. We do not compare results, but only test whether the cases complete.
 We test inside the NixOS VM and also in an ad-hoc shell environment provided by the \texttt{nix develop} command.
@@
 We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial.
-The official preCICE VM uses a prebuilt version of swak4foam, however also patching the prebuilt binary did not work.
+The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
 The dependency is, however, no longer required in the v202404.0 release of the distribution.
@@
 We investigated on whether Nix und NixOS are a potential solution to enable full reproducibility of research software environments.
-As a case study, we tried to package and test all components of the preCICE ecosystem -- a very heterogeneous set of legacy software packages, which is, however, representative for the state of research software today in our opinion.
+As a case study, we tried to package and test all components of the preCICE ecosystem -- a very heterogeneous set of legacy software packages, which is, however, representative for the state of research software today in our experience.
 
@@
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
-Beyond this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
+Beyond the content of this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
 

### 2024-05-17 paper/paper.tex
chore: set-dune-vars script
Shorten DUNE sentence
feat: nutils section

@@
 Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix's sandboxed builds.
+The two environment variables must also be set at runtime.
+This can be achieved by sourcing the \texttt{set-dune-vars} script, which we provide.
 
@@
 Finally, to find the preCICE installation, we need to set the configuration flags \texttt{--with-include} and \texttt{--with-lib}.
+
+\subsubsection{nutils}
+
+The nutils~\cite{nutils} project can be build using \texttt{buildPythonPackage}, one test needs to be disabled, the rest of the testsuite passed without any issues.
+While the preCICE distribution ships with nutils v7.0, one tutorial case requires v6.3 because of the usage of a deprecated function.
 

### 2024-05-17 paper/paper.tex
Extend tutorials section with all pitfalls

@@
 We test inside the NixOS VM and also in an ad-hoc shell environment provided by the \texttt{nix develop} command.
-All tests pass except the \textit{Turek-Hron FSI3} and the \textit{flow over heated plate steady state} cases.
-We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial.
+All tests pass except the \textit{Turek-Hron FSI3}, parts of the \textit{partitioned heat conduction}, parts of the \textit{perpendicular flap}, and the \textit{flow over heated plate steady state} cases.
+We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial and the OpenFOAM participants of the partitioned heat conduction tutorial.
 The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
 The dependency is, however, no longer required in the v202404.0 release of the distribution.
+We could, moreover, not package the third-party solids4foam solver, which is one of many options in the perpendicular flap tutorial. 
 The flow over heated plate steady state tutorial requires Code\_Aster.

### 2024-05-17 paper/paper.tex
Tweak Nutils section

@@
 
-\subsubsection{nutils}
+\subsubsection{Nutils}
 
-The nutils~\cite{nutils} project can be build using \texttt{buildPythonPackage}, one test needs to be disabled, the rest of the testsuite passed without any issues.
-While the preCICE distribution ships with nutils v7.0, one tutorial case requires v6.3 because of the usage of a deprecated function.
+The Nutils~\cite{nutils} solver can be build using \texttt{buildPythonPackage}.
+One test needs to be disabled, but the rest of the testsuite passes without issues.
+There is no dedicated Nutils-preCICE adapter, but preCICE is typically directly called from Nutils application scripts.
 

### 2024-08-08 paper/paper.tex
Sort solver subsubsections
feat(paper): Set link to code to tag
feat(paper): Add the feedback provided by peer review

@@
 
-\keywords{Reproducibility, Nix, NixOS, Spack, EasyBuild, preCICE, HPC}
+\keywords{Reproducibility, Nix, NixOS}
 
@@
 
-There are, moreover, commercial, domain specifig solutions to achieve reproducibility, e.g., CodeOcean\footnote{\url{https://codeocean.com/}} mainly for bioinformatics or Weights and Biases\footnote{\url{https://wandb.ai/site}} for machine learning.
+There are, moreover, commercial, domain specific solutions to achieve reproducibility, e.g., CodeOcean\footnote{\url{https://codeocean.com/}} mainly for bioinformatics or Weights and Biases\footnote{\url{https://wandb.ai/site}} for machine learning.
 With these archiving platforms, experiments can be rerun using technologies such as Docker.
@@
 At FOSDEM 2018, Kenneth Hoste compares Spack, EasyBuild, and Nix with each other~\cite{Hoste}.
-Minor deficits regarding reproducibility of Spack and EasyBuild are that they link against core system libraries, which can break independently.
+Minor deficits regarding reproducibility of Spack and EasyBuild are that they link against core system libraries, e.g., glibc, which can break independently.
 The Spack developers are aware of these shortcomings and are working on improvements\footnote{\url{https://github.com/spack/spack/issues/39560}}$^,$\footnote{\url{https://github.com/spack/spack/pull/42082}}.
@@
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
+Beyond the content of this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
 
@@
 In our case study, we target the distribution v2211.0~\cite{preciceDistribution}.
-To execute simulations, such as the included tutorials, requires the actual solvers next to the distribution.
+To execute simulations, such as the included tutorials, the actual solvers are required besides the distribution.
 These are included, for example, in the preCICE VM\footnote{\url{https://precice.org/installation-vm.html}} and also part of our case study.
@@
     preCICE & library & cmake & 2.5.0 \\ \hline
-    Fortran module & bindings & make & 9e3f405 \\ \hline
-    Julia bindings & bindings & julia & 2.5.0 \\ \hline
-    MATLAB bindings & bindings & matlab build script & 2.5.0.0 \\ \hline
-    Python bindings & bindings & setup.py & 2.5.0.1 \\ \hline
     ASTE & tooling & cmake & 3.0.0 \\ \hline
     Config visualizer & tooling & setup.py & 60f2165 \\ \hline
+    MATLAB bindings & bindings & matlab build script & 2.5.0.0 \\ \hline
+    Fortran module & bindings & make & 9e3f405 \\ \hline
+    Python bindings & bindings & setup.py & 2.5.0.1 \\ \hline
+    Julia bindings & bindings & julia & 2.5.0 \\ \hline
     preCICE-CalculiX adapter & solver + adapter & makefile & 2.20 \\ \hline
@@
 We present challenges, solutions, and workarounds for packaging all components of the preCICE ecosystem.
-The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/}}.
+The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission}}.
 
@@
 
-The preCICE library itself is already available in the nixpkgs repository\footnote{\url{https://github.com/NixOS/nixpkgs}}, so we do not have to package the software.
+The preCICE library itself is already available in the nixpkgs repository\footnote{\url{https://github.com/NixOS/nixpkgs/blob/nixos-24.05/pkgs/development/libraries/precice/default.nix\#L41}}, so we do not have to package the software.
 Looking at the package definition and the preCICE source code, the library is quite easy to build and package as it uses CMake as a build system.
@@
 
-The preCICE distribution includes several tools around preCICE.
+The preCICE distribution includes two tools that enhance the preCICE user experience.
 
-\subsubsection{ASTE}
-
-ASTE\footnote{\url{https://github.com/precice/aste}} stands for Artificial Solver Testing Environment.
+One of the tools is ASTE\footnote{\url{https://github.com/precice/aste}}, which stands for Artificial Solver Testing Environment.
 It is a thin wrapper around the preCICE API, which allows, for instance, testing of data mapping with real geometries.
-ASTE requires VTK9\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
+ASTE requires VTK\footnote{\url{https://vtk.org/}}, a visualization toolkit, which is built without python support in the nixpkgs repository to reduce compilation time.
 This feature is needed for ASTE, however.
@@
 
-\subsubsection{Config visualizer}
-
-The preCICE config visualizer\footnote{\url{https://github.com/precice/config-visualizer}}, as the name suggests, visualizes preCICE configuration files.
-Similarly to preCICE itself, the tool is already packaged upstream.
-As a python package, it is easy to handle with Nix.
+The second tool is the preCICE config visualizer.
+As the name suggests, it visualizes preCICE configuration files.
+Similarly to preCICE itself, the tool is already packaged upstream\footnote{\url{https://github.com/NixOS/nixpkgs/blob/nixos-24.05/pkgs/tools/misc/precice-config-visualizer/default.nix\#L25}}.
+This tool is easy to handle with Nix, as it is a python package.
 
@@
 There were several attempts to package MATLAB\footnote{\url{https://de.mathworks.com/products/matlab.html}} for NixOS in the past\footnote{\url{https://github.com/NixOS/nixpkgs/issues/56887}}, yet there was no success so far.
-This might be due to the fact, that MATLAB needs to be installed by running an installation wizard that downloads files during the installation process and verifies the licence.
-Also, as MATLAB is not installed in the preCICE VM, we do not proceed in packaging the software.
+This might be due to the fact, that MATLAB needs to be installed by running an installation wizard that downloads files during the installation process and verifies the license.
+As MATLAB has some licensing issues and is also not installed in the preCICE VM, we do not proceed in packaging the software.
 
@@
 
-The \texttt{pyprecice} package is already available upstream, but did not compile as it is not able to find the python module \texttt{pkgconfig}.
+The \texttt{pyprecice} package is already available upstream, but does not compile as it is not able to find the python module \texttt{pkgconfig}.
 After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.
-This solution was contributed to the upstream nixpkgs repository.
+This solution was contributed to the upstream nixpkgs repository\footnote{\url{https://github.com/NixOS/nixpkgs/commit/fd8962162ac21be59fc3a05fb6a250eeab6b2bec}}.
 
@@
 
-Julia~\cite{bezanson2017julia} support in NixOS is currently still in its early stages and cannot be declaratively be defined by Nix\footnote{\url{https://github.com/NixOS/nixpkgs/issues/20649}}.
+Julia~\cite{bezanson2017julia} support in NixOS is currently still in its early stages and cannot be declaratively defined by Nix at the current state\footnote{\url{https://github.com/NixOS/nixpkgs/issues/20649}}.
 
@@
 The Makefile also has hard-coded locations of its dependencies: SPOOLES, ARPACK, and yaml-cpp. These need to be replaced with the equivalent pkg-config calls.
-Additionally, there is no install target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to \texttt{\$out}.
+Additionally, there is no install target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to the \texttt{\$out} placeholder, which maps to the resulting store path.
 
@@
   pname = "precice-dealii-adapter";
-  version = "unstable-2022-09-23";
+  version = "unstable-2022-09-23"; # could also be a git tag
 
-  src = fetchFromGitHub {
+  src = fetchFromGitHub { # Defining where to get the source from
     owner = "precice"; repo = "dealii-adapter";
@@
 
-  nativeBuildInputs = [ cmake ];
-  buildInputs = [ precice dealii ];
+  nativeBuildInputs = [ cmake ]; # build-time dependencies
+  buildInputs = [ precice dealii ]; # run-time dependencies
 
@@
 
+  # nix' default phases can be overwritten such as:
   installPhase = ''
@@
 We clone all of the required DUNE repositories into a directory and additionally clone the DUNE-preCICE adapter into the same directory.
-For the build and install process, we need to manually set the \texttt{\$DUNE\_CONTROL\_PATH} and the \texttt{\$DUNE\_PY\_DIR} environment variables. Both variables also need to be set at runtime.
+For the build and install process, we need to manually set the \texttt{\$DUNE\_CONTROL\_PATH} and the \texttt{\$DUNE\_PY\_DIR} environment variables.
 Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix's sandboxed builds.
@@
 FEniCS~\cite{fenics} is already packaged upstream, however, not with all features the FEniCs-preCICE adapter~\cite{Rodenberg2021} relies on, in particular PETSc support and mshr, the FEniCS mesh generator.
-This is why, we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
+This is why we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
 The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code and pin the cython version to \texttt{0.29.34}~\footnote{\url{https://gitlab.com/petsc/petsc/-/issues/1359}}.
@@
 
+\subsubsection{Nutils}
+
+The Nutils~\cite{nutils} solver can be built using \texttt{buildPythonPackage}.
+One test needs to be disabled, but the rest of the testsuite passes without issues.
+There is no dedicated Nutils-preCICE adapter, but preCICE is typically directly called from Nutils application scripts.
+
 \subsubsection{OpenFOAM}
@@
 For Nix, these properties are unfortunate. To compile OpenFOAM with Nix, we need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
-We also use a shell script, which exports the environment variables to the current shell session.
+We also use a shell script which exports the environment variables to the current shell session.
 For simplicity, we hard-code all parameters, such as the processor architecture.
@@
 
-\subsubsection{Nutils}
-
-The Nutils~\cite{nutils} solver can be build using \texttt{buildPythonPackage}.
-One test needs to be disabled, but the rest of the testsuite passes without issues.
-There is no dedicated Nutils-preCICE adapter, but preCICE is typically directly called from Nutils application scripts.
+Since preCICE distribution v202404, this adapter is replaced by a python-based solution
 
@@
 The current preCICE VM\footnote{\url{https://precice.org/installation-vm.html}} is built on top of Vagrant\footnote{\url{https://www.vagrantup.com/}} using Ubuntu as a base image.
-When run for the first time, it further installs software, compiles programs, and clones the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
-This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the \textit{latest} version, e.g.
+When provisioned for the first time, it further installs software, compiles programs, and clones the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
+This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the \textit{latest} version, i.e.
 the \texttt{main} branch of the preCICE repository.
@@
 We use Nix to define a VM image with NixOS as a base, which can be built reproducibly.
-The image contains all components of the case study and some additional custom tools that can only be seen in the official VM, such as the \texttt{preciceToPNG} command.
+The image contains all components of the case study and some additional custom tools that can only be seen in the official VM\footnote{\url{https://github.com/precice/vm}}, such as the \texttt{preciceToPNG} command.
 Also, we generate an iso and a Vagrant VirtualBox file of the VM as additional outputs.
@@
 The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
-The dependency is, however, no longer required in the v202404.0 release of the distribution.
+The dependency is, however, no longer required in the v202404.0 release\footnote{\url{https://precice.org/installation-distribution.html#v24040}} of the preCICE distribution.
 We could, moreover, not package the third-party solids4foam solver, which is one of many options in the perpendicular flap tutorial.
@@
 We were able to package nearly all components and run nearly all test cases showing the potential of Nix in the domain of scientific software.
-Many packages required workarounds, however, which might be difficult to achieve for none-experienced Nix users.
+Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.
 Nix also comes with a few peculiarities, which complicate workarounds.
@@
 Components of the preCICE ecosystem that already adhere to this standard, were among the easiest ones to package -- even though xSDK  ultimately targets Spack and not Nix.
-
-This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
-Beyond the content of this paper, the report also includes a detailed comparison of Nix to Spack and EasyBuild, and a discussion on how to use Nix on HPC systems.
 

### 2024-08-12 paper/paper.tex
feat: integrate some more feedback into paper and fix tests in fenics adapter

@@
 More precisely, we disable HDF5, as it is already packaged in the upstream nixpkgs repository, even at the required version 5.1.10.
-HDF5 is packaged with multiple outputs.
-One package, \texttt{out}, contains the libraries and binaries and another, \texttt{dev} contains the header files, since most of the time, users installing HDF5 only need the libraries and binaries, and it helps to reduce installation size of this rather large package.
 Additionally, Code\_Aster requires Medfile and Scotch, which are both available upstream and can be used as dependencies.
@@
 Mumps uses a Makefile and can be configured using Makefile.inc.
-By default, it provides a few configurations for systems such as Ubuntu.
-We use one of these provided configuration files to write our own custom config file.
-It defines the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
+We write our own custom Makefile.inc by providing the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
 Almost all dependencies are available upstream, but we need to recompile Scotch with additional build flags.
@@
 The issue is precisely, that the custom build system does not correctly calculate the python site-packages directory.
-It unconditionally slices the first 3 chars from \texttt{sys.version}, which works for version 3.9.x, but not for version 3.10.x.
+It unconditionally slices the first 3 chars from \texttt{sys.version}, which works for python version 3.9.x, but not for version 3.10.x.
 After patching, Code\_Aster successfully installed into \texttt{\$out/14.6/}, so we move around some files until we have a valid directory structure with \texttt{\$out/bin}, \texttt{\$out/lib}, providing symlinks for \texttt{\$out/14.6/} and \texttt{\$out/stable/}.
@@
 Building the FEniCS-preCICE adapter is now directly possible with \texttt{buildPythonPackage}.
-All but two of the adapter's tests run successfully, which means, however, that we need to disable the tests.
+We then validate the correctness by including an import check as well as successfully running the full test suite.
 
@@
 When provisioned for the first time, it further installs software, compiles programs, and clones the preCICE tutorials\footnote{\url{https://github.com/precice/tutorials/}}.
-This inherently breaks reproducibility as the VM fetches files from the internet that are at this very point in time the \textit{latest} version, i.e.
-the \texttt{main} branch of the preCICE repository.
+This means that we only have reproducibility when fetching the VM from vagrant cloud~\footnote{\url{https://app.vagrantup.com/precice/boxes/precice-vm}}, as we can not rebuild the VM from scratch, because the provision step currently fetches files from the internet that are at this very point in time the \textit{latest} version, i.e.
+the \texttt{main} branch of the preCICE repository rather than a pinned tag or commit.
+Making us dependent on vagrant cloud to continue to serve the box.
 
@@
 The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
-The dependency is, however, no longer required in the v202404.0 release\footnote{\url{https://precice.org/installation-distribution.html#v24040}} of the preCICE distribution.
+The dependency is, however, no longer required in the v202404.0 release\footnote{\url{https://precice.org/installation-distribution.html\#v24040}} of the preCICE distribution.
 We could, moreover, not package the third-party solids4foam solver, which is one of many options in the perpendicular flap tutorial.

### 2024-08-13 paper/paper.tex
feat: more feedback

@@
 We try to build the complete ecosystem using Nix and run all tutorials.
+In section~\ref{sec:background} the background to Nix and preCICE is introduced, section~\ref{sec:results} presents the case study on packaging the preCICE ecosystem including its challenges and workarounds.
+After this we discuss the results in section~\ref{sec:discussion}, followed by the conclusion of this paper in section~\ref{sec:conclusion}.
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
@@
 \section{Background}
+\label{sec:background}
 
@@
 \section{Results}
+\label{sec:results}
 
@@
 \section{Discussion}
+\label{sec:discussion}
 
@@
 
-The xSDK policies also requires portable installations.
-Several components in the preCICE ecosystem require installation into \texttt{\$HOME} or any required files to be located at specific locations.
+The xSDK policies also requires portable installations. Several components in the preCICE ecosystem require installation into \texttt{\$HOME} or any required files to be located at specific locations.
 Both, the CalculiX-preCICE adapter and the SU2-preCICE adapter require the original source code to be present as they add new features on top or patch the source code.
-The current solution of cloning the code at the specific location works, but it might be better to look into Git submodules.
+The current solution that these build systems require cloning the code to a specific directory works, but it might be better to look into Git submodules, as this would result in a more portable solution.
 This would moreover allow pinning the version of the original code compared to only documenting it.
@@
 \section{Conclusion}
+\label{sec:conclusion}
 

### 2024-08-13 paper/paper.tex
feat: push remaining feedback changes

@@
 
-We were able to package nearly all components and run nearly all test cases showing the potential of Nix in the domain of scientific software.
+Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselfs, four were already packaged upstream and two were not packageable as described earlier.
+Additionally we were able to run 52 participant examples out of the 61 total examples found in the preCICE tutorials, showing the potential of Nix in the domain of scientific software.
 Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.

### 2024-08-14 paper/paper.tex
feat: integrate last feedback

@@
     \includegraphics[width=1\textwidth]{figures/coreutils.png}
-    \caption{The dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -Grankdir=LR -ocoreutils.png}}
+    \caption{The full dependency graph of the coreutils package, generated by \texttt{nix-store --query --graph \$(nix build nixpkgs\#coreutils --print-out-paths) | dot -Tpng -Grankdir=LR -ocoreutils.png}}
     \label{fig:nix-graph}
@@
 As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
-Inputs of these functions are parameters such as the package version or the location of the source code.
+Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
+Nix then offers built-in functions for common buildsystems and languages such as cmake, rust, go and python by implementing different phases.
+There are a wide range of phases that always get executed in the same order and package maintainers can then decide weather the default phase is sufficient or needs to be overriden in order to produce a working build.
+The most common phases to override are the \texttt{configurePhase}, which is mostly used for cmake or autotools configuration, the \texttt{buildPhase}, \texttt{installPhase} and the \texttt{checkPhase} which can for example be used to run unit tests.
+It is also worth mentioning that most phases have a pre and post step, which users can use to do slight adjustments when they need to override excisting packages.
 Evaluating these functions with the same inputs yields the same outputs.
@@
 All contents, including source files and resulting build artifacts, are stored inside the Nix store.
-Per default, the Nix store resides in the path \texttt{/nix/store} on the file system.
+Per default, the Nix store resides in the path \path{/nix/store} on the file system.
 The naming scheme for packages includes the above mentioned hash, for preCICE v2.5.0, for example,
@@
     \includegraphics[width=0.5\textwidth]{figures/precice.png}
-    \caption{Software setup of preCICE from the preCICE website \url{https://precice.org}}
+    \caption{Software setup of preCICE from the preCICE website \url{https://precice.org}.
+It shows how a custom solver can be coupled using an adapter with existing adapters for OpenFOAM and CalculiX.}
     \label{fig:precice}
@@
 
-We present challenges, solutions, and workarounds for packaging all components of the preCICE ecosystem.
-The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission}}.
+In this section we will look and discuss all components within the preCICE ecosystem.
+We then present challenges, solutions and workarounds that occur when packaging these components.
+This section is structured as follows, we first look into preCICE itself, followed by additional tooling and bindings of the preCICE library in other languages.
+Afterwards we present solutions for all solvers and their respected adapters.
+Lastly we show that Nix additionally be utilized to generate iso files as well as qemu and Vagrant VM images.
+The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission}} and can be used to download and run any of the presented packages.
 
@@
 
-The \texttt{pyprecice} package is already available upstream, but does not compile as it is not able to find the python module \texttt{pkgconfig}.
+The pyprecice package is already available upstream, but does not compile as it is not able to find the python module pkgconfig.
 After adding this single dependency as an input to the package definition, the python package builds and can be used as expected.
@@
 The remaining two dependencies, Mumps and Tfel are not available upstream and need to be packaged.
-Mumps uses a Makefile and can be configured using Makefile.inc.
+Mumps uses a Makefile and can be configured using \texttt{Makefile.inc}.
 We write our own custom Makefile.inc by providing the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
@@
     \begin{minted}{nix}
-{ lib, stdenv, fetchFromGitHub, cmake, precice,
-  dealii, enable3d ? false }:
+{
+  # Inputs including nix functions/packages or additional flags
+  lib, stdenv, fetchFromGitHub, cmake,  dealii,
+  # Optional features that users can enable when overriding
+  enable3d ? false }:
 stdenv.mkDerivation rec {
@@
 
-  nativeBuildInputs = [ cmake ]; # build-time dependencies
-  buildInputs = [ precice dealii ]; # run-time dependencies
+  # dependencies only available at build-time
+  nativeBuildInputs = [ cmake ];
+  # dependencies that also need to be installed at run-time
+  # mostly libraries needed for linking and for execution
+  buildInputs = [ precice dealii ];
 
@@
   installPhase = ''
+    # $out contains the final path inside the nix store.
+    # The resulting build files are copied into this directory
     mkdir -p $out/bin && cp elasticity $out/bin/elasticity
@@
 This is why we need to package PETSc4py, the python bindings package of PETSc, which is not available upstream.
-The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code and pin the cython version to \texttt{0.29.34}~\footnote{\url{https://gitlab.com/petsc/petsc/-/issues/1359}}.
+The build process uses the internal \texttt{buildPythonApplication} build tool and runs successfully once we add the option \texttt{build\_src --force} to rebuild cython code and pin the cython version to 0.29.34~\footnote{\url{https://gitlab.com/petsc/petsc/-/issues/1359}}.
 We could not enable tests for PETSc4py, however, because they depend on OpenMPI and the network, which is not available within Nix's build sandbox.
@@
 
-OpenFOAM uses its own custom build system called \texttt{wmake}, which is typically called with a \texttt{Allwmake} wrapper script.
+OpenFOAM uses its own custom build system called wmake, which is typically called with a Allwmake wrapper script.
 The build system sets 36 environment variables, one step at a time by checking several parameters, e.g.~the CPU architecture or the location of the source code.
@@
 This means that we only have reproducibility when fetching the VM from vagrant cloud~\footnote{\url{https://app.vagrantup.com/precice/boxes/precice-vm}}, as we can not rebuild the VM from scratch, because the provision step currently fetches files from the internet that are at this very point in time the \textit{latest} version, i.e.
-the \texttt{main} branch of the preCICE repository rather than a pinned tag or commit.
+the main branch of the preCICE repository rather than a pinned tag or commit.
 Making us dependent on vagrant cloud to continue to serve the box.
@@
 
-Finally, hard-coding libraries to \texttt{/usr/lib} and \texttt{/usr/include} might work on one system, but might make it hard or even impossible to install a package on a different system.
-Configurable solutions, such as \texttt{pkg-config}, should always be preferred and work out of the box with common paths such as \texttt{/usr/lib}.
+Finally, hard-coding libraries to \path{/usr/lib} and \path{/usr/include} might work on one system, but might make it hard or even impossible to install a package on a different system.
+Configurable solutions, such as pkg-config, should always be preferred and work out of the box with common paths such as \path{/usr/lib}.
 
@@
 Nix also comes with a few peculiarities, which complicate workarounds.
-In fact, every piece of software has to lie within \texttt{/nix/store} and each path therein is its own-isolated tree inspired by the Filesystem Hierarchy Standard (FHS).
+In fact, every piece of software has to lie within \path{/nix/store} and each path therein is its own-isolated tree inspired by the Filesystem Hierarchy Standard (FHS).
 

### 2024-08-14 paper/paper.tex
chore: fix some typos

@@
 \usepackage{xcolor}
-%\newcommand\todo[1]{\textcolor{red}{TODO #1}}
 
@@
 Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
-Nix then offers built-in functions for common buildsystems and languages such as cmake, rust, go and python by implementing different phases.
-There are a wide range of phases that always get executed in the same order and package maintainers can then decide weather the default phase is sufficient or needs to be overriden in order to produce a working build.
-The most common phases to override are the \texttt{configurePhase}, which is mostly used for cmake or autotools configuration, the \texttt{buildPhase}, \texttt{installPhase} and the \texttt{checkPhase} which can for example be used to run unit tests.
-It is also worth mentioning that most phases have a pre and post step, which users can use to do slight adjustments when they need to override excisting packages.
+Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
+There are a wide range of phases that always get executed in the same order and package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working build.
+The most common phases to override are the \texttt{configurePhase}, which is mostly used for CMake or Autotools configuration, the \texttt{buildPhase}, \texttt{installPhase} and the \texttt{checkPhase} which can for example be used to run unit tests.
+It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.
 Evaluating these functions with the same inputs yields the same outputs.
@@
     \bfseries Package & \bfseries Type & \bfseries Build system & \bfseries Version \\ \hline
-    preCICE & library & cmake & 2.5.0 \\ \hline
-    ASTE & tooling & cmake & 3.0.0 \\ \hline
+    preCICE & library & CMake & 2.5.0 \\ \hline
+    ASTE & tooling & CMake & 3.0.0 \\ \hline
     Config visualizer & tooling & setup.py & 60f2165 \\ \hline
@@
     Julia bindings & bindings & julia & 2.5.0 \\ \hline
-    preCICE-CalculiX adapter & solver + adapter & makefile & 2.20 \\ \hline
+    preCICE-CalculiX adapter & solver + adapter & Makefile & 2.20 \\ \hline
     Code\_Aster & solver & custom system in python & 14.6.0 \\ \hline
     Code\_Aster-preCICE adapter & adapter & custom system in python & ce995e0 \\ \hline
-    deal.II & solver & cmake & 9.4.1 \\ \hline
-    preCICE-deal.II adapter & adapter & cmake & dbb25bea \\ \hline
+    deal.II & solver & CMake & 9.4.1 \\ \hline
+    preCICE-deal.II adapter & adapter & CMake & dbb25bea \\ \hline
     DUNE-FEM & solver library & setup.py & 2.8.0 \\ \hline
-    preCICE-DUNE adapter& adapter & scripts + cmake & 5f2364d \\ \hline
-    FEniCS & solver library& cmake + setup.py & 2019.1.0 \\ \hline
+    preCICE-DUNE adapter& adapter & scripts + CMake & 5f2364d \\ \hline
+    FEniCS & solver library& CMake + setup.py & 2019.1.0 \\ \hline
     preCICE-FEniCS adapter & adapter & setup.py & 1.4.0 \\ \hline
@@
     preCICE-OpenFOAM adapter & adapter & wmake/Allwmake & 1.2.1 \\ \hline
-    preCICE-SU2 adapter & solver + adapter & patch script + autoconf & 6.0.0 \\ \hline
+    preCICE-SU2 adapter & solver + adapter & patch script + Autoconf & 6.0.0 \\ \hline
   \end{tabular}
@@
 This might be due to the fact, that MATLAB needs to be installed by running an installation wizard that downloads files during the installation process and verifies the license.
-As MATLAB has some licensing issues and is also not installed in the preCICE VM, we do not proceed in packaging the software.
+As MATLAB has some licensing issues and is also not installed in the preCICE VM, we do not proceed in packaging the preCICE bindings for the language.
 
@@
 
-The preCICE adapter for the structural mechanics code CalculiX~\cite{Uekermann2017_Adapters} requires the original CalculiX source code and provides a new makefile in the adapter repository.
-By default, this makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
+The preCICE adapter for the structural mechanics code CalculiX~\cite{Uekermann2017_Adapters} requires the original CalculiX source code and provides a new Makefile in the adapter repository.
+By default, this Makefile expects the source code to reside in \texttt{\$HOME}, but it is possible to override this location with a make variable.
 It then builds the original code and the adapter code together, resulting in a combined binary.
@@
 The Makefile also has hard-coded locations of its dependencies: SPOOLES, ARPACK, and yaml-cpp. These need to be replaced with the equivalent pkg-config calls.
-Additionally, there is no install target provided by the makefile, thus the resulting binary needs to be installed manually by copying it to the \texttt{\$out} placeholder, which maps to the resulting store path.
+Additionally, there is no install target provided by the Makefile, thus the resulting binary needs to be installed manually by copying it to the \texttt{\$out} placeholder, which maps to the resulting store path.
 
@@
 Mumps uses a Makefile and can be configured using \texttt{Makefile.inc}.
-We write our own custom Makefile.inc by providing the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
+We write our own custom \texttt{Makefile.inc} by providing the locations for dependencies, such as Scotch, METIS, ParMETIS, or Blacs.
 Almost all dependencies are available upstream, but we need to recompile Scotch with additional build flags.
@@
 DUNE~\cite{bastian2020dune} uses a combination of CMake files and custom build scripts making the build process in Nix tedious.
-There is an ongoing migration from autotools to CMake for DUNE, which might improve the situation in the future.
+There is an ongoing migration from Autotools to CMake for DUNE, which might improve the situation in the future.
 Another minor issue is that the DUNE project is rather a collection of repositories than a monorepo.
@@
 Additionally, we need to patch the python install process because the current CMake file tries to access the internet with \texttt{pip install}, which is not allowed in Nix's sandboxed builds.
-The two environment variables must also be set at runtime.
+The two environment variables must also be set at run-time.
 This can be achieved by sourcing the \texttt{set-dune-vars} script, which we provide.
@@
 The build system sets 36 environment variables, one step at a time by checking several parameters, e.g.~the CPU architecture or the location of the source code.
-Also during runtime, these variables need to be set, either by a wrapper script or by manually sourcing a file inside the installation directory of OpenFOAM.
+Also during run-time, these variables need to be set, either by a wrapper script or by manually sourcing a file inside the installation directory of OpenFOAM.
 For Nix, these properties are unfortunate. To compile OpenFOAM with Nix, we need to patch the shebangs\footnote{\url{https://foldoc.org/shebang}} of wmake to make it run during the build.
@@
 To this end, the adapter provides a script to run before the build phase.
-We do this in Nix' patch phase. Afterwards, \texttt{stdenv} automatically recognizes that SU2 uses autotools for building and runs the configure phase.
+We do this in Nix' patch phase. Afterwards, \texttt{stdenv} automatically recognizes that SU2 uses Autotools for building and runs the configure phase.
 Finally, to find the preCICE installation, we need to set the configuration flags \texttt{--with-include} and \texttt{--with-lib}.
@@
 The flow over heated plate steady state tutorial requires Code\_Aster.
-Even though we manage to package the solver, we observe a segmentation fault at runtime when the python code tries to call a Fortran subroutine -- a problem we did not manage to debug.
+Even though we manage to package the solver, we observe a segmentation fault at run-time when the python code tries to call a Fortran subroutine -- a problem we did not manage to debug.
 
@@
 OpenFOAM and Allwmake are prime examples of this issue.
-Package managers have to understand a whole new build system for a single package -- a fact that is often ignored when considering the trade-off between maintaining a custom build system versus pivoting to an industry standard, such as CMake or autotools.
+Package managers have to understand a whole new build system for a single package -- a fact that is often ignored when considering the trade-off between maintaining a custom build system versus pivoting to an industry standard, such as CMake or Autotools.
 Such standard build tools generally provide interfaces for dependency management and optional features.
@@
 The authors of xSDK\footnote{\url{https://xsdk.info/}} come to a similar conclusion.
-They provide a list of package policies~\cite{xSDK2023}, which, for example, specify that packages must have an appropriate build system, which includes CMake and autoconf as examples.
+They provide a list of package policies~\cite{xSDK2023}, which, for example, specify that packages must have an appropriate build system, which includes CMake and Autotools as examples.
 If a package supports all policies, they can be added to the growing list of packages in xSDK.
@@
 
-Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselfs, four were already packaged upstream and two were not packageable as described earlier.
+Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream and two were not packageable as described earlier.
 Additionally we were able to run 52 participant examples out of the 61 total examples found in the preCICE tutorials, showing the potential of Nix in the domain of scientific software.

### 2024-08-14 paper/paper.tex
style: simplify phase description in section 2

@@
 Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
-There are a wide range of phases that always get executed in the same order and package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working build.
-The most common phases to override are the \texttt{configurePhase}, which is mostly used for CMake or Autotools configuration, the \texttt{buildPhase}, \texttt{installPhase} and the \texttt{checkPhase} which can for example be used to run unit tests.
+Phases are bash scripts and are already predefined for each buildsystem and always execute in the same order.
+Package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working package.
+The most important phases are the \texttt{configurePhase}, which is mostly used for buildsystem configuration.
+The \texttt{buildPhase} defines the build steps e.g., calling \texttt{make} or \texttt{cargo build}.
+Afterwards the \texttt{installPhase} is executed which copies the results into the output directory.
+Additionally a \texttt{checkPhase} can be enabled which can be used to run unit tests to insure that the package works as expected.
 It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.

### 2024-08-15 paper/paper.tex
fix(paper): typo, buildsystem -> build system, remove double space

@@
 Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
-Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
-Phases are bash scripts and are already predefined for each buildsystem and always execute in the same order.
-Package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working package.
-The most important phases are the \texttt{configurePhase}, which is mostly used for buildsystem configuration.
+Nix then offers built-in functions for common build system and languages such as CMake, Rust, Go and python by implementing different phases.
+Phases are bash scripts and are already predefined for each build system and always execute in the same order.
+Package maintainers can then decide whether the default phase is sufficient or needs to be overridden in order to produce a working package.
+
+For building packages, the four key phases are the \texttt{configurePhase}, the \texttt{buildPhase}, the \texttt{installPhase} and the \texttt{checkPhase}.
+During the \texttt{configurePhase}, the build system is configured.
 The \texttt{buildPhase} defines the build steps e.g., calling \texttt{make} or \texttt{cargo build}.
@@
 It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.
+
 Evaluating these functions with the same inputs yields the same outputs.

### 2024-08-15 paper/paper.tex
feat(paper): Add paragraph to conclusion

@@
 
+Despite encountering challenges in packaging the numerous software tools, we were able to observe the adaptability of the Nix package manager and the pleasant user experience that results upon completion of the packaging process.
+The results we have obtained indicate that Nix is a suitable solution for replicating identical software environments while simplifying the use and reuse of package definitions for users.
+We would love to see more software being packaged for Nix, so that even difficult to install software will be accessible to researchers with ease.
+
 \acknowledge{

### 2024-08-18 paper/paper.tex
Iterate on review changes

--- 2024-08-14 paper/paper.tex
@@
 We try to build the complete ecosystem using Nix and run all tutorials.
-In section~\ref{sec:background} the background to Nix and preCICE is introduced, section~\ref{sec:results} presents the case study on packaging the preCICE ecosystem including its challenges and workarounds.
-After this we discuss the results in section~\ref{sec:discussion}, followed by the conclusion of this paper in section~\ref{sec:conclusion}.
+Section~\ref{sec:background} briefly introduces the background of Nix and preCICE. Section~\ref{sec:results} then presents the case study including challenges, workarounds, and open problems.
+Afterwards, we discuss the results in section~\ref{sec:discussion}, followed by the conclusions in section~\ref{sec:conclusion}.
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
@@
 As Nix is purely functional, it relies on functions that, without side-effects, realize the package derivations.
-Inputs of these functions are parameters such as the package version, the location of the source code and its dependencies.
-Nix then offers built-in functions for common buildsystems and languages such as CMake, Rust, Go and python by implementing different phases.
-Phases are bash scripts and are already predefined for each buildsystem and always execute in the same order.
-Package maintainers can then decide weather the default phase is sufficient or needs to be overridden in order to produce a working package.
-The most important phases are the \texttt{configurePhase}, which is mostly used for buildsystem configuration.
-The \texttt{buildPhase} defines the build steps e.g., calling \texttt{make} or \texttt{cargo build}.
-Afterwards the \texttt{installPhase} is executed which copies the results into the output directory.
-Additionally a \texttt{checkPhase} can be enabled which can be used to run unit tests to insure that the package works as expected.
-It is also worth mentioning that most phases have a \texttt{pre} and \texttt{post} phases, which users can use to make slight adjustments when they need to override existing packages.
-Evaluating these functions with the same inputs yields the same outputs.
+Inputs of these functions are parameters such as the package version or the location of the source code and its dependencies.
+Nix offers built-in functions for common build systems and languages, such as CMake, Rust, Go, and Python by implementing different phases.
+Phases are bash scripts and are always executed in the same order. They are predefined for each build system, but can be overwritten if necessary.
+
+The four key phases are the configure phase, the build phase, the install phase, and the check phase.
+During the configure phase, the build system is configured.
+The build phase defines the build steps, such as calling \texttt{make} or \texttt{cargo build}.
+Afterwards, the install phase copies the results into the output directory.
+The check phase can be enabled to run any tests.
+Most phases have a pre and post step, which can be uses to make adjustments.
+
+Evaluating these Nix functions with the same inputs yields the same outputs.
 This is a critical factor for Nix's reproducibility.
@@
     \caption{Software setup of preCICE from the preCICE website \url{https://precice.org}.
-It shows how a custom solver can be coupled using an adapter with existing adapters for OpenFOAM and CalculiX.}
+The setup shows a custom solver coupled to OpenFOAM and CalculiX for conjugate-heat transfer (CHT) and fluid-structure interaction (FSI) simulations as an example.}
     \label{fig:precice}
@@
 
-\section{Results}
+\section{Case study}
 \label{sec:results}
 
-In this section we will look and discuss all components within the preCICE ecosystem.
-We then present challenges, solutions and workarounds that occur when packaging these components.
-This section is structured as follows, we first look into preCICE itself, followed by additional tooling and bindings of the preCICE library in other languages.
-Afterwards we present solutions for all solvers and their respected adapters.
-Lastly we show that Nix additionally be utilized to generate iso files as well as qemu and Vagrant VM images.
+We now study the packaging of all components of the preCICE ecosystem one by one, including challenges, solutions, and workarounds.
+We first look at preCICE itself, followed by additional tooling and bindings.
+Afterwards, we study all solvers and their respected adapters.
+Next, we show that Nix can also be used to generate iso files as well as qemu and Vagrant VM images. Finally, we try running all preCICE tutorials.
 The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission}} and can be used to download and run any of the presented packages.
@@
 Both, the CalculiX-preCICE adapter and the SU2-preCICE adapter require the original source code to be present as they add new features on top or patch the source code.
-The current solution that these build systems require cloning the code to a specific directory works, but it might be better to look into Git submodules, as this would result in a more portable solution.
-This would moreover allow pinning the version of the original code compared to only documenting it.
+The current solution of requiring to clone the code into a specific directory works, but it might be better to look into Git submodules, as this would result in a more portable solution.
+This would, moreover, allow pinning the version of the original code compared to only documenting it.
 
@@
 
-Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream and two were not packageable as described earlier.
-Additionally we were able to run 52 participant examples out of the 61 total examples found in the preCICE tutorials, showing the potential of Nix in the domain of scientific software.
+Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream, and two were not packageable in their current state.
+We were able to run 52 participant examples out of the 61 total examples found in the preCICE tutorials.
 Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.
@@
 
+Despite encountering challenges in packaging the numerous software tools, we were able to observe the adaptability of the Nix package manager.
+Our results indicate that, in general, Nix is a suitable solution for reproducing software environments.
+
 \acknowledge{
-We thankfully acknowledge the funding by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy EXC 2075 -- 390740016 and the support by the Stuttgart Center for Simulation Science (SimTech).
+We thankfully acknowledge the funding by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) under Germany's Excellence Strategy EXC 2075 -- 390740016 and under project number 528693298,  and the support by the Stuttgart Center for Simulation Science (SimTech).
 }

### 2024-08-20 paper/paper.tex
feat: consistent naming for section

@@
 Section~\ref{sec:background} briefly introduces the background of Nix and preCICE. Section~\ref{sec:results} then presents the case study including challenges, workarounds, and open problems.
-Afterwards, we discuss the results in section~\ref{sec:discussion}, followed by the conclusions in section~\ref{sec:conclusion}.
+Afterwards, we discuss the results in Section~\ref{sec:discussion}, followed by the conclusions in Section~\ref{sec:conclusion}.
 This paper is the result of a student research project and a streamlined version of its report~\cite{studentreport}.
@@
 
-This chapter gives a short introduction into Nix and preCICE.
+This Section gives a short introduction into Nix and preCICE.
 

### 2024-08-20 paper/paper.tex
feat(paper): Minor revisions

@@
 
-The four key phases are the configure phase, the build phase, the install phase, and the check phase.
-During the configure phase, the build system is configured.
-The build phase defines the build steps, such as calling \texttt{make} or \texttt{cargo build}.
-Afterwards, the install phase copies the results into the output directory.
-The check phase can be enabled to run any tests.
+For building packages, the four key phases are the \texttt{configurePhase}, the \texttt{buildPhase}, the \texttt{installPhase} and the \texttt{checkPhase}.
+During the \texttt{configurePhase}, the build system is configured.
+The \texttt{buildPhase} defines the build steps, such as calling \texttt{make} or \texttt{cargo build}.
+Afterwards, the \texttt{installPhase} copies the results into the output directory.
+The \texttt{checkPhase} can be enabled to run any tests.
 Most phases have a pre and post step, which can be uses to make adjustments.
@@
 Despite encountering challenges in packaging the numerous software tools, we were able to observe the adaptability of the Nix package manager.
-Our results indicate that, in general, Nix is a suitable solution for reproducing software environments.
+Our results indicate that, in general, Nix is a suitable solution for reproducing software environments and a viable solution to package management in a scientific domain.
 

### 2024-08-20 paper/paper.tex
fix(paper): There are 60 participant cases
feat(paper): Make Nix' phases consistent

@@
 
-We were able to package the Fortran module of preCICE by providing custom phases for the build phase and the install phase of Nix.
+We were able to package the Fortran module of preCICE by providing custom contents for the \texttt{buildPhase} and the \texttt{installPhase} of Nix.
 The resulting binary could run the example that comes with the module.
@@
 
-After successfully packaging all dependencies, the build phase of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
+After successfully packaging all dependencies, the \texttt{buildPhase} of Code\_Aster completes, but the installation part has another issue, which also affects new versions of Ubuntu and other distributions using a python version greater than 3.9.
 There is a forum post\footnote{\url{https://forum.code-aster.org/public/d/26475-problem-installing-code-aster-version-14-6/11}} without resolution, so we need to manually patch the bug.
@@
 These could be parametrized, however, based on the Nix inputs to allow for optimized builds.
-We use the script during the build phase to source all variables such that, for example, \texttt{OPENFOAM\_SRC\_PATH} points to the default location \texttt{/build/openfoam}.
+We use the script during the \texttt{buildPhase} to source all variables such that, for example, \texttt{OPENFOAM\_SRC\_PATH} points to the default location \path{/build/openfoam}.
 Afterwards, \texttt{./Allwmake -j -q} is sufficient to start the build.
-The install phase then copies the necessary files and directories to \texttt{\$out}, replaces the mock \texttt{OPENFOAM\_SRC\_PATH} by the value of \texttt{\$out}, and creates a wrapper for the \texttt{openfoam} shell script.
+The \texttt{installPhase} then copies the necessary files and directories to \texttt{\$out}, replaces the mock \texttt{OPENFOAM\_SRC\_PATH} by the value of \texttt{\$out}, and creates a wrapper for the \texttt{openfoam} shell script.
 
@@
 The SU2-preCICE adapter~\cite{Uekermann2017_Adapters} patches and extends the original CFD code SU2~\cite{su2}.
-To this end, the adapter provides a script to run before the build phase.
-We do this in Nix' patch phase. Afterwards, \texttt{stdenv} automatically recognizes that SU2 uses Autotools for building and runs the configure phase.
+To this end, the adapter provides a script to run before the \texttt{buildPhase}.
+We do this in Nix' \texttt{patchPhase}, which is responsible for patching the source before running the \texttt{buildPhase}.
+Afterwards, \texttt{stdenv} automatically recognizes that SU2 uses Autotools for building and runs the \texttt{configurePhase}.
 Finally, to find the preCICE installation, we need to set the configuration flags \texttt{--with-include} and \texttt{--with-lib}.
@@
 Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream, and two were not packageable in their current state.
-We were able to run 52 participant examples out of the 61 total examples found in the preCICE tutorials.
+We were able to run 52 participant examples out of the 60 total examples found in the preCICE tutorials.
 Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.

### 2024-08-20 paper/paper.tex
Fix spelling mistake (Section)
feat(paper): Explain the coupled solvers

@@
 
-This Section gives a short introduction into Nix and preCICE.
+This section gives a short introduction into Nix and preCICE.
 
@@
 The preCICE tutorials included in the distribution cannot only be used as reference examples on how to couple different solvers, but also to verify functionality of solver and adapter installations by executing them.
-To this end, we execute all 23 tutorials of the distribution such that each participant option is at least run once. We do not compare results, but only test whether the cases complete.
+To this end, we execute all 23 tutorials of the distribution such that each coupled solver option is at least run once. We do not compare results, but only test whether the cases complete.
 We test inside the NixOS VM and also in an ad-hoc shell environment provided by the \texttt{nix develop} command.
 All tests pass except the \textit{Turek-Hron FSI3}, parts of the \textit{partitioned heat conduction}, parts of the \textit{perpendicular flap}, and the \textit{flow over heated plate steady state} cases.
-We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial and the OpenFOAM participants of the partitioned heat conduction tutorial.
+We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial and the OpenFOAM coupled solver participant of the partitioned heat conduction tutorial.
 The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
@@
 Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream, and two were not packageable in their current state.
-We were able to run 52 participant examples out of the 60 total examples found in the preCICE tutorials.
+In total, there are 22 tutorials in the preCICE v2211 distribution, many having several different coupled solver examples, resulting in a total of 60 coupled solver examples.
+We were able to run 52 coupled solvers out of the total number found in the preCICE tutorials.
 Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.

### 2024-08-21 paper/paper.tex
Simplify coupled solvers

@@
 All tests pass except the \textit{Turek-Hron FSI3}, parts of the \textit{partitioned heat conduction}, parts of the \textit{perpendicular flap}, and the \textit{flow over heated plate steady state} cases.
-We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial and the OpenFOAM coupled solver participant of the partitioned heat conduction tutorial.
+We could not build swak4foam, an add-on to OpenFOAM, which is needed in the Turek-Hron FSI3 tutorial and the OpenFOAM solver of the partitioned heat conduction tutorial.
 The official preCICE VM uses a prebuilt version of swak4foam, but also patching the prebuilt binary did not work.
@@
 Out of the 20 components of the preCICE distribution, we were able to package 14 components ourselves, four were already packaged upstream, and two were not packageable in their current state.
-In total, there are 22 tutorials in the preCICE v2211 distribution, many having several different coupled solver examples, resulting in a total of 60 coupled solver examples.
-We were able to run 52 coupled solvers out of the total number found in the preCICE tutorials.
+In total, there are 22 tutorials in the preCICE v2211 distribution, many having several different coupled solvers, resulting in a total of 60 coupled solvers.
+We were able to run 52 out of these.
 Many packages required workarounds, however, which might be difficult to achieve for non-experienced Nix users.

### 2024-08-21 paper/paper.tex
chore: migrate to lstlisting so we can diff those

@@
 
-\usepackage{minted}
 \usepackage{subfig}
 
+\usepackage{listings}
 \usepackage{xcolor}
+
+\definecolor{identifiercolor}{rgb}{0,0,0}
+\definecolor{keywordcolor}{HTML}{1a177d}
+\definecolor{commentcolor}{HTML}{3d7a7a}
+\definecolor{stringcolor}{HTML}{ba2121}
+\definecolor{keywordcolor2}{HTML}{870000}
+\definecolor{keywordcolor3}{HTML}{008000}
+\definecolor{keywordcolor4}{HTML}{666666}
+\lstdefinelanguage{Nix}{
+  % Anything betweeen $ becomes LaTeX math mode
+  mathescape=false,
+  % Comments may or not include Latex commands
+  texcl=false,
+  keywords=[1]{pname,version,src,owner,rev,repo,sha256,nativeBuildInputs,buildInputs,installPhase},
+  % Comments delimiters, we do turn this off for the manual
+  comment=[l]{\#},
+  % Spaces are not displayed as a special character
+  showstringspaces=false,
+  % String delimiters
+  morestring=[b]",
+  morestring=[d]'',
+  % Size of tabulations
+  tabsize=2,
+  % Enables ASCII chars 128 to 255
+  extendedchars=false,
+  % Case sensitivity
+  sensitive=true,
+  % Automatic breaking of long lines
+  breaklines=false,
+  % Default style for listings
+  basicstyle=\ttfamily,
+  % Position of captions is bottom
+  captionpos=b,
+  % flexible columns
+  columns=[l]fixed,
+  % Style for (listings') identifiers
+  identifierstyle={\ttfamily\color{identifiercolor}},
+  % Style for declaration keywords
+  keywordstyle=[1]{\ttfamily\color{keywordcolor}},
+  % Style for strings
+  stringstyle={\ttfamily\color{stringcolor}},
+  % Style for comments
+  commentstyle={\ttfamily\color{commentcolor}},
+  alsoletter={=}
+  %moredelim=**[is][\ttfamily\color{red}]{/&}{&/},
+  morekeywords=[2]{false},
+  keywordstyle=[2]{\ttfamily\color{keywordcolor2}},
+  morekeywords=[3]{rec},
+  keywordstyle=[3]{\ttfamily\color{keywordcolor3}},
+  morekeywords=[4]{=},
+  keywordstyle=[4]{\ttfamily\color{keywordcolor4}},
+}[keywords,comments,strings]
 
@@
     \normalsize
-    \begin{minted}{nix}
+    \begin{lstlisting}[language=Nix]
 {
@@
   # Optional features that users can enable when overriding
-  enable3d ? false }:
+  enable3d ? false
+}:
 stdenv.mkDerivation rec {
@@
 }
-    \end{minted}
+    \end{lstlisting}
     \caption{Nix code to package the deal.II-preCICE adapter}

### 2024-08-21 paper/paper.tex
feat: set updated link for tag

@@
 Next, we show that Nix can also be used to generate iso files as well as qemu and Vagrant VM images. Finally, we try running all preCICE tutorials.
-The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission}} and can be used to download and run any of the presented packages.
+The Nix code for all components is available online\footnote{\url{https://github.com/precice/nix-packages/releases/tag/deRSE24-paper-submission-v2}} and can be used to download and run any of the presented packages.
 
